### 2020-09-22 paper.Rmd
push paper draft start

---
title: "Catchy title here"
authors: Robert(1), James (2)
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(out.width = "75%",
                      dev = "png",
                      fig.align = "center",
                      echo = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(naniar)
library(visdat)
library(patchwork)

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-09-23 paper.Rmd
push
push

@@
 ---
-title: "Catchy title here"
+title: "waRne - Putting cricket analytics in a spin"
 authors: Robert(1), James (2)
@@
     keep_md: true
-bibliography: references.bib
+
 ---
@@
 
-## R Markdown
+## Abstract
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+The importance of reproducibility, and the related issue of open access to data, has received a lot of recent attention.
+Momentum on these issues is gathering in the sports analytics community.
+While cricket is the second largest commercial sport in the world, unlike other popular international sports, there has been no mechanism for the public to access comprehensive statistics on players and teams.
+Expert commentary currently relies heavily on data that isn't made readily accessible and this produces an unnecessary barrier for the development of an inclusive sports analytics community.
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+## Introduction
 
-```{r cars}
-summary(cars)
-```
+## What is waRne
 
-## Including Plots
+We need to think about what it is waRne should be?
 
-You can also embed plots, for example:
+1.  provide the csv we used for datajam and then people just add games to it
 
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
+2.  people can not have the csv but just download random tournaments
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+## Why cricket needs reproducibility
+
+maybe include stuff around cricviz interesting tweets but fans can recreate or investigate further.
+
+## Cool cricket examples
+
+-   Dhoni spin/pace/yorkr
+
+    -   shows off joins to player information and extracting information from play by play
+
+-   espncricinfo articles but instead of ipl do same tables for big bash
+
+-   recreate stats from espncricinfo like this [page](https://www.espncricinfo.com/series/19297/statistics/1187665/new-zealand-vs-england-1st-t20i-england-in-new-zealand-2019-20
+)
+
+| Things to do                 | person                                                                   |
+|------------------------------|--------------------------------------------------------------------------|
+| clean scrapers/csv for waRne | James                                                                    |
+| data dictionary              | Robert                                                                   |
+| examples                     | Robert - start (assume same csv being used) - cool cricket example stuff |
+
+: things to do

### 2020-09-23 code/scraping_functions.R
Added the code to make the master.csv

@@ -1,0 +1,1052 @@
+# This function will alter the input year to the correct format for the url in ESPNcricinfo
+get_season <- function(input_year){
+	
+	print(paste("Scraping the", input_year, "data"))
+	
+	# year is a 4 or 8 length numeric value to indicate either a single year, or a season that crossed two
+	# years (ie 20102011)
+	
+	# There is a specific way of formatting season that span 1 or 2 years, and this function will alter
+	# the input, say 2019 or 20192020 to 2019 or 2019%2F20. 
+	
+	if(floor(log10(input_year))+1==4){
+		
+		year <- as.character(input_year)
+		
+		half_season <- as.character(input_year+1)
+		half_season <- paste0(str_sub(half_season, 1, 1), "F", str_sub(half_season, 3, 4), sep = "")
+		
+		not_year <- paste(year, "%", half_season, sep = "")
+		
+		links <- 
+			tourney_list_site_data %>%
+			html_nodes(".season-links") %>%
+			html_nodes("a") %>%
+			html_attr(name = "href") %>% 
+			as_tibble() %>% 
+			rename("links" = value) %>%
+			dplyr::filter(grepl(year, links)) %>%
+			dplyr::filter(!grepl(not_year, links)) %>%
+			pull()
+		
+	} else if(floor(log10(input_year))+1==8){
+		
+		first_year <- str_sub(input_year, 1, 4)
+		second_year <- paste0(str_sub(input_year, 5, 5), "F", str_sub(input_year, 7, 8), sep = "")
+		year <- paste(first_year, "%", second_year, sep= "")
+		
+		links <- 
+			tourney_list_site_data %>%
+			html_nodes(".season-links") %>%
+			html_nodes("a") %>%
+			html_attr(name = "href") %>% 
+			as_tibble() %>% 
+			rename("links" = value) %>%
+			dplyr::filter(grepl(year, links)) %>%
+			pull()
+		
+		Sys.sleep(sample(5:15, 1))
+		
+	} else{
+		
+		print("Problem with input values, please enter either a single year (i.e. 2019), or a season spanning two seasons (i.e. 20192020)")
+	}
+	
+	return(links)
+	
+}
+
+# The 2 inputs are:
+# year_input - just the season. For example 2019, or 20182019. Must be input as a numeric input, 4 or 8
+# characters in length
+# root_path - You need to select a root path for the data to be saved on your computer. For example,
[…]

### 2020-09-23 code/run_scraper.R
ok try this rob

@@ -1,0 +1,61 @@
+library(tidyverse)
+library(RSelenium)
+library(XML)
+library(rvest)
+library(htmltools)
+library(lubridate)
+library(magrittr)
+source("/Users/James/James/Reproducibility-in-Cricket/code/scraping_functions.R")
+
+# You will need to create an image to do this in docker
+system('docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome')
+# So if it does break, its a good idea to clear your container. the scraper will start a fresh one
+system('docker kill $(docker ps -q)')
+# Run this loop, it catches a few errors that are common and can be ignored. If the tournament has
+# a broken link, remove_master underneth willt take the year and row number as an argument and 
+# remove it for you
+complete = 0
+while(complete!=1){
+	commentary <- try(grab_tournament_links(2018, "/Users/James/James/Reproducibility-in-Cricket"))
+	if(class(commentary)=="try-error"){
+		error_type <- attr(commentary, "condition")
+		print(class(error_type))
+		print(error_type$message)
+	}
+	if(error_type$message == "HTTP error 504."){
+		print("504 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "HTTP error 500."){
+		print("500 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "HTTP error 501."){
+		print("501 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "HTTP error 502."){
+		print("502 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "HTTP error 503."){
+		print("503 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "object 'starting_date' not found"){
+		print("I think it failed to get the results page loaded, 
+          taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
[…]

### 2020-09-23 code/scraping_functions.R

@@ -1,5 +1,5 @@
 # This function will alter the input year to the correct format for the url in ESPNcricinfo
-get_season <- function(input_year){
+get_season <- function(input_year, input_link){
 	
 	print(paste("Scraping the", input_year, "data"))
 	
@@ -19,7 +19,7 @@ get_season <- function(input_year){
 		not_year <- paste(year, "%", half_season, sep = "")
 		
 		links <- 
-			tourney_list_site_data %>%
+			input_link %>%
 			html_nodes(".season-links") %>%
 			html_nodes("a") %>%
 			html_attr(name = "href") %>% 
@@ -36,7 +36,7 @@ get_season <- function(input_year){
 		year <- paste(first_year, "%", second_year, sep= "")
 		
 		links <- 
-			tourney_list_site_data %>%
+			input_link %>%
 			html_nodes(".season-links") %>%
 			html_nodes("a") %>%
 			html_attr(name = "href") %>% 
@@ -88,7 +88,8 @@ grab_tournament_links <- function(year_input, root_path){
 	}
 	# Create the master spreadsheet for the year if it doesn't exist. If it does just read it in
 	if(!file.exists(master_path)){
-		year_link<- get_season(year_input)
+		year_link<- get_season(year_input, 
+													 tourney_list_site_data)
 		full_link <- paste(web_prefix, year_link, sep = "")
 		# The get season function just changes the input year from a numeric to the strange format they
 		# use on cric info. It then reads the season archive page, finds the correct link, and outputs
@@ -1049,4 +1050,47 @@ extract_year <- function(this_date){
 	out <- list(year = year, 
 							date_minus_year = this_date)
 	return(year)
+}
+
+# Webscraping is a bit messy, so occasionally it's easier to just drop a bad link from the scraping
+# process altogether. These functions will drop them for you
+remove_commentary_row <- function(year_input, comp_type, row_to_drop){
+	if(comp_type == "pc"){
+		data_path <- "D:\\Cricket_Thesis\\Data"
+		year_path <- paste(data_path, "\\", year_input, sep = "")
+		year_master_spreadsheet_path <- paste(year_path, "\\master.csv", sep = "")
+		year_master_scorecard_spreadsheet_path <- paste(year_path, "\\master_scorecard.csv", sep = "")
[…]

### 2020-09-23 code/run_scraper.R
finished uploading scraping code with comments for users

@@ -8,9 +8,10 @@ library(magrittr)
 source("/Users/James/James/Reproducibility-in-Cricket/code/scraping_functions.R")
 
 # You will need to create an image to do this in docker
-system('docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome')
+# system('docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome')
 # So if it does break, its a good idea to clear your container. the scraper will start a fresh one
 system('docker kill $(docker ps -q)')
+system('docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome')
 # Run this loop, it catches a few errors that are common and can be ignored. If the tournament has
 # a broken link, remove_master underneth willt take the year and row number as an argument and 
 # remove it for you
@@ -58,4 +59,55 @@ while(complete!=1){
 }
 
 # We leave this commented out when not in use to avoid accidentally removing a good row
-# remove_master_row(20202021, "mac", 10)
+# remove_master_row(20202021, "/Users/James/James/Reproducibility-in-Cricket", 10)
+
+# Now we scrape the season for the new data
+complete = 0
+while(complete!=1){
+	commentary <- try(collect_commentary_html_new_new_format(2020, "mac"))
+	if(class(commentary)=="try-error"){
+		error_type <- attr(commentary, "condition")
+		print(class(error_type))
+		print(error_type$message)
+	}
+	if(error_type$message == "HTTP error 504."){
+		print("504 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "HTTP error 500."){
+		print("500 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "HTTP error 501."){
+		print("501 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "HTTP error 502."){
+		print("502 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "HTTP error 503."){
+		print("503 error, taking a 2-4 minute break before trying again")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0
+	}else if(error_type$message == "object 'starting_date' not found"){
[…]

### 2020-09-23 code/scraping_functions.R

@@ -64,6 +64,7 @@ get_season <- function(input_year, input_link){
 # folders, beginning with a Data folder to hold each season
 
 grab_tournament_links <- function(year_input, root_path){
+	system('docker kill $(docker ps -q)')
 	# tourney_list_site is the season archive page for cricinfo
 	tourney_list_site <- "http://www.espncricinfo.com/ci/engine/series/index.html"
 	# basic prefix
@@ -144,6 +145,10 @@ grab_tournament_links <- function(year_input, root_path){
 	# a webpage that changes based on some buttons without altering the url. Docker is a free program
 	# that you will need to install to scrape this website
 	
+	total_tournaments <- 
+		master_spreadsheet %>% 
+		distinct(tournament_name, tournament_type) %>% 
+		nrow()
 	# We'll scrape from the bottom up to the top of the spreadsheet
 	for(i in nrow(master_spreadsheet):1){
 		this_row <- 
@@ -157,9 +162,10 @@ grab_tournament_links <- function(year_input, root_path){
 			 !"commentary_link" %in% colnames(master_spreadsheet) & 
 			 !"status" %in% colnames(master_spreadsheet) & 
 			 !"date" %in% colnames(master_spreadsheet)){
-			print(paste(i, "/", nrow(master_spreadsheet), " remaining: ",
+			print(paste(i, "/", total_tournaments, " remaining: ",
 									this_row$tournament_name, ", a ", this_row$tournament_type, 
 									" tournament in ", year_input, " is being scraped", sep = ""))
+			print(this_row$tournament_link)
 			
 			# The default is to load the "fixtures" page, which has upcoming tournaments. Results has to be
 			# loaded separately
@@ -334,9 +340,10 @@ grab_tournament_links <- function(year_input, root_path){
 			# column.y, which we will coalesce into a single column. Otherwie nothing changes from above
 			if(is.na(this_row$scorecard_link)){
 				# If somehting broke the loop, it will continue from where it left off
-				print(paste(i, "/", nrow(master_spreadsheet), " remaining: ",
+				print(paste(i, "/", total_tournaments, " remaining: ",
 										this_row$tournament_name, ", a ", this_row$tournament_type, 
 										" tournament in ", year_input, " is being scraped", sep = ""))
+				print(this_row$tournament_link)
 				
 				this_tournament_upcoming <- 
[…]

### 2020-09-23 code/run_scraper.R
fixed a problem with the scraping functions breaking

@@ -13,8 +13,8 @@ source("/Users/James/James/Reproducibility-in-Cricket/code/scraping_functions.R"
 system('docker kill $(docker ps -q)')
 system('docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome')
 # Run this loop, it catches a few errors that are common and can be ignored. If the tournament has
-# a broken link, remove_master underneth willt take the year and row number as an argument and 
-# remove it for you
+# a broken link, remove_master underneth will take the year and row number as an argument and 
+# remove it for you. The links are printed for each 
 complete = 0
 while(complete!=1){
 	commentary <- try(grab_tournament_links(2018, "/Users/James/James/Reproducibility-in-Cricket"))
@@ -59,7 +59,7 @@ while(complete!=1){
 }
 
 # We leave this commented out when not in use to avoid accidentally removing a good row
-# remove_master_row(20202021, "/Users/James/James/Reproducibility-in-Cricket", 10)
+# remove_master_row(2018, "/Users/James/James/Reproducibility-in-Cricket", 98)
 
 # Now we scrape the season for the new data
 complete = 0

### 2020-09-23 code/scraping_functions.R

@@ -64,7 +64,7 @@ get_season <- function(input_year, input_link){
 # folders, beginning with a Data folder to hold each season
 
 grab_tournament_links <- function(year_input, root_path){
-	system('docker kill $(docker ps -q)')
+	# system('docker kill $(docker ps -q)')
 	# tourney_list_site is the season archive page for cricinfo
 	tourney_list_site <- "http://www.espncricinfo.com/ci/engine/series/index.html"
 	# basic prefix
@@ -1089,7 +1089,7 @@ remove_master_row <- function(year_input, root_path, row_to_drop){
 # So just use the same root as before, to the folder that will contain all the code, and data, and
 # whatever else you may eventially put in here in your cricket analysis
 collect_commentary_html_new_new_format <- function(year_input, root_path){
-	system('docker kill $(docker ps -q)')
+	# system('docker kill $(docker ps -q)')
 	# Set up the appropriate paths first
 	
 	data_path <- paste(root_path, "/", "Data", sep = "")
@@ -1232,7 +1232,7 @@ collect_commentary_html_new_new_format <- function(year_input, root_path){
 					# Read in commentary page
 					remDr <- RSelenium::remoteDriver(remoteServerAddr = "localhost",
 																					 port = 4444L,
-																					 browserName = "firefox")
+																					 browserName = "chrome")
 					Sys.sleep(5)
 					remDr$open(silent = TRUE)
 					Sys.sleep(5)
@@ -1381,7 +1381,7 @@ collect_commentary_html_new_new_format <- function(year_input, root_path){
 					
 					remDr <- RSelenium::remoteDriver(remoteServerAddr = "localhost",
 																					 port = 4444L,
-																					 browserName = "firefox")
+																					 browserName = "chrome")
 					Sys.sleep(5)
 					remDr$open(silent = TRUE)
 					Sys.sleep(5)

### 2020-09-24 paper.Rmd
fed
fsg
beaten by con artists
proof?

@@
 
+James and I entered a reproducibility competition and the winning entry provided a statistic sort of a gateway into the importance of their ''project'' was was specifically in 2019 Australian Big Bash T20 season.
+The statistic given was in the 2019 big bash season teams that batted first won only $43$%$ of games.
+Thankfully we can try to reproduce that statistic now through the use of waRne
+
+Wikipedia tells us that there were the following:
+
+In the 2018-19 Big Bash League there were 59 matches played. Of which we had 3 DLS results.
+
+| Games                             | Result                        |
+|-----------------------------------|-------------------------------|
+| 21 DEC 2018 - THUNDER VS STARS    | THUNDER WON BY 15 RUNS (D/L)  |
+| 2 FEB 2019 - THUNDER VS SIXERS    | SIXERS WON BY 9 WICKETS (D/L) |
+| 8 JAN 2019 HEAT VS THUNDER        | HEAT BY 15 RUNS (D/L)         |
+| 17th January 2019 Thunder vs Heat | no result                     |
+
+: Games under consideration in 2018-19 Big Bash
+
+```{r, investigation}
+df<-read.csv("data/combined_2020_09_11.csv")
+
+blah<-df%>%
+	dplyr::select(duckworth_lewis,year,match_id, tournament_name, tournament_type, team, inning, binary_result)%>%
+	dplyr::filter(tournament_type%in% c("Twenty20 Internationals", "Twenty20"  ))%>%
+	dplyr::filter(tournament_name%in% c("Twenty20 Big Bash","Big Bash League" ))%>%
+									dplyr::filter(inning==1)%>%
+									dplyr::group_by()%>%
+									distinct()
+# dataset doesn't contain duckworth lewis results. 
+table(blah$year,blah$binary_result)
+````
+```
+
+So here we can see that we in the big bash season 2018-19 we have a current winning percentage batting first of 22/(22+33) =0.4 but this doesn't include the duckworth lewis games or the washed out completely game.
+
+If we were to add in the duckworth lewis games we get 24/(24+34) which is equal to 0.414. Which isn't the result.
+
+We can also clearly see that in the 20192020 big bash season the team batting first wins 39/(39+20) or 0.66 so what could be going on? Did Quantium mean only games played in the actual year 2019?
+
+In that case from the 2018-2019 season we would remove the following games
+
+| Date                 | Game                    | Result                   | Winning Team |
+|----------------------|-------------------------|--------------------------|--------------|
+| 19 December 2018     | Heat Vs Strickers       | Strickers 5 Wickets      | Chased       |
+| 20th December 2018   | Scorchers Vs Renegades  | Renegades by 4 wickets   | Chased       |
+| 21 December 2018     | Thunder vs Stars        | Thunder by 15 runs (D/L) | Batted First |
+| 22 December 2018     | Sixers vs Scorchers     | Sixers by 17 runs        | Batted First |
+| 22 December          | Hurricanes vs Heat      | Hurricanes by 15 runs    | Batted First |
+| 23 December 2018     | Strikers vs Renegades   | Renegades by 5 wickets   | Chased       |
+| 24 December 2018     | Stars vs Hurricanes     | Hurricanes by 6 wickets  | Chased       |
+| 24th December 2018   | Thunder vs sixers       | Thunder by 21 runs       | Batted First |
+| 26th December 2018   | Strickers vs scorchers  | Scorchers by 7 wickets   | Chased       |
+| 27th December        | Sixers vs Stars         | stars by 5 wickets       | Chased       |
+| 28th December        | thunder vs hurricanes   | hurricanes by 7 wickets  | Chased       |
+| 29th december        | sixers vs renegades     | sixers by 33 runs        | Batted First |
+| 30th decemeber 2018  | scorchers vs hurricanes | hurricanes by 6 wickets  | Chased       |
+| 31st decemember 2018 | strickers vs thunder    | strickers by 20 runs     | Batted First |
+
+: Games in 2018 of the 2018-19 Big Bash Season
+
+i.e. we remove 14 games here of which 6 teams won batting first 16/(22+33-14)
+
+Then from here we have to add on games in the 2019-2020 big bash season that were played in 2019
+
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| Date               | Teams                   | Result                                | Winning Team             |
++====================+=========================+=======================================+==========================+
+| 17 December        | Thunder vs Heat         | Thunder by 29 runs                    | Batted First             |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 18th December      | Scorchers vs Sixers     | Sixers by 8 wickets                   | Chased                   |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 19th december      | Renegades vs thunder    | thunder by 6 wickets                  | chased                   |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 20th decemeber     | hurricanes vs sixers    | hurricanes by 25 runs                 | batted first             |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 20th decemember    | stars vs heat           | stars by 22 runs                      | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 21st decemember    | strickers vs thunder    | no result                             | no result                |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 21st december      | scorchers vs renegades  | scorchers by 11 runs                  | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 22 deceember       | stars vs hurricanes     | stars by 52 runs                      | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 22 decemember      | heat vs sixers          | heat by 48 runs                       | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 23 dec             | strickers vs scorchers  | strickers by 15 (DLS)                 | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 24th dec           | renegades vs hurricanes | hurricanes by 7 wickets               | chased                   |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 26 decemember      | sixers vs scorchers     | sixers by 48 runs                     | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 27 dec             | strickers vs stars      | strikers by 5 runs                    | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 28th december      | thunder vs sixers       | sixers won super over (batting first) | sixers by superover      |
+|                    |                         |                                       |                          |
+|                    |                         |                                       | or sixers batting second |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 29th dec           | strikers vs renegades   | strikers by 18 runs                   | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 30th dec           | hurricanes vs stars     | stars by 4 runs (DLS)                 | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+| 31st december 2019 | thunder vs strikers     | thunder by 3 runs                     | batting first            |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+|                    |                         |                                       |                          |
++--------------------+-------------------------+---------------------------------------+--------------------------+
+
+: Games played in 2019 of the 2019-20 Big Bash
+
 ## Cool cricket examples
@@
 
--   recreate stats from espncricinfo like this [page](https://www.espncricinfo.com/series/19297/statistics/1187665/new-zealand-vs-england-1st-t20i-england-in-new-zealand-2019-20
-)
+-   recreate stats from espncricinfo like this [page](https://www.espncricinfo.com/series/19297/statistics/1187665/new-zealand-vs-england-1st-t20i-england-in-new-zealand-2019-20)
 
++------------------------------+--------------------------------------------------------------------------+
 | Things to do                 | person                                                                   |
-|------------------------------|--------------------------------------------------------------------------|
++==============================+==========================================================================+
 | clean scrapers/csv for waRne | James                                                                    |
++------------------------------+--------------------------------------------------------------------------+
 | data dictionary              | Robert                                                                   |
++------------------------------+--------------------------------------------------------------------------+
 | examples                     | Robert - start (assume same csv being used) - cool cricket example stuff |
++------------------------------+--------------------------------------------------------------------------+
 
 : things to do
+
+## How would you recommend evaluating the selection of players to reduce the required run rate?
+Is there a tool that would could be used to let coaches or selectors do this themselves or is it would it be more appropriate to have an analyst do this prior?
+
+##

### 2020-09-24 paper.Rmd
ROBBED

@@
 James and I entered a reproducibility competition and the winning entry provided a statistic sort of a gateway into the importance of their ''project'' was was specifically in 2019 Australian Big Bash T20 season.
-The statistic given was in the 2019 big bash season teams that batted first won only $43$%$ of games.
+The statistic given was in the 2019 big bash season teams that batted first won only $43$%\$ of games.
 Thankfully we can try to reproduce that statistic now through the use of waRne
@@
 table(blah$year,blah$binary_result)
-````
 ```
+
+\`\`\`
 
@@
 +--------------------+-------------------------+---------------------------------------+--------------------------+
-|                    |                         |                                       |                          |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
 
 : Games played in 2019 of the 2019-20 Big Bash
+
+12 batting first and 3 batting second
+
+(16+12)/(12+3+22+33-14)
+
+So over the past 2 years (not including DLS) we have
+
+(22+39)/(20+39+33+22)
+
+i.e. \>50% of teams win chasing so what does that mean?
 

### 2020-09-26 paper.Rmd
keep on writing

@@
     keep_md: true
+bibliography: references.bib
 
@@
 
-## What is waRne
+Data access is a key enabler for any analytics community.
+Most major sports have easy access to match statistics, for example [nbastatR](https://github.com/abresler/nbastatR) for NBA, [Lahman](https://cran.r-project.org/web/packages/Lahman/Lahman.pdf) for baseball, [deuce](https://github.com/skoval/deuce) for tennis and [nflfastR](https://github.com/mrcaseb/nflfastR) for NFL.
+Through access to sports data and reproducible findings and metrics fans clubs and researchers are better able to understand the game, predict match outcomes and rate players.
+For example the way teams tackled 4th down decisions in the NFL has changed since [@romer2006firms] seminial work.
+As teams have changed their 4th down decision making this allows follow up research such as [@yam2019lost] which looked at more granular data to see if this happens in practice.
 
-We need to think about what it is waRne should be?
+By making data more accessible and more advanced metrics more accessible fans data journalism in sports has grown in recent years.
+For example in [@horowitz2017nflscrapr] has enabled EPA to enter popular discussion among fans.
 
-1.  provide the csv we used for datajam and then people just add games to it
+Cricket is the second largest sport in the world.
+However, unfortunetaly there is no easy accessible way to access ball by ball data nor aggregated statistics of teams and players.
+Data while available on sites like [espncricinfo](https://www.espncricinfo.com/) are not in an easy to use form.
+For example, each match is listed on different webpages so hours upon hours of time would be required to copy and paste a single season.
+Hence, there are significant logistical barriers for prospective fans and analysts studying the game, which stagnats understanding of cricket.
 
-2.  people can not have the csv but just download random tournaments
+This paper describes the waRne package, the first to provide free and easy access to data for cricket for fans.
+Web scraping tools are avaiable for fans to easily scrape the play by play commentary data on espncricinfo.
+For the first time fans can evaluate their favourite teams and players and do so in a reproducible and accessible manner.
 
 ## Why cricket needs reproducibility
+
+Data accessibility enables fans, analyts and researchers to better understand the game. Through being able to reproduce common popular metrics, visualisations and article findings.
+
+Through being able to reproduce, fans are able to make accessible findings for others and importantly they are able to extend and grow concepts.
+Unfortuntely what we see through leading cricket analytics providers is a track record of confusing output for fans.
+This can lead to lower engagement and dismissial of cricket analytics.
 
@@
 James and I entered a reproducibility competition and the winning entry provided a statistic sort of a gateway into the importance of their ''project'' was was specifically in 2019 Australian Big Bash T20 season.
-The statistic given was in the 2019 big bash season teams that batted first won only $43$%\$ of games.
+The statistic given was in the 2019 big bash season teams that batted first won only $43$% of games.
 Thankfully we can try to reproduce that statistic now through the use of waRne
@@
 
+## Easier engagement of fans
+
+To the surprise of many, its hard to engage cricket fans into the analytics behind the game. Reasons for this are generally centred around reproducibility and explainability to fans.
+
+Without an easy accessible medium how can crickets version of an analytics community grow.
+
+something something look towards how EPA has changed the way fans are engaged in NFL analytics - so maybe the change in run rate can be like EPA?
+
+```{r}
+library(tidyverse)
+df<-read.csv("data/combined_2020_09_11.csv")
+
+df%>%
+	group_by(match_id)%>%
+	filter(inning==2)%>%
+	group_by()
+
+```
+
 ## Cool cricket examples
@@
 
-: things to do
+## Dhoni Dilemma
 
-## How would you recommend evaluating the selection of players to reduce the required run rate?
-Is there a tool that would could be used to let coaches or selectors do this themselves or is it would it be more appropriate to have an analyst do this prior?
-
-##
+With this new dataset, we allow fans to put on their analyst thats and to be able to ask themselves, what exactly makes a good closer and what exactly is the Dhoni Dilemma.

### 2020-09-26 paper.Rmd
keep on writing
writing

@@
 
-maybe include stuff around cricviz interesting tweets but fans can recreate or investigate further.
+For example in this [series of tweets](https://twitter.com/benjonescricket/status/1219898578719723520) we see the narrative being pushed that Steve Smith is a good player vs pace bowling unfortunetly just a few months prior the same company and journalist published an article which had Steve Smith doing much worse against pace (balls above 140km/h).
+Unlike a similar sport baseball, fans have no easily accisible way of seeing if Steve Smith vs pace is a strength as alluded to in the original tweet, or a weakness like the same persons published online article.
+In comparsion, fans are able to get a breakdown of Mike Trout vs fastballs from using [baseballr](https://billpetti.github.io/2018-02-19-build-statcast-database-rstats/) which provides access through statcast data from [baseballsavant]([https://baseballsavant.mlb.com)](https://baseballsavant.mlb.com)) .
+
+Unfortunetly this is just one of many examples whereby a relatively simple statistic is provided by media and fans have no mechanism to fact check.
 

### 2020-09-26 paper.Rmd
push
push
writing
writing

@@
 Unfortunetly this is just one of many examples whereby a relatively simple statistic is provided by media and fans have no mechanism to fact check.
+Fact checking is an important avenue for fans to not only engage and understand statistics, but having this mechanism also stops analytics companies from putting out misleading conclusions and findings.
+
+However though using waRne we are able to not only compare Steve Smiths performance vs bowling type but we are able to easily compare Steve Smiths performance vs bowling types with other players.
+
+### Steve Smith vs Bowling Type Chart
+
+James here need to figure out the code, do we keep it seperate so in the example we still have to join on bowling type to the ball by ball, or will it just be an extra column joined already.
+
+```{r}
+library(tidyverse)
+
+# steve smith vs pace graph
+
+```
+
+Another reason why is that cricket is a weird sport with a lot of old school rules of thumb
 

### 2020-09-26 paper.Rmd
writing
writing

@@
 
-Another reason why is that cricket is a weird sport with a lot of old school rules of thumb
+Another reason why is that cricket is a weird sport with a lot of old school rules of thumb that are followed/talked about.
+
+For example it has become popular conjecture that its better that teams should bat second (chase down a total set by the oppoition). Through waRne we can answer such question for the T20 game.
+
+For example we can explore do teams win more often when chasing in the big bash over the history of the big bash, or we can plot a line chart to see if there is a movement towards batting last and winning.
+
+```{r}
+
+# insert above plots here
+```
+
+We can also use waRne and other R packages as an easy way to engage students in learning statistical concepts.
+
+For example using the above we can not only look at the answers graphically, but we can run a statistical test.
+The example we would use is instead of testing is a coin biased (different from 50/50) we can test to see if the winning \\% is different from $0.5$ for teams chasing.
+
+```{r}
+
+```
 

### 2020-09-26 paper.Rmd
writing
writing

@@
 ```
+
+## Using waRne to teach undergraduate statistics
 
@@
 
-```{r}
-library(tidyverse)
-df<-read.csv("data/combined_2020_09_11.csv")
-
-df%>%
-	group_by(match_id)%>%
-	filter(inning==2)%>%
-	group_by()
-
-```
-
 ## Cool cricket examples
@@
 With this new dataset, we allow fans to put on their analyst thats and to be able to ask themselves, what exactly makes a good closer and what exactly is the Dhoni Dilemma.
+
+To understand the Dhoni Deilemma, we need to see some interesting things to do with the change in RRR.
+
+Cricket is an obvious sort of sport, to win in the end you have to get one run more than your opposition.
+A commonly used statistic during the broadcast is the required run rate, for example if you need 120 to win off 20 overs your RRR is \$6\$ runs an over
+
+What follows is then obvious is the chasing team wins if they are able to lower the RRR that is score quicker.
+
+Espncricinfo did article on teams that win games get a higher proportion of their runs in the powerplay and first 10 overs.
+But proportion of team runs in T20 games doesn't really make sense as a batting metric as we are not just concerned with how many runs they get, but how quickly they get them.
+So we introduce a new way to measure batsman change in RRR.
+
+```{r}
+# graph of powerplay
+```
+
+This graph makes sense as we can see the best openers in the world lower the required run rate so whats the Dhoni Dilemma?
+
+```{r}
+# middle overs (enter in the middle)
+```
+
+Dhoni is renowed for being one of the worlds best closers but we can see it seems as though in the middle he is letting the game get out of hand. How can one of the worlds best closers

### 2020-09-26 paper_rob_working .Rmd
why
just in case stuff up the whole master thing or if I end up conflicting stuff while your on flight etc

--- 2020-09-26 paper.Rmd
@@
 Dhoni is renowed for being one of the worlds best closers but we can see it seems as though in the middle he is letting the game get out of hand. How can one of the worlds best closers
+
+## notes for James
+
+So I don't think its like a complex paper, because of the way cricket is at the moment I think the biggest influence this stuff can make is
+
+-   being able to fact check
+
+-   having batsman metrics that are beyond average, total runs and SR
+
+-   What we can maybe mention?
+is that change in RRR while simple seems to be quite effective, a better measurement would be the DLS which we know can be derived from the ball by ball data or alternatively if someone has a resource table they can instead of change in RRR use change in resources or whatever
+
+For the examples, I think its pretty easy to get the graphs to look nice, its more about whats the dataset look like?
+i.e.
+do we need the first few lines for reproducibility to be joins or is it sort of like done already

### 2020-09-28 code/run_scraper.R
push
push

@@ -5,7 +5,7 @@ library(rvest)
 library(htmltools)
 library(lubridate)
 library(magrittr)
-source("/Users/James/James/Reproducibility-in-Cricket/code/scraping_functions.R")
+source("/Users/robertnguyen/Library/Mobile Documents/com~apple~CloudDocs/OneDrive/Reproducibility-in-Cricket/code/scraping_functions.R")
 
 # You will need to create an image to do this in docker
 # system('docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome')
@@ -17,7 +17,7 @@ system('docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome')
 # remove it for you. The links are printed for each 
 complete = 0
 while(complete!=1){
-	commentary <- try(grab_tournament_links(2018, "/Users/James/James/Reproducibility-in-Cricket"))
+	commentary <- try(grab_tournament_links(2018, "/Users/robertnguyen/Library/Mobile Documents/com~apple~CloudDocs/OneDrive/Reproducibility-in-Cricket"))
 	if(class(commentary)=="try-error"){
 		error_type <- attr(commentary, "condition")
 		print(class(error_type))
@@ -43,7 +43,7 @@ while(complete!=1){
 		print("503 error, taking a 2-4 minute break before trying again")
 		Sys.sleep(sample(120:240, 1))
 		complete <- 0
-	}else if(error_type$message == "object 'starting_date' not found"){
+	}else if(error_type$message == "x object 'starting_date' not found"){
 		print("I think it failed to get the results page loaded, 
           taking a 2-4 minute break before trying again")
 		Sys.sleep(sample(120:240, 1))
@@ -53,6 +53,10 @@ while(complete!=1){
           taking a 2-4 minute break before trying again")
 		Sys.sleep(sample(120:240, 1))
 		complete <- 0
+	}else if(error_type$message == "Problem with `mutate()` input `starting_date`.\n\033[31mx\033[39m object 'starting_date' not found\n\033[34mℹ\033[39m Input `starting_date` is `as.Date(starting_date)`."){
+		print("x object 'starting_date' not found")
+		Sys.sleep(sample(120:240, 1))
+		complete <- 0		
 	}else{
 		complete <- 1
 	}

### 2020-09-29 paper_rob_working .Rmd
push
push

@@
 
-The importance of reproducibility, and the related issue of open access to data, has received a lot of recent attention.
-Momentum on these issues is gathering in the sports analytics community.
+The importance of reproducibility, and the related issue of open access to data, has received a lot of recent attention.\^[why this not work] Momentum on these issues is gathering in the sports analytics community.
 While cricket is the second largest commercial sport in the world, unlike other popular international sports, there has been no mechanism for the public to access comprehensive statistics on players and teams.
@@
 Data while available on sites like [espncricinfo](https://www.espncricinfo.com/) are not in an easy to use form.
-For example, each match is listed on different webpages so hours upon hours of time would be required to copy and paste a single season.
+For example, each match is listed on different webpages so hours upon hours of time would be required to copy and paste a single season, not to mention the added difficulty of linking players between games and competitions in different countries.
 Hence, there are significant logistical barriers for prospective fans and analysts studying the game, which stagnats understanding of cricket.
@@
 For the first time fans can evaluate their favourite teams and players and do so in a reproducible and accessible manner.
+We hope that this package can be used for fans to better understand the game, for teachers to use for interesting examples in class and for analysts in clubland who might not have access to ball by ball data.
 
@@
 Unlike a similar sport baseball, fans have no easily accisible way of seeing if Steve Smith vs pace is a strength as alluded to in the original tweet, or a weakness like the same persons published online article.
-In comparsion, fans are able to get a breakdown of Mike Trout vs fastballs from using [baseballr](https://billpetti.github.io/2018-02-19-build-statcast-database-rstats/) which provides access through statcast data from [baseballsavant]([https://baseballsavant.mlb.com)](https://baseballsavant.mlb.com)) .
+In comparsion, fans are able to get a breakdown of Mike Trout vs fastballs from using [baseballr](https://billpetti.github.io/2018-02-19-build-statcast-database-rstats/) which provides access through statcast data from [baseballsavant](https://baseballsavant.mlb.com/).
 
 Unfortunetly this is just one of many examples whereby a relatively simple statistic is provided by media and fans have no mechanism to fact check.
-Fact checking is an important avenue for fans to not only engage and understand statistics, but having this mechanism also stops analytics companies from putting out misleading conclusions and findings.
+Fact checking is an important avenue for fans to not only engage and understand statistics, but having this mechanism also stops analytics people/companies from putting out misleading conclusions and findings.
 
@@
 
-James here need to figure out the code, do we keep it seperate so in the example we still have to join on bowling type to the ball by ball, or will it just be an extra column joined already.
+Fans of cricket and Australian cricket especially might be interested to see if this is a statistical qwerk of small samples or if Steve Smith generally does perform as an elite cricketer vs pace bowling.
+Teachers of statistics classes might use this as an interesting example to introduce [tidyverse](https://www.tidyverse.org/learn/) principals [@wickham2019welcome].
 
@@
 ```
-
-Another reason why is that cricket is a weird sport with a lot of old school rules of thumb that are followed/talked about.
-
-For example it has become popular conjecture that its better that teams should bat second (chase down a total set by the oppoition). Through waRne we can answer such question for the T20 game.
-
-For example we can explore do teams win more often when chasing in the big bash over the history of the big bash, or we can plot a line chart to see if there is a movement towards batting last and winning.
 
@@
 For example using the above we can not only look at the answers graphically, but we can run a statistical test.
-The example we would use is instead of testing is a coin biased (different from 50/50) we can test to see if the winning \\% is different from $0.5$ for teams chasing.
+A common classroom example for learning the binomial distribution is to ask if a coin is based given a proportion of heads and tails given a sample size [@mcelreath2020statistical].
+Instead of asking about coins, fans of sport and cricket might be interested to ask the question; after winning the coin toss, should a team decide to bat first or bat second.
 
-```{r}
+While a coin toss is a seemingly trivial example, like in most professional sports the winner of the coin toss gets to decide what to do.
+In cricket, the winner of the coin toss gets to decide if they want to bat first and thus set the total that the opposition needs to pass by a single run to win, or if they want to bowl first and thus chase down the total set.
+Deciding what to do after winning a coin toss has proved to be a popular media piece in recent years \\footnote{[https://www.espncricinfo.com/video/clip/\_/id/23230682](https://www.espncricinfo.com/video/clip/_/id/23230682])} \^[[https://www.espncricinfo.com/story/\_/id/21489056/stuart-wark-cricket-move-away-coin-toss?]](https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss?%5D) \^[[https://www.espncricinfo.com/story/\_/id/20429499/lehmann-backs-scrapping-toss]](https://www.espncricinfo.com/story/_/id/20429499/lehmann-backs-scrapping-toss%5D) \^[[https://www.espncricinfo.com/story/\_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams]](https://www.espncricinfo.com/story/_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams%5D) \^[<https://www.forbes.com/sites/tristanlavalette/2018/08/20/are-cricket-matches-being-decided-by-the-luck-of-a-coin-toss/#735456837eff>]
 
-```
+\^[<https://www.statsinsider.com.au/bbl/how-important-is-winning-the-toss-in-the-big-bash-league>]
 
-James and I entered a reproducibility competition and the winning entry provided a statistic sort of a gateway into the importance of their ''project'' was was specifically in 2019 Australian Big Bash T20 season.
-The statistic given was in the 2019 big bash season teams that batted first won only $43$% of games.
-Thankfully we can try to reproduce that statistic now through the use of waRne
+\^[<https://www.espncricinfo.com/story/_/id/18568387/tim-wigmore-how-batting-second-become-more-fruitful-more-popular>]
 
-Wikipedia tells us that there were the following:
+Fans and analysts of the game might want to make this decision based on a simple statistic, for example they might make it based on answer these questions instead.
 
-In the 2018-19 Big Bash League there were 59 matches played. Of which we had 3 DLS results.
+-   Do teams that bat second have a higher winning percentage than those who bat first?
 
-| Games                             | Result                        |
-|-----------------------------------|-------------------------------|
-| 21 DEC 2018 - THUNDER VS STARS    | THUNDER WON BY 15 RUNS (D/L)  |
-| 2 FEB 2019 - THUNDER VS SIXERS    | SIXERS WON BY 9 WICKETS (D/L) |
-| 8 JAN 2019 HEAT VS THUNDER        | HEAT BY 15 RUNS (D/L)         |
-| 17th January 2019 Thunder vs Heat | no result                     |
-
-: Games under consideration in 2018-19 Big Bash
-
-```{r, investigation}
-df<-read.csv("data/combined_2020_09_11.csv")
-
-blah<-df%>%
-	dplyr::select(duckworth_lewis,year,match_id, tournament_name, tournament_type, team, inning, binary_result)%>%
-	dplyr::filter(tournament_type%in% c("Twenty20 Internationals", "Twenty20"  ))%>%
-	dplyr::filter(tournament_name%in% c("Twenty20 Big Bash","Big Bash League" ))%>%
-									dplyr::filter(inning==1)%>%
-									dplyr::group_by()%>%
-									distinct()
-# dataset doesn't contain duckworth lewis results. 
-table(blah$year,blah$binary_result)
-```
-
-\`\`\`
-
-So here we can see that we in the big bash season 2018-19 we have a current winning percentage batting first of 22/(22+33) =0.4 but this doesn't include the duckworth lewis games or the washed out completely game.
-
-If we were to add in the duckworth lewis games we get 24/(24+34) which is equal to 0.414. Which isn't the result.
-
-We can also clearly see that in the 20192020 big bash season the team batting first wins 39/(39+20) or 0.66 so what could be going on? Did Quantium mean only games played in the actual year 2019?
-
-In that case from the 2018-2019 season we would remove the following games
-
-| Date                 | Game                    | Result                   | Winning Team |
-|----------------------|-------------------------|--------------------------|--------------|
-| 19 December 2018     | Heat Vs Strickers       | Strickers 5 Wickets      | Chased       |
-| 20th December 2018   | Scorchers Vs Renegades  | Renegades by 4 wickets   | Chased       |
-| 21 December 2018     | Thunder vs Stars        | Thunder by 15 runs (D/L) | Batted First |
-| 22 December 2018     | Sixers vs Scorchers     | Sixers by 17 runs        | Batted First |
-| 22 December          | Hurricanes vs Heat      | Hurricanes by 15 runs    | Batted First |
-| 23 December 2018     | Strikers vs Renegades   | Renegades by 5 wickets   | Chased       |
-| 24 December 2018     | Stars vs Hurricanes     | Hurricanes by 6 wickets  | Chased       |
-| 24th December 2018   | Thunder vs sixers       | Thunder by 21 runs       | Batted First |
-| 26th December 2018   | Strickers vs scorchers  | Scorchers by 7 wickets   | Chased       |
-| 27th December        | Sixers vs Stars         | stars by 5 wickets       | Chased       |
-| 28th December        | thunder vs hurricanes   | hurricanes by 7 wickets  | Chased       |
-| 29th december        | sixers vs renegades     | sixers by 33 runs        | Batted First |
-| 30th decemeber 2018  | scorchers vs hurricanes | hurricanes by 6 wickets  | Chased       |
-| 31st decemember 2018 | strickers vs thunder    | strickers by 20 runs     | Batted First |
-
-: Games in 2018 of the 2018-19 Big Bash Season
-
-i.e. we remove 14 games here of which 6 teams won batting first 16/(22+33-14)
-
-Then from here we have to add on games in the 2019-2020 big bash season that were played in 2019
-
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| Date               | Teams                   | Result                                | Winning Team             |
-+====================+=========================+=======================================+==========================+
-| 17 December        | Thunder vs Heat         | Thunder by 29 runs                    | Batted First             |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 18th December      | Scorchers vs Sixers     | Sixers by 8 wickets                   | Chased                   |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 19th december      | Renegades vs thunder    | thunder by 6 wickets                  | chased                   |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 20th decemeber     | hurricanes vs sixers    | hurricanes by 25 runs                 | batted first             |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 20th decemember    | stars vs heat           | stars by 22 runs                      | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 21st decemember    | strickers vs thunder    | no result                             | no result                |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 21st december      | scorchers vs renegades  | scorchers by 11 runs                  | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 22 deceember       | stars vs hurricanes     | stars by 52 runs                      | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 22 decemember      | heat vs sixers          | heat by 48 runs                       | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 23 dec             | strickers vs scorchers  | strickers by 15 (DLS)                 | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 24th dec           | renegades vs hurricanes | hurricanes by 7 wickets               | chased                   |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 26 decemember      | sixers vs scorchers     | sixers by 48 runs                     | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 27 dec             | strickers vs stars      | strikers by 5 runs                    | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 28th december      | thunder vs sixers       | sixers won super over (batting first) | sixers by superover      |
-|                    |                         |                                       |                          |
-|                    |                         |                                       | or sixers batting second |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 29th dec           | strikers vs renegades   | strikers by 18 runs                   | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 30th dec           | hurricanes vs stars     | stars by 4 runs (DLS)                 | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 31st december 2019 | thunder vs strikers     | thunder by 3 runs                     | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-
-: Games played in 2019 of the 2019-20 Big Bash
-
-12 batting first and 3 batting second
-
-(16+12)/(12+3+22+33-14)
-
-So over the past 2 years (not including DLS) we have
-
-(22+39)/(20+39+33+22)
-
-i.e. \>50% of teams win chasing so what does that mean?
+-   Is this consistent across leagues and levels of cricket?
 

### 2020-09-30 code/html_reader.R
adding the latest work to the repo including html reader functions

@@ -1,0 +1,22 @@
+library(tidyverse)
+library(RSelenium)
+library(XML)
+library(rvest)
+library(htmltools)
+library(lubridate)
+library(magrittr)
+source("/Users/James/James/Reproducibility-in-Cricket/code/html_reader_functions.R")
+
+# Load in the old data, the DLS table, and the player stat table
+load("/Users/James/James/Reproducibility-in-Cricket/data_files/odi_dls.Rdata")
+load("/Users/James/James/Reproducibility-in-Cricket/data_files/player_list.Rdata")
+load("/Users/James/James/Reproducibility-in-Cricket/data_files/cmsac_data.Rdata")
+
+# We read in by year, and later add it to the dataset we already have. This is for html that you have 
+# scraped on your machine only, it does nothing otherise
+data2020 <- extract_data_commentary_scorecards(2020, "/Users/James/James/Reproducibility-in-Cricket")
+
+# Fairly easy to add in your newly acquired data after you scrape it, just replace the year.
+cmsac_data %>% 
+	dplyr::filter(year!=2020) %>% 
+	bind_rows(data2020)

### 2020-09-30 code/html_reader_functions.R

@@ -1,0 +1,806 @@
+# Grabbing the data from the html is fairly straightforward, and by saving the html in the original 
+# scraper we are able to go back and make changes to the scraper over time 
+
+extract_data_commentary_scorecards <- function(year_input, root_path){
+	
+	data_path <- paste(root_path, "/", "data", sep = "")
+	player_path <- paste(data_path, "_files/player_list.Rdata", sep = "")
+	
+	full_commentary_csv <- 
+		read_csv(paste(data_path, "/", year_input, "/master_commentary.csv", sep = "")) %>% 
+		mutate(match_id = 0, 
+					 rescraping_required = "NA", 
+					 row_n = row_number())
+	write_csv(full_commentary_csv, paste(data_path, "/", year_input, "/master_commentary.csv", sep = ""))
+	
+	commentary <- 
+		full_commentary_csv %>%
+		dplyr::filter(tournament_type %in% c("One-Day Internationals", 
+																				 "Twenty20 Internationals", "Twenty20")) %>%
+		dplyr::select(tournament_type, tournament_name, inning_1_commentary, inning_2_commentary, 
+									scorecard_path, scorecard_link, commentary_link) 
+	
+	out <- tibble()
+	
+	for(i in 1:nrow(commentary)){
+		
+		scorecard_link_check <- 
+			commentary$scorecard_link[i]
+		commentary_link_check <- 
+			commentary$commentary_link[i]
+		tournament_name_check <- 
+			commentary$tournament_name[i]
+		tournament_type_check <- 
+			commentary$tournament_type[i]
+		
+		match_id_num <- as.numeric(word(word(scorecard_link_check, 2, sep = "scorecard/"), 1, sep = "\\/"))
+		
+		which_row_to_add_info <- 
+			full_commentary_csv %>% 
+			dplyr::filter(scorecard_link == scorecard_link_check, 
+										commentary_link == commentary_link_check, 
+										tournament_type %in% c("One-Day Internationals", 
+																					 "Twenty20 Internationals", "Twenty20"), 
+										# commentary_available == TRUE, 
+										tournament_name == tournament_name_check, 
+										tournament_type == tournament_type_check) %>% 
+			pull(row_n)
+		
+		print(i)
+		print(commentary$scorecard_link[i])
+		
+		score_path <- 
+			commentary$scorecard_path[i]
[…]

### 2020-09-30 code/run_scraper.R

@@ -5,7 +5,7 @@ library(rvest)
 library(htmltools)
 library(lubridate)
 library(magrittr)
-source("/Users/robertnguyen/Library/Mobile Documents/com~apple~CloudDocs/OneDrive/Reproducibility-in-Cricket/code/scraping_functions.R")
+source("/Users/James/James/Reproducibility-in-Cricket/code/scraping_functions.R")
 
 # You will need to create an image to do this in docker
 # system('docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome')
@@ -17,7 +17,7 @@ system('docker run -d -p 4445:4444 --shm-size 2g selenium/standalone-chrome')
 # remove it for you. The links are printed for each 
 complete = 0
 while(complete!=1){
-	commentary <- try(grab_tournament_links(2018, "/Users/robertnguyen/Library/Mobile Documents/com~apple~CloudDocs/OneDrive/Reproducibility-in-Cricket"))
+	commentary <- try(grab_tournament_links(20192020, "/Users/James/James/Reproducibility-in-Cricket"))
 	if(class(commentary)=="try-error"){
 		error_type <- attr(commentary, "condition")
 		print(class(error_type))

### 2020-09-30 code/scraping_functions.R

@@ -1088,7 +1088,7 @@ remove_master_row <- function(year_input, root_path, row_to_drop){
 
 # So just use the same root as before, to the folder that will contain all the code, and data, and
 # whatever else you may eventially put in here in your cricket analysis
-collect_commentary_html_new_new_format <- function(year_input, root_path){
+collect_commentary_html <- function(year_input, root_path){
 	# system('docker kill $(docker ps -q)')
 	# Set up the appropriate paths first
 	

### 2020-10-01 paper_rob_working .Rmd
push
push

@@
 In cricket, the winner of the coin toss gets to decide if they want to bat first and thus set the total that the opposition needs to pass by a single run to win, or if they want to bowl first and thus chase down the total set.
-Deciding what to do after winning a coin toss has proved to be a popular media piece in recent years \\footnote{[https://www.espncricinfo.com/video/clip/\_/id/23230682](https://www.espncricinfo.com/video/clip/_/id/23230682])} \^[[https://www.espncricinfo.com/story/\_/id/21489056/stuart-wark-cricket-move-away-coin-toss?]](https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss?%5D) \^[[https://www.espncricinfo.com/story/\_/id/20429499/lehmann-backs-scrapping-toss]](https://www.espncricinfo.com/story/_/id/20429499/lehmann-backs-scrapping-toss%5D) \^[[https://www.espncricinfo.com/story/\_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams]](https://www.espncricinfo.com/story/_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams%5D) \^[<https://www.forbes.com/sites/tristanlavalette/2018/08/20/are-cricket-matches-being-decided-by-the-luck-of-a-coin-toss/#735456837eff>]
+Deciding what to do after winning a coin toss has proved to be a popular media piece in recent years \footnote{https://www.espncricinfo.com/video/clip/\_/id/23230682}
+
+(https://www.espncricinfo.com/video/clip/_/id/23230682%5D)} \^[[https://www.espncricinfo.com/story/\_/id/21489056/stuart-wark-cricket-move-away-coin-toss?]](https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss?%5D) \^[[https://www.espncricinfo.com/story/\_/id/20429499/lehmann-backs-scrapping-toss]](https://www.espncricinfo.com/story/_/id/20429499/lehmann-backs-scrapping-toss%5D) \^[[https://www.espncricinfo.com/story/\_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams]](https://www.espncricinfo.com/story/_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams%5D) \^[<https://www.forbes.com/sites/tristanlavalette/2018/08/20/are-cricket-matches-being-decided-by-the-luck-of-a-coin-toss/#735456837eff>]
 
@@
 
+```{r}
+```
+
+Instead of just looking at this graphically, fans of crickets and teachers of undergraduate statistics might use the dataset as a ''biased coin'' example.
+So instead of asking, given a proportion of heads over \$n\$ tosses instructers could ask is a team better off batting first or second.
+
+```{r}
+```
+
 ## Easier engagement of fans
@@
 something something look towards how EPA has changed the way fans are engaged in NFL analytics - so maybe the change in run rate can be like EPA?
+
+<https://twitter.com/cricvizanalyst/status/1311260989267087361>
 
@@
 -   recreate stats from espncricinfo like this [page](https://www.espncricinfo.com/series/19297/statistics/1187665/new-zealand-vs-england-1st-t20i-england-in-new-zealand-2019-20)
-
-+------------------------------+--------------------------------------------------------------------------+
-| Things to do                 | person                                                                   |
-+==============================+==========================================================================+
-| clean scrapers/csv for waRne | James                                                                    |
-+------------------------------+--------------------------------------------------------------------------+
-| data dictionary              | Robert                                                                   |
-+------------------------------+--------------------------------------------------------------------------+
-| examples                     | Robert - start (assume same csv being used) - cool cricket example stuff |
-+------------------------------+--------------------------------------------------------------------------+
 
@@
 
-Dhoni is renowed for being one of the worlds best closers but we can see it seems as though in the middle he is letting the game get out of hand. How can one of the worlds best closers
+Dhoni is renowed for being one of the worlds best closers but we can see it seems as though in the middle he is letting the game get out of hand.
+How can one of the worlds best closers consistently put his teams in worse positions?
 

### 2020-10-01 James_paper_copy.Rmd
adding my copy of the paper with the graphs

--- 2020-09-26 paper.Rmd
@@
 title: "waRne - Putting cricket analytics in a spin"
+bibliography: references.bib
+output:
+  pdf_document:
+    toc: yes
+  html_document:
+    keep_md: yes
+    toc: yes
+    toc_float: yes
 authors: Robert(1), James (2)
-output: 
-  html_document:
-    toc: true
-    toc_float: true
-    keep_md: true
-bibliography: references.bib
-
 ---
@@
 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+library(stringr)
+library(curl)
+library(tidyverse)
+library(magrittr)
+library(lubridate)
+library(ggplot2)
+library(ggalt)
+library(gganimate)
+library(gifski)
+player_database <- 
+  read_csv("/Users/James/James/Cricket_Thesis/Data/new_player_database_crickjam.csv") %>% 
+  dplyr::select(player_id, player_name, batting_style, bowling_style) %>% 
+  distinct() %>% 
+  dplyr::filter(player_name != "Rohit Gurunath Sharma") %>% 
+  bind_rows(read_csv("/Users/James/James/Cricket_Thesis/Data/new_player_database_crickjam.csv") %>% 
+              dplyr::filter(player_id == 34102) %>% 
+              dplyr::select(player_id, player_name, batting_style, bowling_style) %>% 
+              drop_na() %>% 
+              distinct()) %>% 
+  dplyr::filter(player_name !="Sultan Ahmed")
+
+bowler_vars <- 
+  player_database %>% 
+  dplyr::select(-c(batting_style, player_name)) %>% 
+  drop_na()%>% 
+  mutate(bowling_style = recode(bowling_style, "Right-arm fast" = "Pace",
+                                "Slow left-arm orthodox"= "Spin",
+                                "Legbreak"= "Spin",
+                                "Right-arm medium"= "Pace",
+                                "Left-arm medium-fast"= "Pace",
+                                "Right-arm offbreak"= "Spin",
+                                "Legbreak googly"= "Spin",
+                                "Left-arm fast"= "Pace",
+                                "Right-arm medium-fast"= "Pace",
+                                "Left-arm fast-medium"= "Pace",
+                                "Right-arm fast-medium"= "Pace",
+                                "Left-arm medium"= "Pace",
+                                "Slow left-arm chinaman"= "Spin",
+                                "Right-arm offbreak, Legbreak"= "Spin",
+                                "Right-arm medium, Right-arm offbreak"= "Spin",
+                                "Left-arm medium, Slow left-arm orthodox, 
+                                               Slow left-arm chinaman"= "Spin", 
+                                "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" = "Spin"))
+
+batting_vars <- 
+  player_database %>% 
+  dplyr::select(-c(bowling_style, player_name)) %>%
+  drop_na()
+
+cricjam_data <- 
+  read_csv("/Users/James/James/Reproducibility-in-Cricket/Data/combined_2020_09_11.csv") %>% 
+  dplyr::filter(!grepl("No result", match_summary)) %>%
+  mutate(bucketed_year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
+  left_join(batting_vars, by = c("batsman_id" = "player_id")) %>% 
+  left_join(bowler_vars, by = c("bowler_id" = "player_id")) %>% 
+  mutate(batsman_name = str_remove(batsman_name, "\\(\\)"), 
+         batsman_name = str_trim(batsman_name), 
+         runs = extra_runs+earned_runs)
+# 
+# cricjam_data_partner <- 
+#   cricjam_data %>% 
+#   dplyr::select(match_id, inning, batsman_name, batsman_id, batting_order) %>% 
+#   distinct() %>% 
+#   arrange(batting_order) %>% View()
+
+wicket_types <-
+  cricjam_data %>% 
+  dplyr::filter(wicket==TRUE) %>% 
+  dplyr::select(wicket_description, play_description, bowler, batsman, bowler_name, batsman_name, 
+                match_id, inning, over) %>%
+  mutate(caught_out = if_else(str_detect(wicket_description, " c "), TRUE, FALSE), 
+         bowled_out = if_else(str_detect(wicket_description, " b "), TRUE, FALSE), 
+         run_out = if_else(str_detect(wicket_description, " run out "), TRUE, FALSE), 
+         stumped_out = if_else(str_detect(wicket_description, " st "), TRUE, FALSE), 
+         leg_before_wicket = if_else(str_detect(wicket_description, " lbw "), TRUE, FALSE), 
+         retired_hurt = if_else(str_detect(wicket_description, " retired hurt "), TRUE, FALSE), 
+         obstructing_the_field = if_else(str_detect(wicket_description, " obstructing the field "), TRUE, FALSE), 
+         retired_out = if_else(str_detect(wicket_description, " retired out "), TRUE, FALSE), 
+         null = if_else(str_detect(wicket_description, " null "), TRUE, FALSE)) %>% 
+  # all_conditions = if_else(caught_out == TRUE |
+  #                          bowled_out == TRUE | caught_out == TRUE |
+  #                            run_out == TRUE |retired_out == TRUE |
+  #                            retired_hurt == TRUE |leg_before_wicket == TRUE |
+  #                            obstructing_the_field == TRUE |null == TRUE |
+  #                            stumped_out == TRUE, TRUE, FALSE))
+  mutate(out_condition = case_when(caught_out==TRUE |
+                                     caught_out==TRUE & bowled_out == TRUE ~ "Caught", 
+                                   run_out == TRUE |
+                                     run_out==TRUE & bowled_out == TRUE ~ "Run Out", 
+                                   stumped_out == TRUE |
+                                     stumped_out==TRUE & bowled_out == TRUE ~ "Stumped", 
+                                   leg_before_wicket == TRUE |
+                                     leg_before_wicket==TRUE & bowled_out == TRUE ~ "Leg Before Wicket", 
+                                   retired_hurt == TRUE |
+                                     retired_hurt==TRUE & bowled_out == TRUE ~ "Retired Hurt", 
+                                   obstructing_the_field == TRUE |
+                                     obstructing_the_field==TRUE & bowled_out == TRUE ~ "Obstructing the Field", 
+                                   retired_out == TRUE |
+                                     retired_out==TRUE & bowled_out == TRUE ~ "Retired Out", 
+                                   null == TRUE |
+                                     null==TRUE & bowled_out == TRUE ~ "Caught", 
+                                   bowled_out == TRUE & caught_out == FALSE |
+                                     bowled_out == TRUE & stumped_out == FALSE |
+                                     bowled_out == TRUE & run_out == FALSE |
+                                     bowled_out == TRUE & leg_before_wicket == FALSE |
+                                     bowled_out == TRUE & retired_hurt == FALSE |
+                                     bowled_out == TRUE & obstructing_the_field == FALSE |
+                                     bowled_out == TRUE & retired_out == FALSE |
+                                     bowled_out == TRUE & null == FALSE ~"Bowled"), 
+         out_batsman = case_when(out_condition == "Bowled" ~ word(wicket_description, 1, sep = " b "), 
+                                 out_condition == "Caught" & null == FALSE ~ word(wicket_description, 1, sep = " c "), 
+                                 out_condition == "Caught" & null == TRUE ~ word(wicket_description, 1, sep = " null "), 
+                                 out_condition == "Run Out" ~ word(wicket_description, 1, sep = " run out "), 
+                                 out_condition == "Stumped" ~ word(wicket_description, 1, sep = " st "), 
+                                 out_condition == "Leg Before Wicket" ~ word(wicket_description, 1, sep = " lbw "), 
+                                 out_condition == "Retired Hurt" ~ word(wicket_description, 1, sep = " retired hurt "), 
+                                 out_condition == "Obstructing the Field" ~ word(wicket_description, 1, sep = " obstructing the field "), 
+                                 out_condition == "Retired Out" ~ word(wicket_description, 1, sep = " retired out ")),  
+         out_batsman = str_trim(out_batsman)) %>% 
+  dplyr::select(-c(caught_out, bowled_out, stumped_out, leg_before_wicket, retired_hurt, 
+                   obstructing_the_field, retired_hurt, null, run_out, retired_out)) %>% 
+  mutate(batsman_or_partner_out = if_else(out_batsman == batsman_name, "Batsman","Partner"))
+
+cricjam_data <- 
+  cricjam_data %>% 
+  left_join(wicket_types, by = c("wicket_description" = "wicket_description", 
+                                 "play_description" = "play_description", 
+                                 "bowler" = "bowler", 
+                                 "batsman" = "batsman", 
+                                 "batsman_name" = "batsman_name", 
+                                 "match_id" = "match_id", 
+                                 "inning" = "inning", 
+                                 "over" = "over", 
+                                 "bowler_name" = "bowler_name")) %>% 
+  rename(inning_summary = inning_total)
+
+only_big_t20_matches <- 
+  cricjam_data %>% 
+  dplyr::filter(tournament_type == "Twenty20", 
+                tournament_name %in% c("Twenty20 Big Bash", "Indian Premier League", "Big Bash League", 
+                                       "Champions League Twenty20", "Pepsi Indian Premier League")) %>% 
+  dplyr::select(match_id) %>% 
+  distinct() %>% 
+  pull(match_id)
+
+cricjam_data_big <- 
+  cricjam_data %>% 
+  dplyr::filter(tournament_type == "Twenty20 Internationals") %>% 
+  bind_rows(cricjam_data %>% 
+              dplyr::filter(match_id %in% only_big_t20_matches))
+```
+
+\color{red}{\text{Yeah I forgot I wouldn't have any internet on the plane, so everything I wrote is not referencing anything from a link or package I didn't have downloaded prior to boarding}}
+
 ## Abstract
@@
 The importance of reproducibility, and the related issue of open access to data, has received a lot of recent attention.
-Momentum on these issues is gathering in the sports analytics community.
-While cricket is the second largest commercial sport in the world, unlike other popular international sports, there has been no mechanism for the public to access comprehensive statistics on players and teams.
+Momentum on these issues is gathering in the sports analytics community, particularly in North America where open access to quality sports data has become the norm.
+Cricket is the second largest commercial sport in the world, but unlike other popular international sports, there has been no mechanism for the public to access comprehensive statistics on players and teams.
 Expert commentary currently relies heavily on data that isn't made readily accessible and this produces an unnecessary barrier for the development of an inclusive sports analytics community.
@@
 Hence, there are significant logistical barriers for prospective fans and analysts studying the game, which stagnats understanding of cricket.
+The Duckworth-Lewis-Stern table, used to reset targets for matches that experience rain delays in all of cricket, is maintained by one man in Australia.
+Not even teams have access to the table.
+\color{red}{text{what do they do, call it into the ICC to get a temporary copy or something when there's a delay?}}
 
@@
 For the first time fans can evaluate their favourite teams and players and do so in a reproducible and accessible manner.
+We have included the data for ODI and T20 from the 2010 archived season on ESPCcricinfo up to the matches on September 8, 2020.
+If a person wishes to expand this data to include more recent games, this package can scrape and update the dataset.
 
@@
 
-Data accessibility enables fans, analyts and researchers to better understand the game. Through being able to reproduce common popular metrics, visualisations and article findings.
+Data accessibility enables fans, analysts and researchers to better understand the game. Through being able to reproduce common popular metrics, visualisations and article findings.
 
@@
 
+\color{red}{\text{I'm not sure if we should slam the datajam competition directly.
+I think the guy tweeting out conflicting news is a good point and I want to keep that in because he was a dick about his results, but we just lost to Quantium and I worry it will make us seem petty or look like sore losers.
+We could just indirectly show that they were wrong, but not mention them at all, I would feel fine with that}}
+
 For example in this [series of tweets](https://twitter.com/benjonescricket/status/1219898578719723520) we see the narrative being pushed that Steve Smith is a good player vs pace bowling unfortunetly just a few months prior the same company and journalist published an article which had Steve Smith doing much worse against pace (balls above 140km/h).
 Unlike a similar sport baseball, fans have no easily accisible way of seeing if Steve Smith vs pace is a strength as alluded to in the original tweet, or a weakness like the same persons published online article.
-In comparsion, fans are able to get a breakdown of Mike Trout vs fastballs from using [baseballr](https://billpetti.github.io/2018-02-19-build-statcast-database-rstats/) which provides access through statcast data from [baseballsavant]([https://baseballsavant.mlb.com)](https://baseballsavant.mlb.com)) .
+In comparsion, fans are able to get a breakdown of Mike Trout vs fastballs from using [baseballr](https://billpetti.github.io/2018-02-19-build-statcast-database-rstats/) which provides access through statcast data from [baseballsavant]([https://baseballsavant.mlb.com)](https://baseballsavant.mlb.com)).
 
@@
 
-However though using waRne we are able to not only compare Steve Smiths performance vs bowling type but we are able to easily compare Steve Smiths performance vs bowling types with other players.
+With the commentary data on cricinfo, we are not able to get the same resolution on delivery speed and location as the statcast data provides in baseball, but we can often learn the type of delivery the batsman faces through the years.
+This represents an enormous leap in publically available data, and we hope will lead to the eventual release of the ball tracking data used butnot shared in crickets leagues around the world
+
+Armed with this new dataset and the waRne package, we are able to not only compare Steve Smiths performance vs bowling type but we are able to easily compare Steve Smiths performance vs bowling types with other players, across multiple seasons, leagues, and cricket formats.
 
@@
 
+\color{red}{\text{So the way the scraper is set up, I grab the play links when scraping the scorecard and commentary along with the id and name, and then run a separate scraper for the batsman/bowler profiles.
+So they are separate datasets that have to be joined together, and we drop mostly everything except batting and bowling style.
+I can probably make an all in one scraper, but it'll be a bit tricky and if we want to put this out Friday it might be best to leave them separate and just walk people through it in the readme.
+Otherwise it might not work at all, and the scraper already breaks down so often you have to pay attention more than you should}}
+
 ```{r}
-library(tidyverse)
 
-# steve smith vs pace graph
+cricjam_data_big %>% 
+	dplyr::filter(batsman_id == 267192, 
+								!is.na(bowling_style)) %>% 
+	group_by(bowling_style, earned_runs) %>% 
+	summarise(total = n()) %>% 
+	group_by(bowling_style) %>% 
+	mutate(totals = sum(total),
+				 percent = total/totals, 
+				 earned_runs = as.factor(earned_runs)) %>% 
+	ggplot(aes(y = bowling_style, x = percent, color = earned_runs))+
+	geom_point() +
+	labs(color = "Runs", 
+			 x = "% Balls Faced", 
+			 y = "Bowling Style", 
+			 title = "Steve Smith's Distribution of Runs Against Spin and Pace")
 
@@
 
-# insert above plots here
+cricjam_data_big %>% 
+	dplyr::filter(tournament_name %in% c("Twenty20 Big Bash", "Big Bash League"), 
+								inning == 2) %>% 
+	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
+	group_by(match_id) %>% 
+	slice(1) %>% 
+	dplyr::select(batting_team_result, year) %>% 
+	mutate(binary_result = if_else(batting_team_result=="win", 1, 0)) %>% 
+	group_by(batting_team_result, year) %>% 
+	summarise(win_count = n()) %>% 
+	ungroup() %>% 
+	pivot_wider(names_from = batting_team_result, values_from = win_count) %>% 
+	rename(superover = 'super-over') %>% 
+	replace_na(list(superover=0)) %>% 
+	mutate(total_games = lose+win+superover, 
+				 win_percentage = win/total_games, 
+				 year = as.factor(year)) %>% 
+	ggplot(aes(x = year, y = win_percentage, group = 1)) +
+	geom_line(color = 'red')
+
 ```
@@
 ```{r}
+
+dat <- 
+	cricjam_data_big %>% 
+	dplyr::filter(tournament_name %in% c("Twenty20 Big Bash", "Big Bash League"), 
+								inning == 2) %>% 
+	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
+	group_by(match_id) %>% 
+	slice(1) %>% 
+	dplyr::select(batting_team_result, year) %>% 
+	mutate(binary_result = if_else(batting_team_result=="win", 1, 0)) %>% 
+	group_by(batting_team_result) %>% 
+	summarise(win_count = n()) %>% 
+	dplyr::filter(batting_team_result != 'super-over')
+
+N = sum(dat$win_count)
+wins = dat$win_count[2]
+
+z = (wins/N-0.5)/sqrt((0.5^2)/N)
+
+1-pnorm(abs(z))
+
+# No eveidence to suggest the 2 inning team wins more
 

### 2020-10-01 James_paper_copy.Rmd
update on the dhoni dilema

@@
 output:
-  pdf_document:
-    toc: yes
   html_document:
@@
     toc_float: yes
+  pdf_document:
+    toc: yes
 authors: Robert(1), James (2)
@@
   drop_na()%>% 
-  mutate(bowling_style = recode(bowling_style, "Right-arm fast" = "Pace",
+  mutate(pace_spin = bowling_style, "Right-arm fast" = "Pace",
                                 "Slow left-arm orthodox"= "Spin",
@@
                                                Slow left-arm chinaman"= "Spin", 
-                                "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" = "Spin"))
+                                "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" = "Spin")
 
@@
 
-1-pnorm(abs(z))
+2*(pnorm(z))
 
@@
 With this new dataset, we allow fans to put on their analyst thats and to be able to ask themselves, what exactly makes a good closer and what exactly is the Dhoni Dilemma.
-
 To understand the Dhoni Deilemma, we need to see some interesting things to do with the change in RRR.
@@
 Cricket is an obvious sort of sport, to win in the end you have to get one run more than your opposition.
-A commonly used statistic during the broadcast is the required run rate, for example if you need 120 to win off 20 overs your RRR is \$6\$ runs an over
-
+A commonly used statistic during the broadcast is the required run rate, for example if you need 120 to win off 20 overs your RRR is \$6\$ runs an over.
 What follows is then obvious is the chasing team wins if they are able to lower the RRR that is score quicker.
@@
 # graph of powerplay
+
+opening_batsman_of_interest <- c(
+  # 230559, #pollard
+  48472, #t dilshaan
+  37737, #brendon mccalum
+  8180, #shane watson
+  47270, #greme smith, 
+  28763, #gautham gambhir
+  226492, #martn guptil
+  49289, #mahela jayawardene
+  56194, #tamim iqbal
+  55608, #hamilton masakadza
+  51880, #gayle
+  41434, #mohommad hafeez
+  259410, #ahmed shehzad
+  24611, #william porterfield
+  # 49209, #sanath jayasuriya
+  43906, #hashim amla
+  16406, #michael lumb
+  # 42683, #salman butt
+  # 325012, #stoinis
+  # 272450, #MR Marsh
+  # 28081, #Dhoni
+  # 253802, #kholi
+  5334, #finch
+  219889, #warner
+  # 326434, #carey
+  267192, #smith
+  # 505120, #agar
+  # 530011, #head
+  # 5961, #henriques
+  # 325026, #maxwell
+  # 25913, #nabu
+  # 4864, #christian,
+  # 270484, #faulkner
+  # 230193, #wade
+  # 233680, #wells
+  # 500268, #turner
+  # 334337, #handscomb
+  # 603410, #mcdermott
+  # 308798  #short
+  # 787987, #labuschagne
+  # 1124282, # josh phillipe
+  # 326637, #Chris Lyrnn
+  # 5939, #hussey
+  # 931581 #pant
+  308967, # jos buttler
+  253802, #Kohli
+  249866  # alex hales
+)
+
+# Dot plot for openers
+cricjam_data_big %>% 
+  dplyr::filter(inning == 2, 
+                batsman_id %in% opening_batsman_of_interest, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over<6) %>% 
+  group_by(batsman_name) %>% 
+  summarise(change = mean(change_in_requireed_run_rate), 
+            total_balls_faced = n()) %>% 
+  ggplot(aes(x = total_balls_faced, y = change, label = batsman_name)) +
+  geom_point(size = 2) +
+  xlim(c(50, 1250))+
+  geom_text(vjust = 1.4, color = "black") +
+  theme_minimal()+
+  labs(x = "Balls Faced", 
+       y = "Average Change in Required Run Rate", 
+       title = "Top Opening Batsmen T20")
+  theme(axis.text = element_text(size = 12), 
+        axis.title = element_text(size = 16),
+        plot.title = element_text(size = 16))
+
 ```
 
-This graph makes sense as we can see the best openers in the world lower the required run rate so whats the Dhoni Dilemma?
+This graph makes sense as we can see some of the best openers in the world lower the required run rate. So what is the Dhoni Dilemma?
+
+Dhoni is considered the greatest closing batsman in the world.
+He has pulled of many spectacular saves, and has T20 and T20I strike rates after 14 years in the game.
+But how does his required run rate compare to other middle order and closing batsman
 
@@
 # middle overs (enter in the middle)
+
+middle_batsman_of_interest <- c(
+  230559, #pollard
+  32498, #yusef pathan
+  321777, #david miller
+  44936, #AB De Villiers
+  276298, #andre russell
+  8180, #shane watson
+  34102, #rohit sharma
+  # 51880, #gayle
+  # 325012, #stoinis
+  # 272450, #MR Marsh
+  28081, #Dhoni
+  # 253802, #kholi
+  # 5334, #finch
+  # 219889, #warner
+  # 326434, #carey
+  # 267192, #smith
+  # 505120, #agar
+  # 530011, #head
+  # 5961, #henriques
+  325026 #maxwell
+  # 25913, #nabu
+  # 4864, #christian,
+  # 270484, #faulkner
+  # 230193, #wade
+  # 233680, #wells
+  # 500268, #turner
+  # 334337, #handscomb
+  # 603410, #mcdermott
+  # 308798  #short
+  # 787987, #labuschagne
+  # 1124282, # josh phillipe
+  # 326637, #Chris Lyrnn
+  # 5939, #hussey
+  # 931581 #pant
+  # 308967, # job buttler
+  # 249866  # alex hales
+)
+
+cricjam_data_big %>% 
+  dplyr::filter(inning == 2, 
+                batsman_id %in% middle_batsman_of_interest, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # min(batsman_balls_faced_this_match)==1,
+                over<16, 
+                over>=6) %>% 
+  group_by(batsman_id, match_id) %>% 
+  mutate(min_balls = if_else(min(batsman_balls_faced_this_match)==1, 1, 0)) %>% 
+  ungroup() %>% 
+  # dplyr::filter(min_balls==1) %>%
+  group_by(batsman_name) %>% 
+	# dplyr::filter(batsman_id == 28081) %>% ggplot(aes(x = change_in_requireed_run_rate))+geom_histogram()
+  summarise(change = mean(change_in_requireed_run_rate), 
+            total_balls_faced = n()) %>% 
+  ggplot(aes(x = total_balls_faced, y = change, label = batsman_name)) +
+  geom_point() +
+	xlim(c(300, 1300))+
+  # geom_text(vjust = 1, color = "black") +
+  labs(x = "Balls Faced", 
+       y = "Average Change in Required Run Rate", 
+       title = "Top Middle Batsmen T20")+
+  geom_text(vjust = 1.4, color = "black") +
+  theme_minimal() 
+  theme(axis.text = element_text(size = 12), 
+        axis.title = element_text(size = 16),
+        plot.title = element_text(size = 16))
+
 ```
 
-Dhoni is renowed for being one of the worlds best closers but we can see it seems as though in the middle he is letting the game get out of hand. How can one of the worlds best closers
+Dhoni is renowed for being one of the worlds best closers but we can see it seems as though in the middle he is letting the game get out of hand.
+How can one of the worlds best closers let the run rate go up so consistently in the middle overs?
+
+One possible explanation comes from looking at the acceleration of Dhoni compared to these other middle order batsmen.
+Dhoni comes in usually around overs 11/12, and from the plot below we can see that he is a slow starter, it takes several overs for him to get his strike rate up, which consistently stays high throughout the game even when other batsmen's begins to drop off.
+
+```{r}
+
+cricjam_data_big %>%
+	dplyr::filter(batsman_id == 28081, 
+								batsman_balls_faced_this_match == 1) %>%
+	group_by(bucketed_over) %>% 
+	summarise(total = n()) %>% 
+	ungroup() %>% 
+	mutate(mean(total))
+
+cricjam_data_big %>% 
+    dplyr::filter(batsman_id %in% middle_batsman_of_interest, 
+                  !is.infinite(required_run_rate), 
+                  over>=6, 
+    							# over <16
+                  # batsman_id == 5961
+                  # batsman_balls_faced_this_match == 1
+                  ) %>% 
+    # dplyr::select(over, earned_runs, play_description) %>%
+    # group_by(earned_runs) %>% 
+    # summarise(n())
+    # group_by(batsman_id, match_id) %>% 
+    # mutate(min_balls = if_else(min(batsman_balls_faced_this_match)==1, 1, 0)) %>% 
+    # ungroup() %>%
+    group_by(batsman_balls_faced_this_match, batsman_name) %>%  # from here to the next comment is for sr, the rest is for 
+    summarise(sr = 100*sum(extra_runs+earned_runs)/n(), 
+              count = n()) %>% 
+    # dplyr::filter(batsman_balls_faced_this_match<50) %>% 
+    ggplot(aes(x = batsman_balls_faced_this_match, y = sr)) + 
+    geom_smooth() +
+    geom_hline(yintercept = 120) +
+    ylim(c(0,200))+
+    labs(x = "Balls Faced", 
+         y = "Strike Rate", 
+         title = "Batsman Acceleration In Overs 11-20")+
+    theme_minimal() +
+    theme(axis.text = element_text(size = 12), 
+          axis.title = element_text(size = 16),
+          plot.title = element_text(size = 16)) +
+    # ylim(c(50, 200))+
+    facet_wrap(~batsman_name)
+
+```
+
+Another is the effort of Dhoni's partners in moving the game along while Dhoni is up
+
+```{r}
+
+middle_batsman_of_interest <- c(
+  230559, #pollard
+  32498, #yusef pathan
+  321777, #david miller
+  44936, #AB De Villiers
+  276298, #andre russell
+  8180, #shane watson
+  34102, #rohit sharma
+  28081, #Dhoni
+  325026 #maxwell
+  )  
+maxwell <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 325026, 
+                inning == 2, resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Maxwell = mean(diff))  %>% 
+	mutate(batsman_partner_Maxwell = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 325026, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Maxwell = if_else(batsman_id == 325026, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Maxwell) %>% 
+  summarise(avg_Maxwell = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Maxwell", 
+				 avg = avg_Maxwell, 
+				 batsman_partner = batsman_partner_Maxwell) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+dhoni <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 28081, 
+                inning == 2, 
+  							resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Dhoni = mean(diff)) %>% 
+	mutate(batsman_partner_Dhoni = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 28081, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+														resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Dhoni = if_else(batsman_id == 28081, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Dhoni) %>% 
+  summarise(avg_Dhoni = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Dhoni", 
+				 avg = avg_Dhoni, 
+				 batsman_partner = batsman_partner_Dhoni) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+# sharma <-
+#   cricjam_data_big %>% 
+#   dplyr::filter(batsman_id %in% 34102, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr) %>% 
+#   group_by(bucketed_over) %>% 
+#   summarise(avg_Sharma = mean(diff)) %>% 
+# 	mutate(batsman_partner_Sharma = "Partnership") %>% 
+# 	bind_rows(cricjam_data_big %>% 
+# 							dplyr::filter(batsman_id %in% 34102, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr, 
+#   			 batsman_partner_Sharma = if_else(batsman_id == 34102, "Batsman", "Partner")) %>% 
+#   group_by(bucketed_over, batsman_partner_Sharma) %>% 
+#   summarise(avg_Sharma = mean(diff)) %>% 
+#   ungroup()) %>% 
+# 	mutate(batsman_name = "Sharma", 
+# 				 avg = avg_Sharma, 
+# 				 batsman_partner = batsman_partner_Sharma) %>% 
+# 	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+# watson <-
+#   cricjam_data_big %>% 
+#   dplyr::filter(batsman_id %in% 8180, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr) %>% 
+#   group_by(bucketed_over) %>% 
+#   summarise(avg_Watson = mean(diff)) %>% 
+# 	mutate(batsman_partner_Watson = "Partnership") %>% 
+# 	bind_rows(cricjam_data_big %>% 
+# 							dplyr::filter(batsman_id %in% 8180, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr, 
+#   			 batsman_partner_Watson = if_else(batsman_id == 8180, "Batsman", "Partner")) %>% 
+#   group_by(bucketed_over, batsman_partner_Watson) %>% 
+#   summarise(avg_Watson = mean(diff)) %>% 
+#   ungroup()) %>% 
+# 	mutate(batsman_name = "Watson", 
+# 				 avg = avg_Watson, 
+# 				 batsman_partner = batsman_partner_Watson) %>% 
+# 	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+russell <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 276298, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+  							resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Russell = mean(diff)) %>% 
+	mutate(batsman_partner_Russell = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 276298, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+														resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Russell = if_else(batsman_id == 276298, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Russell) %>% 
+  summarise(avg_Russell = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Russell", 
+				 avg = avg_Russell, 
+				 batsman_partner = batsman_partner_Russell) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+# villiers <-
+#   cricjam_data_big %>% 
+#   dplyr::filter(batsman_id %in% 44936, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr) %>% 
+#   group_by(bucketed_over) %>% 
+#   summarise(avg_DeVilliers = mean(diff)) %>% 
+# 	mutate(batsman_partner_DeVilliers = "Partnership") %>% 
+# 	bind_rows(cricjam_data_big %>% 
+# 							dplyr::filter(batsman_id %in% 44936, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr, 
+#   			 batsman_partner_DeVilliers = if_else(batsman_id == 44936, "Batsman", "Partner")) %>% 
+#   group_by(bucketed_over, batsman_partner_DeVilliers) %>% 
+#   summarise(avg_DeVilliers = mean(diff)) %>% 
+#   ungroup()) %>% 
+# 	mutate(batsman_name = "DeVilliers", 
+# 				 avg = avg_DeVilliers, 
+# 				 batsman_partner = batsman_partner_DeVilliers) %>% 
+# 	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+# miller <-
+#   cricjam_data_big %>% 
+#   dplyr::filter(batsman_id %in% 321777, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr) %>% 
+#   group_by(bucketed_over) %>% 
+#   summarise(avg_Miller = mean(diff)) %>% 
+# 	mutate(batsman_partner_Miller = "Partnership") %>% 
+# 	bind_rows(cricjam_data_big %>% 
+# 							dplyr::filter(batsman_id %in% 321777, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr, 
+#   			 batsman_partner_Miller = if_else(batsman_id == 321777, "Batsman", "Partner")) %>% 
+#   group_by(bucketed_over, batsman_partner_Miller) %>% 
+#   summarise(avg_Miller = mean(diff)) %>% 
+#   ungroup()) %>% 
+# 	mutate(batsman_name = "Miller", 
+# 				 avg = avg_Miller, 
+# 				 batsman_partner = batsman_partner_Miller) %>% 
+# 	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+# pathan <-
+#   cricjam_data_big %>% 
+#   dplyr::filter(batsman_id %in% 32498, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr) %>% 
+#   group_by(bucketed_over) %>% 
+#   summarise(avg_Pathan = mean(diff)) %>% 
+# 	mutate(batsman_partner_Pathan = "Partnership") %>% 
+# 	bind_rows(cricjam_data_big %>% 
+# 							dplyr::filter(batsman_id %in% 32498, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr, 
+#   			 batsman_partner_Pathan = if_else(batsman_id == 32498, "Batsman", "Partner")) %>% 
+#   group_by(bucketed_over, batsman_partner_Pathan) %>% 
+#   summarise(avg_Pathan = mean(diff)) %>% 
+#   ungroup())%>% 
+# 	mutate(batsman_name = "Pathan", 
+# 				 avg = avg_Pathan, 
+# 				 batsman_partner = batsman_partner_Pathan) %>% 
+# 	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+pollard <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 230559, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+  							resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Pollard = mean(diff)) %>% 
+	mutate(batsman_partner_Pollard = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 230559, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+														resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Pollard = if_else(batsman_id == 230559, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Pollard) %>% 
+  summarise(avg_Pollard = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Pollard", 
+				 avg = avg_Pollard, 
+				 batsman_partner = batsman_partner_Pollard) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+# marsh <- 
+#   cricjam_data_big %>% 
+#   dplyr::filter(batsman_id %in% 272450, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr) %>% 
+#   group_by(bucketed_over) %>% 
+#   summarise(avg_Marsh = mean(diff)) %>% 
+# 	mutate(batsman_partner_Marsh = "Partnership") %>% 
+# 	bind_rows(cricjam_data_big %>% 
+# 							dplyr::filter(batsman_id %in% 272450, 
+#                 inning == 2, 
+#                 # batsman_balls_faced_this_match<16, 
+#                 !is.infinite(required_run_rate), 
+#                 over < 16,
+#                 over >= 6 
+#                 # batsman_balls_faced_this_match==1, 
+#                 # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+#   ) %>% 
+#   distinct(match_id, bucketed_over) %>% 
+#   left_join(cricjam_data_big %>% 
+#               dplyr::filter(inning == 2, 
+#                             !is.infinite(required_run_rate), 
+#                             over < 16,
+#                             over >= 6), by = c("match_id" = "match_id", 
+#                                                "bucketed_over" = "bucketed_over")) %>% 
+#   group_by(bucketed_over, match_id) %>% 
+#   mutate(starting_rrr = first(required_run_rate), 
+#          ending_rrr = last(required_run_rate), 
+#          diff = ending_rrr-starting_rrr, 
+#   			 batsman_partner_Marsh = if_else(batsman_id == 272450, "Batsman", "Partner")) %>% 
+#   group_by(bucketed_over, batsman_partner_Marsh) %>% 
+#   summarise(avg_Marsh = mean(diff)) %>% 
+#   ungroup()) %>% 
+# 	mutate(batsman_name = "Marsh", 
+# 				 avg = avg_Marsh, 
+# 				 batsman_partner = batsman_partner_Marsh) %>% 
+# 	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+pollard %>% 
+	bind_rows(maxwell, dhoni, russell) %>% 
+	ggplot(aes(x = bucketed_over, y = avg, group = batsman_partner, color = batsman_partner)) +
+  geom_smooth()+
+  geom_hline(yintercept = 0)+
+  labs(y = "Required Run Rate", 
+       x = "Over", 
+       title = "Change in Required Run Rate While Batting (or Partnering)", 
+  		 color = "")+
+  facet_wrap(~batsman_name)+
+  theme_minimal()+
+  theme(axis.text = element_text(size = 12), 
+        axis.title = element_text(size = 16),
+        plot.title = element_text(size = 16))
+
+```
+
+```{r}
+
+middle_batsman_of_interest <- c(
+  230559, #pollard
+  32498, #yusef pathan
+  321777, #david miller
+  44936, #AB De Villiers
+  276298, #andre russell
+  8180, #shane watson
+  34102, #rohit sharma
+  28081, #Dhoni
+  325026 #maxwell
+  )  
+maxwell <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 325026, 
+                inning == 2, 
+  							resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Maxwell = mean(diff))  %>% 
+	mutate(batsman_partner_Maxwell = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 325026, 
+                inning == 2, 
+														resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Maxwell = if_else(batsman_id == 325026, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Maxwell) %>% 
+  summarise(avg_Maxwell = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Maxwell", 
+				 avg = avg_Maxwell, 
+				 batsman_partner = batsman_partner_Maxwell) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+dhoni <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 28081, 
+                inning == 2, 
+  							resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16  
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Dhoni = mean(diff), 
+  					total = n()) %>% 
+	mutate(batsman_partner_Dhoni = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 28081, 
+                inning == 2, 
+														resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Dhoni = if_else(batsman_id == 28081, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Dhoni) %>% 
+  summarise(avg_Dhoni = mean(diff), 
+  					total = n()) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Dhoni", 
+				 avg = avg_Dhoni, 
+				 batsman_partner = batsman_partner_Dhoni) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+russell <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 276298, 
+                inning == 2, 
+  							resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Russell = mean(diff)) %>% 
+	mutate(batsman_partner_Russell = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 276298, 
+                inning == 2, 
+														resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Russell = if_else(batsman_id == 276298, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Russell) %>% 
+  summarise(avg_Russell = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Russell", 
+				 avg = avg_Russell, 
+				 batsman_partner = batsman_partner_Russell) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+pollard <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 230559, 
+  							resources>10,
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Pollard = mean(diff)) %>% 
+	mutate(batsman_partner_Pollard = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 230559, 
+                inning == 2, 
+														resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+# over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Pollard = if_else(batsman_id == 230559, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Pollard) %>% 
+  summarise(avg_Pollard = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Pollard", 
+				 avg = avg_Pollard, 
+				 batsman_partner = batsman_partner_Pollard) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+pollard %>% 
+	bind_rows(maxwell, dhoni, russell) %>%
+	# dplyr::filter(bucketed_over <20) %>% 
+	ggplot(aes(x = bucketed_over, y = avg, group = batsman_partner, color = batsman_partner)) +
+  geom_smooth()+
+  geom_hline(yintercept = 0)+
+  labs(y = "Required Run Rate", 
+       x = "Over", 
+       title = "Change in Required Run Rate While Batting (or Partnering)", 
+  		 color = "")+
+  facet_wrap(~batsman_name)+
+  theme_minimal()+
+  theme(axis.text = element_text(size = 12), 
+        axis.title = element_text(size = 16),
+        plot.title = element_text(size = 16))
+
+```
+
+What we can tell is that in second inning chases, the required run rate for Dhoni is lower than our other top middle order batsmen, contrasting our inital findings.
+One would assume that this was due to Dhoni's partners, but it appears he still performs at an elite level
 

### 2020-10-02 paper_rob_working-.Rmd
push
push

--- 2020-09-26 paper.Rmd
@@
 output: 
+  pdf_document:
+    toc: yes
   html_document:
@@
 bibliography: references.bib
-
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(out.width = "75%",
-                      dev = "png",
-                      fig.align = "center",
-                      echo = FALSE)
+```{r setup,message=FALSE, include=FALSE}
+knitr::opts_chunk$set(message = TRUE, warning = TRUE)
+library(tidyverse)
+player_database <- 
+  read_csv("data/new_player_database_crickjam (2).csv") %>% 
+  dplyr::select(player_id, player_name, batting_style, bowling_style) %>% 
+  dplyr::distinct() %>% 
+  dplyr::filter(player_name != "Rohit Gurunath Sharma") %>% 
+  bind_rows(read_csv("data/new_player_database_crickjam (2).csv") %>% 
+              dplyr::filter(player_id == 34102) %>% 
+              dplyr::select(player_id, player_name, batting_style, bowling_style) %>% 
+              drop_na() %>% 
+              distinct()) %>% 
+  dplyr::filter(player_name !="Sultan Ahmed")
+
+bowler_vars <- 
+  player_database %>% 
+  dplyr::select(-c(batting_style, player_name)) %>% 
+  drop_na()%>% 
+  mutate(bowling_style = recode(bowling_style, "Right-arm fast" = "Pace",
+                                "Slow left-arm orthodox"= "Spin",
+                                "Legbreak"= "Spin",
+                                "Right-arm medium"= "Pace",
+                                "Left-arm medium-fast"= "Pace",
+                                "Right-arm offbreak"= "Spin",
+                                "Legbreak googly"= "Spin",
+                                "Left-arm fast"= "Pace",
+                                "Right-arm medium-fast"= "Pace",
+                                "Left-arm fast-medium"= "Pace",
+                                "Right-arm fast-medium"= "Pace",
+                                "Left-arm medium"= "Pace",
+                                "Slow left-arm chinaman"= "Spin",
+                                "Right-arm offbreak, Legbreak"= "Spin",
+                                "Right-arm medium, Right-arm offbreak"= "Spin",
+                                "Left-arm medium, Slow left-arm orthodox, 
+                                               Slow left-arm chinaman"= "Spin", 
+                                "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" = "Spin"))
+
+batting_vars <- 
+  player_database %>% 
+  dplyr::select(-c(bowling_style, player_name)) %>%
+  drop_na()
+
+cricjam_data <- 
+  read_csv("data/combined_2020_09_11.csv") %>% 
+  dplyr::filter(!grepl("No result", match_summary)) %>%
+  mutate(bucketed_year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
+  left_join(batting_vars, by = c("batsman_id" = "player_id")) %>% 
+  left_join(bowler_vars, by = c("bowler_id" = "player_id")) %>% 
+  mutate(batsman_name = str_remove(batsman_name, "\\(\\)"), 
+         batsman_name = str_trim(batsman_name), 
+         runs = extra_runs+earned_runs)
+# 
+# cricjam_data_partner <- 
+#   cricjam_data %>% 
+#   dplyr::select(match_id, inning, batsman_name, batsman_id, batting_order) %>% 
+#   distinct() %>% 
+#   arrange(batting_order) %>% View()
+
+wicket_types <-
+  cricjam_data %>% 
+  dplyr::filter(wicket==TRUE) %>% 
+  dplyr::select(wicket_description, play_description, bowler, batsman, bowler_name, batsman_name, 
+                match_id, inning, over) %>%
+  mutate(caught_out = if_else(str_detect(wicket_description, " c "), TRUE, FALSE), 
+         bowled_out = if_else(str_detect(wicket_description, " b "), TRUE, FALSE), 
+         run_out = if_else(str_detect(wicket_description, " run out "), TRUE, FALSE), 
+         stumped_out = if_else(str_detect(wicket_description, " st "), TRUE, FALSE), 
+         leg_before_wicket = if_else(str_detect(wicket_description, " lbw "), TRUE, FALSE), 
+         retired_hurt = if_else(str_detect(wicket_description, " retired hurt "), TRUE, FALSE), 
+         obstructing_the_field = if_else(str_detect(wicket_description, " obstructing the field "), TRUE, FALSE), 
+         retired_out = if_else(str_detect(wicket_description, " retired out "), TRUE, FALSE), 
+         null = if_else(str_detect(wicket_description, " null "), TRUE, FALSE)) %>% 
+  # all_conditions = if_else(caught_out == TRUE |
+  #                          bowled_out == TRUE | caught_out == TRUE |
+  #                            run_out == TRUE |retired_out == TRUE |
+  #                            retired_hurt == TRUE |leg_before_wicket == TRUE |
+  #                            obstructing_the_field == TRUE |null == TRUE |
+  #                            stumped_out == TRUE, TRUE, FALSE))
+  mutate(out_condition = case_when(caught_out==TRUE |
+                                     caught_out==TRUE & bowled_out == TRUE ~ "Caught", 
+                                   run_out == TRUE |
+                                     run_out==TRUE & bowled_out == TRUE ~ "Run Out", 
+                                   stumped_out == TRUE |
+                                     stumped_out==TRUE & bowled_out == TRUE ~ "Stumped", 
+                                   leg_before_wicket == TRUE |
+                                     leg_before_wicket==TRUE & bowled_out == TRUE ~ "Leg Before Wicket", 
+                                   retired_hurt == TRUE |
+                                     retired_hurt==TRUE & bowled_out == TRUE ~ "Retired Hurt", 
+                                   obstructing_the_field == TRUE |
+                                     obstructing_the_field==TRUE & bowled_out == TRUE ~ "Obstructing the Field", 
+                                   retired_out == TRUE |
+                                     retired_out==TRUE & bowled_out == TRUE ~ "Retired Out", 
+                                   null == TRUE |
+                                     null==TRUE & bowled_out == TRUE ~ "Caught", 
+                                   bowled_out == TRUE & caught_out == FALSE |
+                                     bowled_out == TRUE & stumped_out == FALSE |
+                                     bowled_out == TRUE & run_out == FALSE |
+                                     bowled_out == TRUE & leg_before_wicket == FALSE |
+                                     bowled_out == TRUE & retired_hurt == FALSE |
+                                     bowled_out == TRUE & obstructing_the_field == FALSE |
+                                     bowled_out == TRUE & retired_out == FALSE |
+                                     bowled_out == TRUE & null == FALSE ~"Bowled"), 
+         out_batsman = case_when(out_condition == "Bowled" ~ word(wicket_description, 1, sep = " b "), 
+                                 out_condition == "Caught" & null == FALSE ~ word(wicket_description, 1, sep = " c "), 
+                                 out_condition == "Caught" & null == TRUE ~ word(wicket_description, 1, sep = " null "), 
+                                 out_condition == "Run Out" ~ word(wicket_description, 1, sep = " run out "), 
+                                 out_condition == "Stumped" ~ word(wicket_description, 1, sep = " st "), 
+                                 out_condition == "Leg Before Wicket" ~ word(wicket_description, 1, sep = " lbw "), 
+                                 out_condition == "Retired Hurt" ~ word(wicket_description, 1, sep = " retired hurt "), 
+                                 out_condition == "Obstructing the Field" ~ word(wicket_description, 1, sep = " obstructing the field "), 
+                                 out_condition == "Retired Out" ~ word(wicket_description, 1, sep = " retired out ")),  
+         out_batsman = str_trim(out_batsman)) %>% 
+  dplyr::select(-c(caught_out, bowled_out, stumped_out, leg_before_wicket, retired_hurt, 
+                   obstructing_the_field, retired_hurt, null, run_out, retired_out)) %>% 
+  mutate(batsman_or_partner_out = if_else(out_batsman == batsman_name, "Batsman","Partner"))
+
+cricjam_data <- 
+  cricjam_data %>% 
+  left_join(wicket_types, by = c("wicket_description" = "wicket_description", 
+                                 "play_description" = "play_description", 
+                                 "bowler" = "bowler", 
+                                 "batsman" = "batsman", 
+                                 "batsman_name" = "batsman_name", 
+                                 "match_id" = "match_id", 
+                                 "inning" = "inning", 
+                                 "over" = "over", 
+                                 "bowler_name" = "bowler_name")) %>% 
+  rename(inning_summary = inning_total)
+
+only_big_t20_matches <- 
+  cricjam_data %>% 
+  dplyr::filter(tournament_type == "Twenty20", 
+                tournament_name %in% c("Twenty20 Big Bash", "Indian Premier League", "Big Bash League", 
+                                       "Champions League Twenty20", "Pepsi Indian Premier League")) %>% 
+  dplyr::select(match_id) %>% 
+  distinct() %>% 
+  pull(match_id)
+
+cricjam_data_big <- 
+  cricjam_data %>% 
+  dplyr::filter(tournament_type == "Twenty20 Internationals") %>% 
+  bind_rows(cricjam_data %>% 
+              dplyr::filter(match_id %in% only_big_t20_matches))
 ```
@@
 library(tidyverse)
-
-# steve smith vs pace graph
+cricjam_data_big %>% 
+	dplyr::filter(batsman_id == 267192, 
+								!is.na(bowling_style)) %>% 
+	group_by(bowling_style, earned_runs) %>% 
+	summarise(total = n()) %>% 
+	group_by(bowling_style) %>% 
+	mutate(totals = sum(total),
+				 percent = total/totals, 
+				 earned_runs = as.factor(earned_runs)) %>% 
+	ggplot(aes(y = bowling_style, x = percent, color = earned_runs))+
+	geom_point() +
+	labs(color = "Runs", 
+			 x = "% Balls Faced", 
+			 y = "Bowling Style", 
+			 title = "Steve Smith's Distribution of Runs Against Spin and Pace")
 
 ```
-
-Another reason why is that cricket is a weird sport with a lot of old school rules of thumb that are followed/talked about.
-
-For example it has become popular conjecture that its better that teams should bat second (chase down a total set by the oppoition). Through waRne we can answer such question for the T20 game.
-
-For example we can explore do teams win more often when chasing in the big bash over the history of the big bash, or we can plot a line chart to see if there is a movement towards batting last and winning.
 
@@
 For example using the above we can not only look at the answers graphically, but we can run a statistical test.
-The example we would use is instead of testing is a coin biased (different from 50/50) we can test to see if the winning \\% is different from $0.5$ for teams chasing.
+A common classroom example for learning the binomial distribution is to ask if a coin is based given a proportion of heads and tails given a sample size [@mcelreath2020statistical].
+Instead of asking about coins, fans of sport and cricket might be interested to ask the question; after winning the coin toss, should a team decide to bat first or bat second [@kvam2004teaching].
+
+While a coin toss is a seemingly trivial example, like in most professional sports the winner of the coin toss gets to decide what to do.
+In cricket, the winner of the coin toss gets to decide if they want to bat first and thus set the total that the opposition needs to pass by a single run to win, or if they want to bowl first and thus chase down the total set.
+Deciding what to do after winning a coin toss has proved to be a popular media piece in recent years\\footnote{<https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss> }\\footnote{https[://www.espncricinfo.com/video/clip/\_/id/23230682%5D](https://www.espncricinfo.com/video/clip/_/id/23230682%5D){.uri}}
+
+\\footnote{[https://www.espncricinfo.com/story/\_/id/21489056/stuart-wark-cricket-move-away-coin-toss](https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss?%5D)}
+
+\^[[https://www.espncricinfo.com/story/\_/id/20429499/lehmann-backs-scrapping-toss]](https://www.espncricinfo.com/story/_/id/20429499/lehmann-backs-scrapping-toss%5D) \^[[https://www.espncricinfo.com/story/\_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams]](https://www.espncricinfo.com/story/_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams%5D) \^[<https://www.forbes.com/sites/tristanlavalette/2018/08/20/are-cricket-matches-being-decided-by-the-luck-of-a-coin-toss/#735456837eff>]
+
+\^[<https://www.statsinsider.com.au/bbl/how-important-is-winning-the-toss-in-the-big-bash-league>]
+
+\^[<https://www.espncricinfo.com/story/_/id/18568387/tim-wigmore-how-batting-second-become-more-fruitful-more-popular>]
+
+Fans and analysts of the game might want to make this decision based on a simple statistic, for example they might make it based on answer these questions instead.
+
+-   Do teams that bat second have a higher winning percentage than those who bat first?
+
+-   Is this consistent across leagues and levels of cricket?
 
@@
 
+cricjam_data_big %>% 
+	dplyr::filter(tournament_name %in% c("Twenty20 Big Bash", "Big Bash League"), 
+								inning == 2) %>% 
+	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
+	group_by(match_id) %>% 
+	slice(1) %>% 
+	dplyr::select(batting_team_result, year) %>% 
+	mutate(binary_result = if_else(batting_team_result=="win", 1, 0)) %>% 
+	group_by(batting_team_result, year) %>% 
+	summarise(win_count = n()) %>% 
+	ungroup() %>% 
+	pivot_wider(names_from = batting_team_result, values_from = win_count) %>% 
+	rename(superover = 'super-over') %>% 
+	replace_na(list(superover=0)) %>% 
+	mutate(total_games = lose+win+superover, 
+				 win_percentage = win/total_games, 
+				 year = as.factor(year)) %>% 
+	ggplot(aes(x = year, y = win_percentage, group = 1)) +
+	geom_line(color = 'red')
 ```
 
-James and I entered a reproducibility competition and the winning entry provided a statistic sort of a gateway into the importance of their ''project'' was was specifically in 2019 Australian Big Bash T20 season.
-The statistic given was in the 2019 big bash season teams that batted first won only $43$% of games.
-Thankfully we can try to reproduce that statistic now through the use of waRne
+Instead of just looking at this graphically, fans of crickets and teachers of undergraduate statistics might use the dataset as a ''biased coin'' example.
+So instead of asking, given a proportion of heads over \$n\$ tosses instructors could ask is a team better off batting first or second.
 
-Wikipedia tells us that there were the following:
+```{r}
 
-In the 2018-19 Big Bash League there were 59 matches played. Of which we had 3 DLS results.
+n<-900
+p<-0.5
+qbinom(c(0.025,0.975),size=n,prob=p)
 
-| Games                             | Result                        |
-|-----------------------------------|-------------------------------|
-| 21 DEC 2018 - THUNDER VS STARS    | THUNDER WON BY 15 RUNS (D/L)  |
-| 2 FEB 2019 - THUNDER VS SIXERS    | SIXERS WON BY 9 WICKETS (D/L) |
-| 8 JAN 2019 HEAT VS THUNDER        | HEAT BY 15 RUNS (D/L)         |
-| 17th January 2019 Thunder vs Heat | no result                     |
-
-: Games under consideration in 2018-19 Big Bash
-
-```{r, investigation}
-df<-read.csv("data/combined_2020_09_11.csv")
-
-blah<-df%>%
-	dplyr::select(duckworth_lewis,year,match_id, tournament_name, tournament_type, team, inning, binary_result)%>%
-	dplyr::filter(tournament_type%in% c("Twenty20 Internationals", "Twenty20"  ))%>%
-	dplyr::filter(tournament_name%in% c("Twenty20 Big Bash","Big Bash League" ))%>%
-									dplyr::filter(inning==1)%>%
-									dplyr::group_by()%>%
-									distinct()
-# dataset doesn't contain duckworth lewis results. 
-table(blah$year,blah$binary_result)
 ```
 
-\`\`\`
+```{r}
+df<-data.frame(
+             Year = c(2010L,2011L,2012L,2013L,2014L,
+                      2015L,2016L,2017L,2018L,2019L,2020L,2021L,2022L),
+  number.of.heads = c(52L,55L,51L,70L,30L,40L,35L,
+                      43L,49L,37L,39L,45L,32L)
+)
 
-So here we can see that we in the big bash season 2018-19 we have a current winning percentage batting first of 22/(22+33) =0.4 but this doesn't include the duckworth lewis games or the washed out completely game.
+library(broom)
+df %>%
+  # create variables to pass to binom.test
+  mutate(n = 100, 
+         x = n - number.of.heads, 
+         p = 0.5,
+         alternative = "two.sided",
+         conf.level = 0.95) %>%
+  # Run a binom.test for each row/year
+  mutate(result = pmap(list(x, n,p), binom.test)) %>% 
+  # Extract results from test into dataframe using broom::tidy
+  mutate(result = map(result, tidy)) %>%
+  mutate(result = map(result, ~select(.x, p.value, conf.low, conf.high))) %>% 
+  unnest(cols = c(result)) %>%
+  # plot the annual confidence intervals
+  ggplot(aes(factor(Year), p.value)) + 
+    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, color = (p.value > 0.05)), width=.1) +
+  theme_minimal() +
+  labs(y = "P(coin is fair)", x = "Year") +
+  ggtitle("Annual coin fairness test")
 
-If we were to add in the duckworth lewis games we get 24/(24+34) which is equal to 0.414. Which isn't the result.
+```
 
-We can also clearly see that in the 20192020 big bash season the team batting first wins 39/(39+20) or 0.66 so what could be going on? Did Quantium mean only games played in the actual year 2019?
+## James Stein Estimator - Why not a cricket example
 
-In that case from the 2018-2019 season we would remove the following games
+<https://bookdown.org/content/922/james-stein.html>
 
-| Date                 | Game                    | Result                   | Winning Team |
-|----------------------|-------------------------|--------------------------|--------------|
-| 19 December 2018     | Heat Vs Strickers       | Strickers 5 Wickets      | Chased       |
-| 20th December 2018   | Scorchers Vs Renegades  | Renegades by 4 wickets   | Chased       |
-| 21 December 2018     | Thunder vs Stars        | Thunder by 15 runs (D/L) | Batted First |
-| 22 December 2018     | Sixers vs Scorchers     | Sixers by 17 runs        | Batted First |
-| 22 December          | Hurricanes vs Heat      | Hurricanes by 15 runs    | Batted First |
-| 23 December 2018     | Strikers vs Renegades   | Renegades by 5 wickets   | Chased       |
-| 24 December 2018     | Stars vs Hurricanes     | Hurricanes by 6 wickets  | Chased       |
-| 24th December 2018   | Thunder vs sixers       | Thunder by 21 runs       | Batted First |
-| 26th December 2018   | Strickers vs scorchers  | Scorchers by 7 wickets   | Chased       |
-| 27th December        | Sixers vs Stars         | stars by 5 wickets       | Chased       |
-| 28th December        | thunder vs hurricanes   | hurricanes by 7 wickets  | Chased       |
-| 29th december        | sixers vs renegades     | sixers by 33 runs        | Batted First |
-| 30th decemeber 2018  | scorchers vs hurricanes | hurricanes by 6 wickets  | Chased       |
-| 31st decemember 2018 | strickers vs thunder    | strickers by 20 runs     | Batted First |
-
-: Games in 2018 of the 2018-19 Big Bash Season
-
-i.e. we remove 14 games here of which 6 teams won batting first 16/(22+33-14)
-
-Then from here we have to add on games in the 2019-2020 big bash season that were played in 2019
-
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| Date               | Teams                   | Result                                | Winning Team             |
-+====================+=========================+=======================================+==========================+
-| 17 December        | Thunder vs Heat         | Thunder by 29 runs                    | Batted First             |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 18th December      | Scorchers vs Sixers     | Sixers by 8 wickets                   | Chased                   |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 19th december      | Renegades vs thunder    | thunder by 6 wickets                  | chased                   |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 20th decemeber     | hurricanes vs sixers    | hurricanes by 25 runs                 | batted first             |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 20th decemember    | stars vs heat           | stars by 22 runs                      | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 21st decemember    | strickers vs thunder    | no result                             | no result                |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 21st december      | scorchers vs renegades  | scorchers by 11 runs                  | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 22 deceember       | stars vs hurricanes     | stars by 52 runs                      | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 22 decemember      | heat vs sixers          | heat by 48 runs                       | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 23 dec             | strickers vs scorchers  | strickers by 15 (DLS)                 | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 24th dec           | renegades vs hurricanes | hurricanes by 7 wickets               | chased                   |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 26 decemember      | sixers vs scorchers     | sixers by 48 runs                     | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 27 dec             | strickers vs stars      | strikers by 5 runs                    | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 28th december      | thunder vs sixers       | sixers won super over (batting first) | sixers by superover      |
-|                    |                         |                                       |                          |
-|                    |                         |                                       | or sixers batting second |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 29th dec           | strikers vs renegades   | strikers by 18 runs                   | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 30th dec           | hurricanes vs stars     | stars by 4 runs (DLS)                 | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-| 31st december 2019 | thunder vs strikers     | thunder by 3 runs                     | batting first            |
-+--------------------+-------------------------+---------------------------------------+--------------------------+
-
-: Games played in 2019 of the 2019-20 Big Bash
-
-12 batting first and 3 batting second
-
-(16+12)/(12+3+22+33-14)
-
-So over the past 2 years (not including DLS) we have
-
-(22+39)/(20+39+33+22)
-
-i.e. \>50% of teams win chasing so what does that mean?
+```{r}
+```
 
@@
 
-To the surprise of many, its hard to engage cricket fans into the analytics behind the game. Reasons for this are generally centred around reproducibility and explainability to fans.
+To the surprise of many, there has not been a real push to engage fans in the analytics of the game of cricket.
+Unlike other sports which has seen a boom through the use of accessible data like Ice Hockey, NFL, Baseball and Basketball for example.
 
@@
 
-+------------------------------+--------------------------------------------------------------------------+
-| Things to do                 | person                                                                   |
-+==============================+==========================================================================+
-| clean scrapers/csv for waRne | James                                                                    |
-+------------------------------+--------------------------------------------------------------------------+
-| data dictionary              | Robert                                                                   |
-+------------------------------+--------------------------------------------------------------------------+
-| examples                     | Robert - start (assume same csv being used) - cool cricket example stuff |
-+------------------------------+--------------------------------------------------------------------------+
-
 ## Dhoni Dilemma
@@
 
-```{r}
+```{r,message=FALSE, echo=FALSE, warning=FALSE}
 # graph of powerplay
+
+opening_batsman_of_interest <- c(
+  # 230559, #pollard
+  48472, #t dilshaan
+  37737, #brendon mccalum
+  8180, #shane watson
+  47270, #greme smith, 
+  28763, #gautham gambhir
+  226492, #martn guptil
+  49289, #mahela jayawardene
+  56194, #tamim iqbal
+  55608, #hamilton masakadza
+  51880, #gayle
+  41434, #mohommad hafeez
+  259410, #ahmed shehzad
+  24611, #william porterfield
+  # 49209, #sanath jayasuriya
+  43906, #hashim amla
+  16406, #michael lumb
+  # 42683, #salman butt
+  # 325012, #stoinis
+  # 272450, #MR Marsh
+  # 28081, #Dhoni
+  # 253802, #kholi
+  5334, #finch
+  219889, #warner
+  # 326434, #carey
+  267192, #smith
+  # 505120, #agar
+  # 530011, #head
+  # 5961, #henriques
+  # 325026, #maxwell
+  # 25913, #nabu
+  # 4864, #christian,
+  # 270484, #faulkner
+  # 230193, #wade
+  # 233680, #wells
+  # 500268, #turner
+  # 334337, #handscomb
+  # 603410, #mcdermott
+  # 308798  #short
+  # 787987, #labuschagne
+  # 1124282, # josh phillipe
+  # 326637, #Chris Lyrnn
+  # 5939, #hussey
+  # 931581 #pant
+  308967, # jos buttler
+  253802, #Kohli
+  249866  # alex hales
+)
+
+# Dot plot for openers
+cricjam_data_big %>% 
+  dplyr::filter(inning == 2, 
+                batsman_id %in% opening_batsman_of_interest, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over<6) %>% 
+  group_by(batsman_name) %>% 
+  summarise(change = mean(change_in_requireed_run_rate), 
+            total_balls_faced = n()) %>% 
+  ggplot(aes(x = total_balls_faced, y = change, label = batsman_name)) +
+  geom_point(size = 2) +
+  xlim(c(50, 1250))+
+  geom_text(vjust = 1.4, color = "black") +
+  labs(x = "Balls Faced", 
+       y = "Average Change in Required Run Rate", 
+       title = "Top Opening Batsmen T20")
+  theme_minimal()
 ```
 
-This graph makes sense as we can see the best openers in the world lower the required run rate so whats the Dhoni Dilemma?
+This graph makes sense as we can see some of the best openers in the world lower the required run rate. So what is the Dhoni Dilemma?
 
-```{r}
+Dhoni is considered the greatest closing batsman in the world.
+He has pulled of many spectacular saves, and has T20 and T20I strike rates after 14 years in the game.
+But how does his required run rate compare to other middle order and closing batsman
+
+```{r,message=FALSE, echo=FALSE, warning=FALSE}
 # middle overs (enter in the middle)
+
+middle_batsman_of_interest <- c(
+  230559, #pollard
+  32498, #yusef pathan
+  321777, #david miller
+  44936, #AB De Villiers
+  276298, #andre russell
+  8180, #shane watson
+  34102, #rohit sharma
+  # 51880, #gayle
+  # 325012, #stoinis
+  # 272450, #MR Marsh
+  28081, #Dhoni
+  # 253802, #kholi
+  # 5334, #finch
+  # 219889, #warner
+  # 326434, #carey
+  # 267192, #smith
+  # 505120, #agar
+  # 530011, #head
+  # 5961, #henriques
+  325026 #maxwell
+  # 25913, #nabu
+  # 4864, #christian,
+  # 270484, #faulkner
+  # 230193, #wade
+  # 233680, #wells
+  # 500268, #turner
+  # 334337, #handscomb
+  # 603410, #mcdermott
+  # 308798  #short
+  # 787987, #labuschagne
+  # 1124282, # josh phillipe
+  # 326637, #Chris Lyrnn
+  # 5939, #hussey
+  # 931581 #pant
+  # 308967, # job buttler
+  # 249866  # alex hales
+)
+
+cricjam_data_big %>% 
+  dplyr::filter(inning == 2, 
+                batsman_id %in% middle_batsman_of_interest, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # min(batsman_balls_faced_this_match)==1,
+                over<16, 
+                over>=6) %>% 
+  group_by(batsman_id, match_id) %>% 
+  mutate(min_balls = if_else(min(batsman_balls_faced_this_match)==1, 1, 0)) %>% 
+  ungroup() %>% 
+  # dplyr::filter(min_balls==1) %>%
+  group_by(batsman_name) %>% 
+	# dplyr::filter(batsman_id == 28081) %>% ggplot(aes(x = change_in_requireed_run_rate))+geom_histogram()
+  summarise(change = mean(change_in_requireed_run_rate), 
+            total_balls_faced = n()) %>% 
+  ggplot(aes(x = total_balls_faced, y = change, label = batsman_name)) +
+  geom_point() +
+	xlim(c(300, 1300))+
+  # geom_text(vjust = 1, color = "black") +
+  labs(x = "Balls Faced", 
+       y = "Average Change in Required Run Rate", 
+       title = "Top Middle Batsmen T20")+
+  geom_text(vjust = 1.4, color = "black") +
+  theme_minimal()
+
 ```
 
-Dhoni is renowed for being one of the worlds best closers but we can see it seems as though in the middle he is letting the game get out of hand. How can one of the worlds best closers
+Dhoni is renowed for being one of the worlds best closers but we can see it seems as though in the middle he is letting the game get out of hand.
+How can one of the worlds best closers let the run rate go up so consistently in the middle overs?
+
+One possible explanation comes from looking at the acceleration of Dhoni compared to these other middle order batsmen.
+Dhoni comes in usually around overs 11/12, and from the plot below we can see that he is a slow starter, it takes several overs for him to get his strike rate up, which consistently stays high throughout the game even when other batsmen's begins to drop off.
+
+```{r,message=FALSE, echo=FALSE, warning=FALSE}
+
+cricjam_data_big %>%
+	dplyr::filter(batsman_id == 28081, 
+								batsman_balls_faced_this_match == 1) %>%
+	group_by(bucketed_over) %>% 
+	summarise(total = n()) %>% 
+	ungroup() %>% 
+	mutate(mean(total))
+
+cricjam_data_big %>% 
+    dplyr::filter(batsman_id %in% middle_batsman_of_interest, 
+                  !is.infinite(required_run_rate), 
+                  over>=6, 
+    							# over <16
+                  # batsman_id == 5961
+                  # batsman_balls_faced_this_match == 1
+                  ) %>% 
+    # dplyr::select(over, earned_runs, play_description) %>%
+    # group_by(earned_runs) %>% 
+    # summarise(n())
+    # group_by(batsman_id, match_id) %>% 
+    # mutate(min_balls = if_else(min(batsman_balls_faced_this_match)==1, 1, 0)) %>% 
+    # ungroup() %>%
+    group_by(batsman_balls_faced_this_match, batsman_name) %>%  # from here to the next comment is for sr, the rest is for 
+    summarise(sr = 100*sum(extra_runs+earned_runs)/n(), 
+              count = n()) %>% 
+    # dplyr::filter(batsman_balls_faced_this_match<50) %>% 
+    ggplot(aes(x = batsman_balls_faced_this_match, y = sr)) + 
+    geom_smooth() +
+    geom_hline(yintercept = 120) +
+    ylim(c(0,200))+
+    labs(x = "Balls Faced", 
+         y = "Strike Rate", 
+         title = "Batsman Acceleration In Overs 11-20")+
+    theme_minimal() +
+     facet_wrap(~batsman_name)
+
+```
+
+Another is the effort of Dhoni's partners in moving the game along while Dhoni is up
+
+```{r, message=FALSE,echo=FALSE, warning=FALSE}
+
+middle_batsman_of_interest <- c(
+  230559, #pollard
+  32498, #yusef pathan
+  321777, #david miller
+  44936, #AB De Villiers
+  276298, #andre russell
+  8180, #shane watson
+  34102, #rohit sharma
+  28081, #Dhoni
+  325026 #maxwell
+  )  
+maxwell <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 325026, 
+                inning == 2, resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Maxwell = mean(diff))  %>% 
+	mutate(batsman_partner_Maxwell = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 325026, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Maxwell = if_else(batsman_id == 325026, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Maxwell) %>% 
+  summarise(avg_Maxwell = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Maxwell", 
+				 avg = avg_Maxwell, 
+				 batsman_partner = batsman_partner_Maxwell) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+dhoni <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 28081, 
+                inning == 2, 
+  							resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Dhoni = mean(diff)) %>% 
+	mutate(batsman_partner_Dhoni = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 28081, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+														resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Dhoni = if_else(batsman_id == 28081, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Dhoni) %>% 
+  summarise(avg_Dhoni = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Dhoni", 
+				 avg = avg_Dhoni, 
+				 batsman_partner = batsman_partner_Dhoni) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+russell <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 276298, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+  							resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Russell = mean(diff)) %>% 
+	mutate(batsman_partner_Russell = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 276298, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+														resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Russell = if_else(batsman_id == 276298, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Russell) %>% 
+  summarise(avg_Russell = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Russell", 
+				 avg = avg_Russell, 
+				 batsman_partner = batsman_partner_Russell) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+pollard <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 230559, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+  							resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,
+              							resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Pollard = mean(diff)) %>% 
+	mutate(batsman_partner_Pollard = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 230559, 
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                over < 16,
+														resources>10,
+                over >= 6 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+                            !is.infinite(required_run_rate), 
+                            over < 16,resources>10,
+                            over >= 6), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Pollard = if_else(batsman_id == 230559, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Pollard) %>% 
+  summarise(avg_Pollard = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Pollard", 
+				 avg = avg_Pollard, 
+				 batsman_partner = batsman_partner_Pollard) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+pollard %>% 
+	bind_rows(maxwell, dhoni, russell) %>% 
+	ggplot(aes(x = bucketed_over, y = avg, group = batsman_partner, color = batsman_partner)) +
+  geom_smooth()+
+  geom_hline(yintercept = 0)+
+  labs(y = "Required Run Rate", 
+       x = "Over", 
+       title = "Change in Required Run Rate While Batting (or Partnering)", 
+  		 color = "")+
+  facet_wrap(~batsman_name)+
+  theme_minimal()
+
+```
+
+```{r,message=FALSE,echo=FALSE, warning=FALSE}
+
+middle_batsman_of_interest <- c(
+  230559, #pollard
+  32498, #yusef pathan
+  321777, #david miller
+  44936, #AB De Villiers
+  276298, #andre russell
+  8180, #shane watson
+  34102, #rohit sharma
+  28081, #Dhoni
+  325026 #maxwell
+  )  
+maxwell <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 325026, 
+                inning == 2, 
+  							resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Maxwell = mean(diff))  %>% 
+	mutate(batsman_partner_Maxwell = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 325026, 
+                inning == 2, 
+														resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Maxwell = if_else(batsman_id == 325026, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Maxwell) %>% 
+  summarise(avg_Maxwell = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Maxwell", 
+				 avg = avg_Maxwell, 
+				 batsman_partner = batsman_partner_Maxwell) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+dhoni <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 28081, 
+                inning == 2, 
+  							resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16  
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Dhoni = mean(diff), 
+  					total = n()) %>% 
+	mutate(batsman_partner_Dhoni = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 28081, 
+                inning == 2, 
+														resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Dhoni = if_else(batsman_id == 28081, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Dhoni) %>% 
+  summarise(avg_Dhoni = mean(diff), 
+  					total = n()) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Dhoni", 
+				 avg = avg_Dhoni, 
+				 batsman_partner = batsman_partner_Dhoni) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+russell <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 276298, 
+                inning == 2, 
+  							resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Russell = mean(diff)) %>% 
+	mutate(batsman_partner_Russell = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 276298, 
+                inning == 2, 
+														resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Russell = if_else(batsman_id == 276298, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Russell) %>% 
+  summarise(avg_Russell = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Russell", 
+				 avg = avg_Russell, 
+				 batsman_partner = batsman_partner_Russell) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+pollard <-
+  cricjam_data_big %>% 
+  dplyr::filter(batsman_id %in% 230559, 
+  							resources>10,
+                inning == 2, 
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+                # over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr) %>% 
+  group_by(bucketed_over) %>% 
+  summarise(avg_Pollard = mean(diff)) %>% 
+	mutate(batsman_partner_Pollard = "Partnership") %>% 
+	bind_rows(cricjam_data_big %>% 
+							dplyr::filter(batsman_id %in% 230559, 
+                inning == 2, 
+														resources>10,
+                # batsman_balls_faced_this_match<16, 
+                !is.infinite(required_run_rate), 
+# over < 16,
+                over >= 16 
+                # batsman_balls_faced_this_match==1, 
+                # year %in% c("2017", "2018", "2019", "20172018", "20182019", "2020", "20192020")
+  ) %>% 
+  distinct(match_id, bucketed_over) %>% 
+  left_join(cricjam_data_big %>% 
+              dplyr::filter(inning == 2, 
+              							resources>10,
+                            !is.infinite(required_run_rate), 
+                            # over < 16,
+                over >= 16 ), by = c("match_id" = "match_id", 
+                                               "bucketed_over" = "bucketed_over")) %>% 
+  group_by(bucketed_over, match_id) %>% 
+  mutate(starting_rrr = first(required_run_rate), 
+         ending_rrr = last(required_run_rate), 
+         diff = ending_rrr-starting_rrr, 
+  			 batsman_partner_Pollard = if_else(batsman_id == 230559, "Batsman", "Partner")) %>% 
+  group_by(bucketed_over, batsman_partner_Pollard) %>% 
+  summarise(avg_Pollard = mean(diff)) %>% 
+  ungroup()) %>% 
+	mutate(batsman_name = "Pollard", 
+				 avg = avg_Pollard, 
+				 batsman_partner = batsman_partner_Pollard) %>% 
+	dplyr::select(c(batsman_name, avg, batsman_partner, bucketed_over))
+
+pollard %>% 
+	bind_rows(maxwell, dhoni, russell) %>%
+	# dplyr::filter(bucketed_over <20) %>% 
+	ggplot(aes(x = bucketed_over, y = avg, group = batsman_partner, color = batsman_partner)) +
+  geom_smooth()+
+  geom_hline(yintercept = 0)+
+  labs(y = "Required Run Rate", 
+       x = "Over", 
+       title = "Change in Required Run Rate While Batting (or Partnering)", 
+  		 color = "")+
+  facet_wrap(~batsman_name)+
+  theme_minimal()
+
+```
+
+What we can tell is that in second inning chases, the required run rate for Dhoni is lower than our other top middle order batsmen, contrasting our inital findings.
+One would assume that this was due to Dhoni's partners, but it appears he still performs at an elite level

### 2020-10-02 paper_rob_working.Rmd
push
push

--- 2020-10-02 paper_rob_working-.Rmd
@@
   html_document:
-    toc: true
-    toc_float: true
-    keep_md: true
+    toc: yes
+    toc_float: yes
+    keep_md: yes
+  word_document:
+    toc: yes
 bibliography: references.bib
@@
 
-```{r setup,message=FALSE, include=FALSE}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(message = TRUE, warning = TRUE)
@@
 
-The importance of reproducibility, and the related issue of open access to data, has received a lot of recent attention.\^[why this not work] Momentum on these issues is gathering in the sports analytics community.
+The importance of reproducibility, and the related issue of open access to data, has received a lot of recent attention.
+Momentum on these issues is gathering in the sports analytics community.
 While cricket is the second largest commercial sport in the world, unlike other popular international sports, there has been no mechanism for the public to access comprehensive statistics on players and teams.
@@
 
-```{r}
+```{r, echo=FALSE, warning=FALSE}
 library(tidyverse)
@@
 
-```{r}
-
-# insert above plots here
-```
-
 ## Using waRne to teach undergraduate statistics
@@
 In cricket, the winner of the coin toss gets to decide if they want to bat first and thus set the total that the opposition needs to pass by a single run to win, or if they want to bowl first and thus chase down the total set.
-Deciding what to do after winning a coin toss has proved to be a popular media piece in recent years\\footnote{<https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss> }\\footnote{https[://www.espncricinfo.com/video/clip/\_/id/23230682%5D](https://www.espncricinfo.com/video/clip/_/id/23230682%5D){.uri}}
-
-\\footnote{[https://www.espncricinfo.com/story/\_/id/21489056/stuart-wark-cricket-move-away-coin-toss](https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss?%5D)}
-
-\^[[https://www.espncricinfo.com/story/\_/id/20429499/lehmann-backs-scrapping-toss]](https://www.espncricinfo.com/story/_/id/20429499/lehmann-backs-scrapping-toss%5D) \^[[https://www.espncricinfo.com/story/\_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams]](https://www.espncricinfo.com/story/_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams%5D) \^[<https://www.forbes.com/sites/tristanlavalette/2018/08/20/are-cricket-matches-being-decided-by-the-luck-of-a-coin-toss/#735456837eff>]
-
-\^[<https://www.statsinsider.com.au/bbl/how-important-is-winning-the-toss-in-the-big-bash-league>]
-
-\^[<https://www.espncricinfo.com/story/_/id/18568387/tim-wigmore-how-batting-second-become-more-fruitful-more-popular>]
+Deciding what to do after winning a coin toss has proved to be a popular media piece in recent years^[https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss]
+^[https://www.espncricinfo.com/video/clip/\_/id/23230682%5D(https://www.espncricinfo.com/video/clip/_/id/23230682%5D]
+^[https://www.espncricinfo.com/story/\_/id/21489056/stuart-wark-cricket-move-away-coin-toss](https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss?%5D]
+^[https://www.espncricinfo.com/story/\_/id/20429499/lehmann-backs-scrapping-toss]](https://www.espncricinfo.com/story/_/id/20429499/lehmann-backs-scrapping-toss%5D] ^[https://www.espncricinfo.com/story/\_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams]
+^[https://www.espncricinfo.com/story/_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams%5D] ^[https://www.forbes.com/sites/tristanlavalette/2018/08/20/are-cricket-matches-being-decided-by-the-luck-of-a-coin-toss/#735456837eff]
+^[https://www.statsinsider.com.au/bbl/how-important-is-winning-the-toss-in-the-big-bash-league]
+^[https://www.espncricinfo.com/story/_/id/18568387/tim-wigmore-how-batting-second-become-more-fruitful-more-popular]
 
@@
 
-```{r}
+```{r, echo=FALSE, warning=FALSE}
 
@@
 Instead of just looking at this graphically, fans of crickets and teachers of undergraduate statistics might use the dataset as a ''biased coin'' example.
-So instead of asking, given a proportion of heads over \$n\$ tosses instructors could ask is a team better off batting first or second.
-
-```{r}
+So instead of asking, given a proportion of heads over $n$ tosses instructors could ask is a team better off batting first or second.
+
+```{r, echo=FALSE, warning=FALSE}
 
@@
 
-```{r}
+```{r, echo=FALSE, warning=FALSE}
 df<-data.frame(
@@
 <https://bookdown.org/content/922/james-stein.html>
-
-```{r}
-```
 
@@
        y = "Average Change in Required Run Rate", 
-       title = "Top Opening Batsmen T20")
+       title = "Top Opening Batsmen T20")+
   theme_minimal()
@@
 cricjam_data_big %>% 
-	dplyr::filter(batsman_id == 28081, 
-								batsman_balls_faced_this_match == 1) %>%
-	group_by(bucketed_over) %>% 
-	summarise(total = n()) %>% 
-	ungroup() %>% 
-	mutate(mean(total))
-
-cricjam_data_big %>% 
     dplyr::filter(batsman_id %in% middle_batsman_of_interest, 

### 2020-10-02 James_paper_copy.Rmd
rpo plots

@@
   drop_na()%>% 
-  mutate(pace_spin = bowling_style, "Right-arm fast" = "Pace",
-                                "Slow left-arm orthodox"= "Spin",
-                                "Legbreak"= "Spin",
-                                "Right-arm medium"= "Pace",
-                                "Left-arm medium-fast"= "Pace",
-                                "Right-arm offbreak"= "Spin",
-                                "Legbreak googly"= "Spin",
-                                "Left-arm fast"= "Pace",
-                                "Right-arm medium-fast"= "Pace",
-                                "Left-arm fast-medium"= "Pace",
-                                "Right-arm fast-medium"= "Pace",
-                                "Left-arm medium"= "Pace",
-                                "Slow left-arm chinaman"= "Spin",
-                                "Right-arm offbreak, Legbreak"= "Spin",
-                                "Right-arm medium, Right-arm offbreak"= "Spin",
-                                "Left-arm medium, Slow left-arm orthodox, 
-                                               Slow left-arm chinaman"= "Spin", 
-                                "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" = "Spin")
+  mutate(pace_spin = case_when(bowling_style == "Right-arm fast" ~ "Pace",
+                                bowling_style == "Slow left-arm orthodox"~ "Spin",
+                                bowling_style == "Legbreak"~ "Spin",
+                                bowling_style == "Right-arm medium"~ "Pace",
+                                bowling_style == "Left-arm medium-fast"~ "Pace",
+                                bowling_style == "Right-arm offbreak"~ "Spin",
+                                bowling_style == "Legbreak googly"~"Spin",
+                                bowling_style == "Left-arm fast"~ "Pace",
+                                bowling_style == "Right-arm medium-fast"~ "Pace",
+                                bowling_style == "Left-arm fast-medium"~ "Pace",
+                                bowling_style == "Right-arm fast-medium"~ "Pace",
+                                bowling_style == "Left-arm medium"~ "Pace",
+                                bowling_style == "Slow left-arm chinaman"~ "Spin",
+                                bowling_style == "Right-arm offbreak, Legbreak"~ "Spin",
+                                bowling_style == "Right-arm medium, Right-arm offbreak"~ "Spin",
+                                bowling_style == "Left-arm medium, Slow left-arm orthodox, 
+                                               Slow left-arm chinaman"~ "Spin", 
+                                bowling_style == "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" ~ "Spin"))
 
@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE, echo = FALSE, error=FALSE, results=FALSE}
 # graph of powerplay
@@
        title = "Top Opening Batsmen T20")
-  theme(axis.text = element_text(size = 12), 
-        axis.title = element_text(size = 16),
-        plot.title = element_text(size = 16))
 
@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE, echo = FALSE, error=FALSE, results=FALSE}
 # middle overs (enter in the middle)
@@
   theme_minimal()
-  theme(axis.text = element_text(size = 12), 
-        axis.title = element_text(size = 16),
-        plot.title = element_text(size = 16))
 
@@
 
-```{r}
-
-cricjam_data_big %>%
-	dplyr::filter(batsman_id == 28081, 
-								batsman_balls_faced_this_match == 1) %>%
-	group_by(bucketed_over) %>% 
-	summarise(total = n()) %>% 
-	ungroup() %>% 
-	mutate(mean(total))
+```{r, message=FALSE, warning=FALSE, echo = FALSE, error=FALSE, results=FALSE}
+
+# cricjam_data_big %>%
+# 	dplyr::filter(batsman_id == 28081, 
+# 								batsman_balls_faced_this_match == 1) %>%
+# 	group_by(bucketed_over) %>% 
+# 	summarise(total = n()) %>% 
+# 	ungroup() %>% 
+# 	mutate(mean(total))
 
@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE, echo = FALSE, error=FALSE, results=FALSE}
 
@@
   facet_wrap(~batsman_name)+
-  theme_minimal()+
-  theme(axis.text = element_text(size = 12), 
-        axis.title = element_text(size = 16),
-        plot.title = element_text(size = 16))
+  theme_minimal()
 
@@
 One would assume that this was due to Dhoni's partners, but it appears he still performs at an elite level
+
+## RPO
+
+```{r}
+
+rpo_players <- c(
+	# 230559, #pollard
+	# 34102, #rohit sharma
+	33335, #suresh raina
+	28235, #siihkar dhawan
+  # 32498, #yusef pathan
+  # 321777, #david miller
+  # 44936, #AB De Villiers
+  # 276298, #andre russell
+  # 8180, #shane watson
+  34102, #rohit sharma
+  # 51880, #gayle
+  # 325012, #stoinis
+  # 272450, #MR Marsh
+  # 28081, #Dhoni
+  253802, #kholi
+  # 5334, #finch
+  219889 #warner
+  # 326434, #carey
+  # 267192, #smith
+  # 505120, #agar
+  # 530011, #head
+  # 5961, #henriques
+  # 325026 #maxwell
+  # 25913, #nabu
+  # 4864, #christian,
+  # 270484, #faulkner
+  # 230193, #wade
+  # 233680, #wells
+  # 500268, #turner
+  # 334337, #handscomb
+  # 603410, #mcdermott
+  # 308798  #short
+  # 787987, #labuschagne
+  # 1124282, # josh phillipe
+  # 326637, #Chris Lyrnn
+  # 5939, #hussey
+  # 931581 #pant
+  # 308967, # job buttler
+  # 249866  # alex hales
+)
+
+cricjam_data_big %>% 
+	dplyr::filter(batsman_id %in% rpo_players, 
+								!is.na(pace_spin)) %>% 
+	group_by(pace_spin, batsman_name) %>% 
+	mutate(rpo = sum(runs)/n()*6) %>% 
+	dplyr::summarise(mean_rpo = mean(rpo)) %>% 
+	ggplot(aes(x = mean_rpo, y = batsman_name, color = pace_spin))+
+	geom_point() +
+	theme_minimal()+
+	labs(x = "RPO", 
+			 title = "RPO By Bowling Style", 
+			 y = "", 
+			 color = "")
+
+```
+
+```{r}
+
+```
 

### 2020-10-02 James_paper_copy.Rmd
variance is still wrong, but I think I know what to do

@@
 
+We recreated this example from the James-Stein Bookdown to show how cricket is similar to baseball in it's utility for teaching a wide range of statistical concepts.
+With a new open source data set available, the statistical community and fans have a rich new data set to work with for both teaching and as fans.
+
+Take the James-Stein estimator as an example.
+We choose batsmen who have face d aminimum of 400 balls in the 2019 and 2019/2020 seasons, and using their final strike rate as the "true" strike rate, we estmate at various points in the season what their final strike rate will be.
+Since T20 cricket is played on the international stage, and at various domestic tournaments which players from around the world are allowed to play in, we will just choose the 10%, 25%, and 50% of their totals balls faced for the season as these estimation points.
+
 ```{r}
+
+players <-
+	cricjam_data_big %>% 
+	group_by(batsman_id, batsman_name) %>% 
+	# dplyr::filter(year %in% c("2019", "20192020")) %>%
+	mutate(balls_faced = n()) %>% 
+	dplyr::filter(balls_faced > 2500) %>% 
+	distinct() %>% 
+	# summarise(total = n())
+	pull(batsman_id)
+
+truth <- 
+	cricjam_data_big %>% 
+  filter(
+  	# year %in% c("2019", "20192020"), 
+  			 batsman_id %in% players) %>% 
+  group_by(batsman_id) %>% 
+  summarise(true_sr = 100*sum(extra_runs+earned_runs)/n(), 
+              count_full = n()) %>% 
+  select(batsman_id,true_sr, count_full)
+
+set.seed(42)
+obs <- 
+	cricjam_data_big %>% 
+  filter(
+  	# year %in% c("2019", "20192020"), 
+  			 batsman_id %in% players) %>% 
+  group_by(batsman_id) %>%
+  do(sample_frac(., 0.1))  %>%       
+  summarise(mle_10 = 100*sum(extra_runs+earned_runs)/n(), 
+  					count_10 = n()) %>% 
+  select(batsman_id,mle_10, count_10) %>% 
+  inner_join(truth,by="batsman_id") %>% 
+	left_join(cricjam_data_big %>% 
+  filter(
+  	# year %in% c("2019", "20192020"), 
+  			 batsman_id %in% players) %>% 
+  group_by(batsman_id) %>%
+  do(sample_frac(., 0.25))  %>%       
+  summarise(mle_25 = 100*sum(extra_runs+earned_runs)/n(), 
+  					count_25 = n()) %>% 
+  select(batsman_id,mle_25, count_25), by = "batsman_id") %>% 
+	left_join(cricjam_data_big %>% 
+  filter(
+  	# year %in% c("2019", "20192020"), 
+  			 batsman_id %in% players) %>% 
+  group_by(batsman_id) %>%
+  do(sample_frac(., 0.50))  %>%       
+  summarise(mle_50 = 100*sum(extra_runs+earned_runs)/n(), 
+  					count_50 = n()) %>% 
+  select(batsman_id,mle_50, count_50), by = "batsman_id")
+
+p_10_=mean(obs$mle_10)
+p_25_=mean(obs$mle_25)
+p_50_=mean(obs$mle_50)
+N = length(obs$mle_25)
+
+# 10%
+count_10 <- obs %>% summarise(median(count_10)) %>% pull()
+
+obs %>% 
+  mutate(sigma2 = sum((mle_10-p_10_)^2)/count_10,
+         JS=p_10_+(1-((N-3)*sigma2/(sum((mle_10-p_10_)^2))))*(mle_10-p_10_)) %>% 
+  select(-count_10,-sigma2) %>% 
+	dplyr::select(batsman_id, true_sr, mle_10, JS) %>% 
+  gather(type,value,2:4) %>% 
+  mutate(is_truth=+(type=="true_sr")) %>% 
+  mutate(type = factor(type, levels = c("true_sr","JS","mle_10"))) %>%
+  arrange(batsman_id, type) %>% 
+  ggplot(aes(x=value,y=type))+
+  geom_path(aes(group=batsman_id),lty=2,color="lightgrey")+
+  scale_color_manual(values=c("lightgrey", "skyblue"))+ #truth==skyblue
+  geom_point(aes(color=factor(is_truth)))+
+  guides(color=FALSE)+ #remove legend
+  labs(title="The James-Stein Estimator Shrinks the MLE")+
+  theme_minimal()
+
+errors <- 
+	obs %>% 
+  mutate(sigma2 = sum((mle_10-p_10_)^2)/count_10,
+         JS=p_10_+(1-((N-3)*sigma2/(sum((mle_10-p_10_)^2))))*(mle_10-p_10_)) %>% 
+  mutate(mle_pred_error_i = (mle_10-true_sr)^2,
+         js_pred_error_i = (JS-true_sr)^2) %>% 
+  summarise(js_pred_error = sum(js_pred_error_i),
+            mle_pred_error = sum(mle_pred_error_i))
+
+# 25%
+count_25 <- obs %>% summarise(median(count_25)) %>% pull()
+	obs %>% 
+  mutate(sigma2 = sum((mle_25-p_25_)^2)/count_25,
+  			 c = (1-((N-3)*sigma2/(sum((mle_25-p_25_)^2)))),
+         JS=p_25_+c*(mle_25-p_25_)) %>% 
+  # select(-count_25,-sigma2) %>% 
+	dplyr::select(batsman_id, true_sr, mle_25, JS) %>% 
+  gather(type,value,2:4) %>% 
+  mutate(is_truth=+(type=="true_sr")) %>% 
+  mutate(type = factor(type, levels = c("true_sr","JS","mle_25"))) %>%
+  arrange(batsman_id, type) %>% 
+  ggplot(aes(x=value,y=type))+
+  geom_path(aes(group=batsman_id),lty=2,color="lightgrey")+
+  scale_color_manual(values=c("lightgrey", "skyblue"))+ #truth==skyblue
+  geom_point(aes(color=factor(is_truth)))+
+  guides(color=FALSE)+ #remove legend
+  labs(title="The James-Stein Estimator Shrinks the MLE")+
+  theme_minimal()
+
+	errors_25 <- 
+	obs %>% 
+  mutate(sigma2 = sum((mle_25-p_25_)^2)/count_25,
+         JS=p_25_+(1-((N-3)*sigma2/(sum((mle_25-p_25_)^2))))*(mle_25-p_25_)) %>% 
+  mutate(mle_pred_error_i = (mle_25-true_sr)^2,
+         js_pred_error_i = (JS-true_sr)^2) %>% 
+  summarise(js_pred_error = sum(js_pred_error_i),
+            mle_pred_error = sum(mle_pred_error_i))
+
+	# 50%
+count_50 <- obs %>% summarise(median(count_50)) %>% pull()
+	obs %>% 
+  mutate(sigma2 = sum((mle_50-p_50_)^2)/count_50,
+  			 c = (1-((N-3)*sigma2/(sum((mle_50-p_50_)^2)))),
+         JS=p_50_+c*(mle_50-p_50_)) %>% 
+  # select(-count_25,-sigma2) %>% 
+	dplyr::select(batsman_id, true_sr, mle_50, JS) %>% 
+  gather(type,value,2:4) %>% 
+  mutate(is_truth=+(type=="true_sr")) %>% 
+  mutate(type = factor(type, levels = c("true_sr","JS","mle_50"))) %>%
+  arrange(batsman_id, type) %>% 
+  ggplot(aes(x=value,y=type))+
+  geom_path(aes(group=batsman_id),lty=2,color="lightgrey")+
+  scale_color_manual(values=c("lightgrey", "skyblue"))+ #truth==skyblue
+  geom_point(aes(color=factor(is_truth)))+
+  guides(color=FALSE)+ #remove legend
+  labs(title="The James-Stein Estimator Shrinks the MLE")+
+  theme_minimal()
+
+	errors_50 <- 
+	obs %>% 
+  mutate(sigma2 = sum((mle_50-p_50_)^2)/count_50,
+         JS=p_50_+(1-((N-3)*sigma2/(sum((mle_50-p_50_)^2))))*(mle_50-p_50_)) %>% 
+  mutate(mle_pred_error_i = (mle_50-true_sr)^2,
+         js_pred_error_i = (JS-true_sr)^2) %>% 
+  summarise(js_pred_error = sum(js_pred_error_i),
+            mle_pred_error = sum(mle_pred_error_i))
+
+```
+
+```{r}
+
+library(Lahman)
+set.seed(42)
+
+# 1.
+# Select 30 players that don't have missing values
+# and have a good amount of data (total hits > 500)
+players <- Batting %>% 
+  group_by(playerID) %>% 
+  summarise(AB_total = sum(AB), 
+            H_total = sum(H)) %>% 
+  na.omit() %>% 
+  filter(H_total>500) %>%  # collect players with a lot of data
+  sample_n(size=30) %>%    # randomly sample 30 players
+  pull(playerID)
+
+TRUTH <- Batting %>% 
+  filter(playerID %in% players) %>% 
+  group_by(playerID) %>% 
+  summarise(AB_total = sum(AB), 
+            H_total = sum(H)) %>% 
+  mutate(TRUTH = H_total/AB_total) %>% 
+  select(playerID,TRUTH)
+
+set.seed(42)
+obs <- Batting %>% 
+  filter(playerID %in% players) %>% 
+  group_by(playerID) %>%
+  do(sample_n(., 5))  %>%       
+  group_by(playerID) %>% 
+  summarise(AB_total = sum(AB), 
+            H_total = sum(H)) %>% 
+  mutate(MLE = H_total/AB_total) %>% 
+  select(playerID,MLE,AB_total) %>% 
+  inner_join(TRUTH,by="playerID")
+
+p_=mean(obs$MLE)
+N = length(obs$MLE)
+obs %>% summarise(median(AB_total))
+
+sim <- tibble(player = letters, 
+              AB_total=300, 
+              TRUTH = rbeta(100,300,n=26),
+              hits_list = map(.f=rbinom,TRUTH,n=300,size=1)
+)
+bats_sim <- sim %>% 
+  unnest(hits_list) %>% 
+  group_by(player) %>% 
+  summarise(Hits = sum(hits_list)) %>% 
+  inner_join(sim, by = "player") %>% 
+  mutate(MLE = Hits/AB_total) %>% 
+  select(-hits_list)
+bats_sim
+
+bats_sim <- bats_sim %>% 
+  mutate(sigma2 = (p_*(1-p_))/300,
+         JS=p_+(1-((N-3)*sigma2/(sum((MLE-p_)^2))))*(MLE-p_)) %>% 
+  select(-AB_total,-Hits,-sigma2)
+
+bats_sim %>% 
+  gather(type,value,2:4) %>% 
+  mutate(is_truth=+(type=="TRUTH")) %>% 
+  mutate(type = factor(type, levels = c("TRUTH","JS","MLE"))) %>%
+  arrange(player, type) %>% 
+  ggplot(aes(x=value,y=type))+
+  geom_path(aes(group=player),lty=2,color="lightgrey")+
+  scale_color_manual(values=c("lightgrey", "skyblue"))+ #truth==skyblue
+  geom_point(aes(color=factor(is_truth)))+
+  guides(color=FALSE)+ #remove legend
+  labs(title="The James-Stein Estimator Shrinks the MLE")+
+  theme_minimal()
 

### 2020-10-02 paper_rob_working_good.Rmd

--- 2020-10-02 paper_rob_working.Rmd
@@
   drop_na()%>% 
-  mutate(bowling_style = recode(bowling_style, "Right-arm fast" = "Pace",
-                                "Slow left-arm orthodox"= "Spin",
-                                "Legbreak"= "Spin",
-                                "Right-arm medium"= "Pace",
-                                "Left-arm medium-fast"= "Pace",
-                                "Right-arm offbreak"= "Spin",
-                                "Legbreak googly"= "Spin",
-                                "Left-arm fast"= "Pace",
-                                "Right-arm medium-fast"= "Pace",
-                                "Left-arm fast-medium"= "Pace",
-                                "Right-arm fast-medium"= "Pace",
-                                "Left-arm medium"= "Pace",
-                                "Slow left-arm chinaman"= "Spin",
-                                "Right-arm offbreak, Legbreak"= "Spin",
-                                "Right-arm medium, Right-arm offbreak"= "Spin",
-                                "Left-arm medium, Slow left-arm orthodox, 
-                                               Slow left-arm chinaman"= "Spin", 
-                                "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" = "Spin"))
+  mutate(pace_spin = case_when(bowling_style == "Right-arm fast" ~ "Pace",
+                                bowling_style == "Slow left-arm orthodox"~ "Spin",
+                                bowling_style == "Legbreak"~ "Spin",
+                                bowling_style == "Right-arm medium"~ "Pace",
+                                bowling_style == "Left-arm medium-fast"~ "Pace",
+                                bowling_style == "Right-arm offbreak"~ "Spin",
+                                bowling_style == "Legbreak googly"~"Spin",
+                                bowling_style == "Left-arm fast"~ "Pace",
+                                bowling_style == "Right-arm medium-fast"~ "Pace",
+                                bowling_style == "Left-arm fast-medium"~ "Pace",
+                                bowling_style == "Right-arm fast-medium"~ "Pace",
+                                bowling_style == "Left-arm medium"~ "Pace",
+                                bowling_style == "Slow left-arm chinaman"~ "Spin",
+                                bowling_style == "Right-arm offbreak, Legbreak"~ "Spin",
+                                bowling_style == "Right-arm medium, Right-arm offbreak"~ "Spin",
+                                bowling_style == "Left-arm medium, Slow left-arm orthodox, 
+                                               Slow left-arm chinaman"~ "Spin", 
+                                bowling_style == "Left-arm medium, Slow left-arm orthodox, Slow left-arm chinaman" ~ "Spin"))
 
@@
 By making data more accessible and more advanced metrics more accessible fans data journalism in sports has grown in recent years.
-For example in [@horowitz2017nflscrapr] has enabled EPA to enter popular discussion among fans.
+For example in [@horowitz2017nflscrapr] has enabled EPA to enter popular discussion among fans
+^[https://twitter.com/minakimes/status/1205591497497817089]
+^[https://twitter.com/KevinColePFF/status/1277974109175951368].
 
 Cricket is the second largest sport in the world.
-However, unfortunetaly there is no easy accessible way to access ball by ball data nor aggregated statistics of teams and players.
+However, unfortunately there is no easy accessible way to access ball by ball data nor aggregated statistics of teams and players.
 Data while available on sites like [espncricinfo](https://www.espncricinfo.com/) are not in an easy to use form.
 For example, each match is listed on different webpages so hours upon hours of time would be required to copy and paste a single season, not to mention the added difficulty of linking players between games and competitions in different countries.
-Hence, there are significant logistical barriers for prospective fans and analysts studying the game, which stagnats understanding of cricket.
+Hence, there are significant logistical barriers for prospective fans and analysts studying the game, which stagnates understanding of cricket.
 
 This paper describes the waRne package, the first to provide free and easy access to data for cricket for fans.
-Web scraping tools are avaiable for fans to easily scrape the play by play commentary data on espncricinfo.
+Web scraping tools are available for fans to easily scrape the play by play commentary data on [espncricinfo](https://www.espncricinfo.com/).
 For the first time fans can evaluate their favourite teams and players and do so in a reproducible and accessible manner.
-We hope that this package can be used for fans to better understand the game, for teachers to use for interesting examples in class and for analysts in clubland who might not have access to ball by ball data.
+We hope that this package can be used for fans to better understand the game, for teachers to use for interesting examples in class and for analysts in clubs who might not have access to ball by ball data.
 
@@
 
-Data accessibility enables fans, analyts and researchers to better understand the game. Through being able to reproduce common popular metrics, visualisations and article findings.
+Data accessibility enables fans, analysts and researchers to better understand the game. Through being able to reproduce common popular metrics, visualisations and article findings.
 
 Through being able to reproduce, fans are able to make accessible findings for others and importantly they are able to extend and grow concepts.
-Unfortuntely what we see through leading cricket analytics providers is a track record of confusing output for fans.
-This can lead to lower engagement and dismissial of cricket analytics.
-
-For example in this [series of tweets](https://twitter.com/benjonescricket/status/1219898578719723520) we see the narrative being pushed that Steve Smith is a good player vs pace bowling unfortunetly just a few months prior the same company and journalist published an article which had Steve Smith doing much worse against pace (balls above 140km/h).
-Unlike a similar sport baseball, fans have no easily accisible way of seeing if Steve Smith vs pace is a strength as alluded to in the original tweet, or a weakness like the same persons published online article.
-In comparsion, fans are able to get a breakdown of Mike Trout vs fastballs from using [baseballr](https://billpetti.github.io/2018-02-19-build-statcast-database-rstats/) which provides access through statcast data from [baseballsavant](https://baseballsavant.mlb.com/).
-
-Unfortunetly this is just one of many examples whereby a relatively simple statistic is provided by media and fans have no mechanism to fact check.
-Fact checking is an important avenue for fans to not only engage and understand statistics, but having this mechanism also stops analytics people/companies from putting out misleading conclusions and findings.
-
-However though using waRne we are able to not only compare Steve Smiths performance vs bowling type but we are able to easily compare Steve Smiths performance vs bowling types with other players.
-
-### Steve Smith vs Bowling Type Chart
-
-Fans of cricket and Australian cricket especially might be interested to see if this is a statistical qwerk of small samples or if Steve Smith generally does perform as an elite cricketer vs pace bowling.
-Teachers of statistics classes might use this as an interesting example to introduce [tidyverse](https://www.tidyverse.org/learn/) principals [@wickham2019welcome].
+Unfortunately what we see through leading cricket analytics providers is a track record of confusing output for fans.
+This can lead to lower engagement and dismissal of cricket analytics.
+
+For example in this [series of tweets](https://twitter.com/benjonescricket/status/1219898578719723520) we see the narrative being pushed that Steve Smith is a good player vs pace bowling unfortunately just a few months prior the same company and journalist published an article which had Steve Smith doing much worse against pace (balls above 140km/h).
+Unlike a similar sport baseball, fans have no easily accessible way of seeing if Steve Smith vs pace is a strength as alluded to in the original tweet, or a weakness like the same persons published online article.
+In comparison, fans are able to get a breakdown of Mike Trout vs fastballs from using [baseballr](https://billpetti.github.io/2018-02-19-build-statcast-database-rstats/) which provides access through [statcast](http://m.mlb.com/glossary/statcast) data from [baseballsavant](https://baseballsavant.mlb.com/).
+
+Another barrier for fans to engage in analytics is the inability to look up insights around their favourite players, for cricket many fans have to take insights as standalone facts, without the ability to explore other players who might be of interest.
+For example after seeing this [tweet](https://twitter.com/cricvizanalyst/status/1311260989267087361) fans might want to know how their favourite player goes against various types of bowling.
+This is in stark contrast to other major sports like the NFL whereby some of the leading analytical writers make their insights more accessible by sharing code^[https://www.opensourcefootball.com/posts/2020-09-09-creating-an-expected-field-goal-metric/] and data sources.^[https://twitter.com/ChiefsAnalytics/status/1310758685325455360]
+^[https://twitter.com/ChiefsAnalytics/status/1307839100896948230]
+^[https://twitter.com/benbbaldwin/status/1311344809823657985].
+
+Without accessible data and a mechanism to get the data easily into peoples hands, how can cricket experience the growth that the NFL has throughout the past few years^[https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics].
+Through waRne, we hope to get the ball rolling on changing things for the cricket community.
+Fans of cricket might be interested to see if their own favourite players perform different against spin vs pace bowling, they might want to explore a different subset of players runs per over.
 
@@
 			 y = "Bowling Style", 
-			 title = "Steve Smith's Distribution of Runs Against Spin and Pace")
+			 title = "Steve Smith's Distribution of Runs Against Spin V Pace")
 
@@
 
-## Using waRne to teach undergraduate statistics
-
-We can also use waRne and other R packages as an easy way to engage students in learning statistical concepts.
-
-For example using the above we can not only look at the answers graphically, but we can run a statistical test.
-A common classroom example for learning the binomial distribution is to ask if a coin is based given a proportion of heads and tails given a sample size [@mcelreath2020statistical].
-Instead of asking about coins, fans of sport and cricket might be interested to ask the question; after winning the coin toss, should a team decide to bat first or bat second [@kvam2004teaching].
-
-While a coin toss is a seemingly trivial example, like in most professional sports the winner of the coin toss gets to decide what to do.
-In cricket, the winner of the coin toss gets to decide if they want to bat first and thus set the total that the opposition needs to pass by a single run to win, or if they want to bowl first and thus chase down the total set.
-Deciding what to do after winning a coin toss has proved to be a popular media piece in recent years^[https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss]
-^[https://www.espncricinfo.com/video/clip/\_/id/23230682%5D(https://www.espncricinfo.com/video/clip/_/id/23230682%5D]
-^[https://www.espncricinfo.com/story/\_/id/21489056/stuart-wark-cricket-move-away-coin-toss](https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss?%5D]
-^[https://www.espncricinfo.com/story/\_/id/20429499/lehmann-backs-scrapping-toss]](https://www.espncricinfo.com/story/_/id/20429499/lehmann-backs-scrapping-toss%5D] ^[https://www.espncricinfo.com/story/\_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams]
-^[https://www.espncricinfo.com/story/_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams%5D] ^[https://www.forbes.com/sites/tristanlavalette/2018/08/20/are-cricket-matches-being-decided-by-the-luck-of-a-coin-toss/#735456837eff]
-^[https://www.statsinsider.com.au/bbl/how-important-is-winning-the-toss-in-the-big-bash-league]
-^[https://www.espncricinfo.com/story/_/id/18568387/tim-wigmore-how-batting-second-become-more-fruitful-more-popular]
-
-Fans and analysts of the game might want to make this decision based on a simple statistic, for example they might make it based on answer these questions instead.
-
--   Do teams that bat second have a higher winning percentage than those who bat first?
-
--   Is this consistent across leagues and levels of cricket?
+```{r, echo=FALSE, warning=FALSE}
+library(tidyverse)
+
+rpo_players <- c(
+	# 230559, #pollard
+	# 34102, #rohit sharma
+	33335, #suresh raina
+	28235, #siihkar dhawan
+  # 32498, #yusef pathan
+  # 321777, #david miller
+  # 44936, #AB De Villiers
+  # 276298, #andre russell
+  # 8180, #shane watson
+  34102, #rohit sharma
+  # 51880, #gayle
+  # 325012, #stoinis
+  # 272450, #MR Marsh
+  # 28081, #Dhoni
+  253802, #kholi
+  # 5334, #finch
+  219889 #warner
+  # 326434, #carey
+  # 267192, #smith
+  # 505120, #agar
+  # 530011, #head
+  # 5961, #henriques
+  # 325026 #maxwell
+  # 25913, #nabu
+  # 4864, #christian,
+  # 270484, #faulkner
+  # 230193, #wade
+  # 233680, #wells
+  # 500268, #turner
+  # 334337, #handscomb
+  # 603410, #mcdermott
+  # 308798  #short
+  # 787987, #labuschagne
+  # 1124282, # josh phillipe
+  # 326637, #Chris Lyrnn
+  # 5939, #hussey
+  # 931581 #pant
+  # 308967, # job buttler
+  # 249866  # alex hales
+)
+
+cricjam_data_big %>% 
+	dplyr::filter(batsman_id %in% rpo_players, 
+								!is.na(pace_spin)) %>% 
+	group_by(pace_spin, batsman_name) %>% 
+	mutate(rpo = sum(runs)/n()*6) %>% 
+	dplyr::summarise(mean_rpo = mean(rpo)) %>% 
+	ggplot(aes(x = mean_rpo, y = batsman_name, color = pace_spin))+
+	geom_point() +
+	theme_minimal()+
+	labs(x = "RPO", 
+			 title = "RPO By Bowling Style", 
+			 y = "", 
+			 color = "")
 
@@
 
-Instead of just looking at this graphically, fans of crickets and teachers of undergraduate statistics might use the dataset as a ''biased coin'' example.
-So instead of asking, given a proportion of heads over $n$ tosses instructors could ask is a team better off batting first or second.
+## Using waRne for teaching statistics and [tidyverse](https://www.tidyverse.org/learn/)
+
+We can also use waRne and other R packages as an easy way to engage students in learning statistical concepts.
+
+For example using the above we can not only look at the answers graphically, but we can run a statistical test.
+A common classroom example for learning the binomial distribution is to ask if a coin is based given a proportion of heads and tails given a sample size [@mcelreath2020statistical].
+Instead of asking about coins, fans of sport and cricket might be interested to ask the question; after winning the coin toss, should a team decide to bat first or bat second [@kvam2004teaching].
+
+While a coin toss is a seemingly trivial example, like in most professional sports the winner of the coin toss gets to decide what to do.
+In cricket, the winner of the coin toss gets to decide if they want to bat first and thus set the total that the opposition needs to pass by a single run to win, or if they want to bowl first and thus chase down the total set.
+Deciding what to do after winning a coin toss has proved to be a popular media piece in recent years
+^[https://www.espncricinfo.com/story/_/id/21489056/stuart-wark-cricket-move-away-coin-toss]
+^[https://www.espncricinfo.com/video/clip/\_/id/23230682%5D]
+^[https://www.espncricinfo.com/story/\_/id/20429499/lehmann-backs-scrapping-toss]
+^[https://www.espncricinfo.com/story/\_/id/28770755/how-much-does-losing-tosses-impact-visiting-teams]
+^[https://www.forbes.com/sites/tristanlavalette/2018/08/20/are-cricket-matches-being-decided-by-the-luck-of-a-coin-toss/#735456837eff]
+^[https://www.statsinsider.com.au/bbl/how-important-is-winning-the-toss-in-the-big-bash-league]
+^[https://www.espncricinfo.com/story/_/id/18568387/tim-wigmore-how-batting-second-become-more-fruitful-more-popular]
+
+Fans and analysts of the game might want to make this decision based on a simple statistic, for example they might make it based on answer these questions instead.
+
+-   Do teams that bat second have a higher winning percentage than those who bat first?
+
+-   Is this consistent across leagues and levels of cricket?
 
@@
 One would assume that this was due to Dhoni's partners, but it appears he still performs at an elite level
+
+```{r, echo=FALSE}
+library(tidyverse)
+
+rpo_players <- c(
+	# 230559, #pollard
+	# 34102, #rohit sharma
+	33335, #suresh raina
+	28235, #siihkar dhawan
+  # 32498, #yusef pathan
+  # 321777, #david miller
+  # 44936, #AB De Villiers
+  # 276298, #andre russell
+  # 8180, #shane watson
+  34102, #rohit sharma
+  # 51880, #gayle
+  # 325012, #stoinis
+  # 272450, #MR Marsh
+  # 28081, #Dhoni
+  253802, #kholi
+  # 5334, #finch
+  219889 #warner
+  # 326434, #carey
+  # 267192, #smith
+  # 505120, #agar
+  # 530011, #head
+  # 5961, #henriques
+  # 325026 #maxwell
+  # 25913, #nabu
+  # 4864, #christian,
+  # 270484, #faulkner
+  # 230193, #wade
+  # 233680, #wells
+  # 500268, #turner
+  # 334337, #handscomb
+  # 603410, #mcdermott
+  # 308798  #short
+  # 787987, #labuschagne
+  # 1124282, # josh phillipe
+  # 326637, #Chris Lyrnn
+  # 5939, #hussey
+  # 931581 #pant
+  # 308967, # job buttler
+  # 249866  # alex hales
+)
+
+cricjam_data_big %>% 
+	dplyr::filter(batsman_id %in% rpo_players, 
+								!is.na(pace_spin)) %>% 
+	group_by(pace_spin, batsman_name) %>% 
+	mutate(rpo = sum(runs)/n()*6) %>% 
+	dplyr::summarise(mean_rpo = mean(rpo)) %>% 
+	ggplot(aes(x = mean_rpo, y = batsman_name, color = pace_spin))+
+	geom_point() +
+	theme_minimal()+
+	labs(x = "RPO", 
+			 title = "RPO By Bowling Style", 
+			 y = "", 
+			 color = "")

### 2020-10-02 CopyOfpaper_rob_working-.Rmd
cleaned up the text a bit, added the win prob for first inning tems

--- 2020-10-02 paper_rob_working-.Rmd
@@
 output: 
-  pdf_document:
-    toc: yes
   html_document:
@@
     keep_md: true
+  pdf_document:
+    toc: yes
 bibliography: references.bib
@@
 player_database <- 
-  read_csv("data/new_player_database_crickjam (2).csv") %>% 
+  read_csv("/Users/James/James/Cricket_Thesis/Data/new_player_database_crickjam.csv") %>% 
   dplyr::select(player_id, player_name, batting_style, bowling_style) %>% 
@@
   dplyr::filter(player_name != "Rohit Gurunath Sharma") %>% 
-  bind_rows(read_csv("data/new_player_database_crickjam (2).csv") %>% 
+  bind_rows(read_csv("/Users/James/James/Cricket_Thesis/Data/new_player_database_crickjam.csv") %>% 
               dplyr::filter(player_id == 34102) %>% 
@@
 cricjam_data <- 
-  read_csv("data/combined_2020_09_11.csv") %>% 
+  read_csv("/Users/James/James/Reproducibility-in-Cricket/Data/combined_2020_09_11.csv") %>% 
   dplyr::filter(!grepl("No result", match_summary)) %>%
@@
 The importance of reproducibility, and the related issue of open access to data, has received a lot of recent attention.\^[why this not work] Momentum on these issues is gathering in the sports analytics community.
-While cricket is the second largest commercial sport in the world, unlike other popular international sports, there has been no mechanism for the public to access comprehensive statistics on players and teams.
-Expert commentary currently relies heavily on data that isn't made readily accessible and this produces an unnecessary barrier for the development of an inclusive sports analytics community.
+While cricket is the second largest commercial sport in the world, unlike other popular international sports, there has been no mechanism for the public to access comprehensive data on players and teams.
+Expert commentary currently relies heavily on data that isn't made readily accessible, and this produces an unnecessary barrier for the development of an inclusive sports analytics community.
 
@@
 
-By making data more accessible and more advanced metrics more accessible fans data journalism in sports has grown in recent years.
+By making data more accessible and more advanced metrics more accessible to fans data journalism in sports has grown in recent years.
 For example in [@horowitz2017nflscrapr] has enabled EPA to enter popular discussion among fans.
@@
 
+Unless otherwis specified, examples are done using data from 2010-2020 T20I, Big Bash, IPL, and Champions League cricket.
+
 ## Why cricket needs reproducibility
@@
 
-Through being able to reproduce, fans are able to make accessible findings for others and importantly they are able to extend and grow concepts.
+Through being able to analyze the data ourselves, fans are able to make accessible findings for others and importantly they are able to extend and grow concepts.
 Unfortuntely what we see through leading cricket analytics providers is a track record of confusing output for fans.
@@
 
-For example in this [series of tweets](https://twitter.com/benjonescricket/status/1219898578719723520) we see the narrative being pushed that Steve Smith is a good player vs pace bowling unfortunetly just a few months prior the same company and journalist published an article which had Steve Smith doing much worse against pace (balls above 140km/h).
-Unlike a similar sport baseball, fans have no easily accisible way of seeing if Steve Smith vs pace is a strength as alluded to in the original tweet, or a weakness like the same persons published online article.
+For example in this [series of tweets](https://twitter.com/benjonescricket/status/1219898578719723520) we see the narrative being pushed that Steve Smith is a good player vs pace bowling.
+Unfortunetly just a few months prior the same company and journalist published an article which had Steve Smith doing much worse against pace (balls above 140km/h).
+Unlike a similar sport baseball, fans have no easily accisible way of seeing if Steve Smith vs pace is a strength as alluded to in the original tweet, or a weakness like the same persons earlier article.
 In comparsion, fans are able to get a breakdown of Mike Trout vs fastballs from using [baseballr](https://billpetti.github.io/2018-02-19-build-statcast-database-rstats/) which provides access through statcast data from [baseballsavant](https://baseballsavant.mlb.com/).
 
-Unfortunetly this is just one of many examples whereby a relatively simple statistic is provided by media and fans have no mechanism to fact check.
+Unfortunately this is just one of many examples whereby a relatively simple statistic is provided by media and fans have no mechanism to fact check.
 Fact checking is an important avenue for fans to not only engage and understand statistics, but having this mechanism also stops analytics people/companies from putting out misleading conclusions and findings.
@@
 Teachers of statistics classes might use this as an interesting example to introduce [tidyverse](https://www.tidyverse.org/learn/) principals [@wickham2019welcome].
+
+The data provided in the waRne package does not include the speed of the deliveries, but the type of ball bowled (and all available statistics and descriptions) can be taken from the personal player pages and joined into the main data set.
+A simple example would be to look at Steve Smiths performance against spin and pace bowling.
 
@@
 
-Fans and analysts of the game might want to make this decision based on a simple statistic, for example they might make it based on answer these questions instead.
-
--   Do teams that bat second have a higher winning percentage than those who bat first?
-
--   Is this consistent across leagues and levels of cricket?
+Fans and analysts of the game might want to make this decision based on a simple statistic, for example they might make it based on answer a different question.
+Do teams that bat second have a higher winning percentage than those who bat first?
+Is this consistent across leagues and levels of cricket?
+We look at this graohically first
 
@@
 	slice(1) %>% 
-	dplyr::select(batting_team_result, year) %>% 
+	dplyr::select(batting_team_result, year, match_id) %>% 
 	mutate(binary_result = if_else(batting_team_result=="win", 1, 0)) %>% 
@@
 	ggplot(aes(x = year, y = win_percentage, group = 1)) +
-	geom_line(color = 'red')
-```
-
-Instead of just looking at this graphically, fans of crickets and teachers of undergraduate statistics might use the dataset as a ''biased coin'' example.
+	geom_line(color = 'red') +
+	theme_minimal()+
+	labs(y = "Win Percentage", 
+			 x = "Year")
+```
+
+Alternatively, fans of crickets and teachers of undergraduate statistics might use the dataset as a ''biased coin'' example.
 So instead of asking, given a proportion of heads over \$n\$ tosses instructors could ask is a team better off batting first or second.
+One thing that is often brought up my analysts and fans is: is there an advantage to batting second?
+We perform a simple hypothesis test on the proportion winning teams that bat first, and found that there was no significant evidence to support the notion that teams batting first win less than 50% of games.
 
@@
 
-n<-900
-p<-0.5
-qbinom(c(0.025,0.975),size=n,prob=p)
-
-```
+dat <- 
+	cricjam_data_big %>% 
+	dplyr::filter(tournament_name %in% c("Twenty20 Big Bash", "Big Bash League"), 
+								inning == 2) %>% 
+	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
+	group_by(match_id) %>% 
+	slice(1) %>% 
+	dplyr::select(batting_team_result, year, match_id) %>% 
+	mutate(binary_result = if_else(batting_team_result=="win", 1, 0)) %>% 
+	group_by(batting_team_result) %>% 
+	summarise(win_count = n()) %>% 
+	dplyr::filter(batting_team_result != 'super-over')
+
+N = sum(dat$win_count)
+wins = dat$win_count[2]
+
+z = (wins/N-0.5)/sqrt((0.5^2)/N)
+
+out <- pnorm(z)
+
+```
+
+Probably drop this since the win prob is done now
 

### 2020-10-02 James_paper_copy.Rmd
cleaned up the text a bit, added the win prob for first inning tems

@@
 We recreated this example from the James-Stein Bookdown to show how cricket is similar to baseball in it's utility for teaching a wide range of statistical concepts.
-With a new open source data set available, the statistical community and fans have a rich new data set to work with for both teaching and as fans.
-
-Take the James-Stein estimator as an example.
-We choose batsmen who have face d aminimum of 400 balls in the 2019 and 2019/2020 seasons, and using their final strike rate as the "true" strike rate, we estmate at various points in the season what their final strike rate will be.
+With a new open source data set available, the statistical community and fans have a rich new data set to work with for any number of personal or professional projects
+
+We choose batsmen who have face a minimum of 400 balls in the 2019 and 2019/2020 seasons, and using their final strike rate as the "true" strike rate, we estmate at various points in the season what their final strike rate will be.
 Since T20 cricket is played on the international stage, and at various domestic tournaments which players from around the world are allowed to play in, we will just choose the 10%, 25%, and 50% of their totals balls faced for the season as these estimation points.
@@
 	group_by(batsman_id, batsman_name) %>% 
-	# dplyr::filter(year %in% c("2019", "20192020")) %>%
+	dplyr::filter(year %in% c("2019", "20192020")) %>%
 	mutate(balls_faced = n()) %>% 
-	dplyr::filter(balls_faced > 2500) %>% 
-	distinct() %>% 
+	dplyr::filter(balls_faced > 400) %>% 
+	distinct(batsman_id) %>% 
 	# summarise(total = n())
@@
   filter(
-  	# year %in% c("2019", "20192020"), 
+  	year %in% c("2019", "20192020"),
   			 batsman_id %in% players) %>% 
@@
 
-errors <- 
+errors_10 <- 
 	obs %>% 
@@
   mutate(sigma2 = sum((mle_50-p_50_)^2)/count_50,
-         JS=p_50_+(1-((N-3)*sigma2/(sum((mle_50-p_50_)^2))))*(mle_50-p_50_)) %>% 
+  			 c = (1-((N-3)*sigma2/(sum((mle_50-p_50_)^2)))),
+         JS=p_50_+c*(mle_50-p_50_)) %>% 
   mutate(mle_pred_error_i = (mle_50-true_sr)^2,

### 2020-10-02 CopyOfpaper_rob_working-.Rmd
added yorks

@@
 	cricjam_data_big %>% 
-	dplyr::filter(tournament_name %in% c("Twenty20 Big Bash", "Big Bash League"), 
-								inning == 2) %>% 
+	dplyr::filter(
+		# tournament_name %in% c("Twenty20 Big Bash", "Big Bash League"), 
+								inning == 1) %>% 
 	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
@@
 
+Maybe we should just skip this, I'm not confident in what I've done and we could instead just out the yorker in
+
 <https://bookdown.org/content/922/james-stein.html>
@@
 ```
+
+## The Yorker
+
+The yorker is considered the most difficult delivery to hit for a number of reason [Rob can you fill this in].
+By pulling the delivery type from the play description on espncricinfo, every time a yorker was bowled we have the players who bowled and batted.
+We can take a look at the ability to hit yorkers for the closers we compared Dhoni to, and see how they all fare against the most difficult delivery in the game.
+
+```{r}
+middle_batsman_of_interest <- c(
+  230559, #pollard
+  32498, #yusef pathan
+  321777, #david miller
+  44936, #AB De Villiers
+  276298, #andre russell
+  8180, #shane watson
+  34102, #rohit sharma
+  28081, #Dhoni
+  325026 #maxwell
+  )
+
+cricjam_data_big %>% 
+  mutate(yorker=if_else(str_detect(play_description,"york"),1,0))%>%
+  select(batsman_id, earned_runs, play_description,yorker, tournament_type, batsman_name)%>%
+  filter(batsman_id%in% middle_batsman_of_interest) %>% 
+  group_by(earned_runs, yorker, batsman_name)%>%
+  tally() %>%
+  filter(yorker==1) %>% 
+  ungroup() %>% 
+	group_by(batsman_name) %>% 
+  mutate(total = sum(n), 
+         percent = n/total, 
+  			 earned_runs = as.factor(earned_runs), 
+  			 yorker_nah = "Yorker") %>% 
+	bind_rows(cricjam_data_big %>% 
+  mutate(yorker=if_else(str_detect(play_description,"york"),1,0))%>%
+  select(batsman_id, earned_runs, play_description,yorker, tournament_type, batsman_name)%>%
+  filter(batsman_id%in% middle_batsman_of_interest) %>% 
+  group_by(earned_runs, yorker, batsman_name)%>%
+  tally() %>%
+  filter(yorker==0) %>% 
+  ungroup() %>% 
+	group_by(batsman_name) %>% 
+  mutate(total = sum(n), 
+         percent = n/total, 
+  			 earned_runs = as.factor(earned_runs), 
+  			 yorker_nah = "Non Yorker")) %>% 
+	ggplot(aes(x = percent, y = batsman_name, color = earned_runs))+
+	geom_point() +
+	labs(color = "Runs", 
+			 y = "", 
+			 x = "%")+
+	facet_wrap(~yorker_nah)
+
+```
+
+Against the yorker most batsmen seem to hit far fewer 6s, though 4s are seemingly unaffected.
+However this doesn't seem to apply to Dhoni, who appears to be hitting more boundaries when facing the yorkers.
+Perhaps this is why he is considered the greatest closer of all time.
+Or perhaps there are other reasons why he appears so much better against the yorker, which this data could help a fan discover.
 

### 2020-10-02 CopyOfpaper_rob_working-.Rmd
editted more dhoni stuff

@@
 
-With this new dataset, we allow fans to put on their analyst thats and to be able to ask themselves, what exactly makes a good closer and what exactly is the Dhoni Dilemma.
-To understand the Dhoni Deilemma, we need to see some interesting things to do with the change in RRR.
+With this new dataset, we allow fans to put on their analyst hats and to be able to ask themselves, what exactly makes a good closer and what exactly is the Dhoni Dilemma.
+To understand the Dhoni Deilemma, we need to see some interesting things to do with the change in required run rate.
 
@@
 But proportion of team runs in T20 games doesn't really make sense as a batting metric as we are not just concerned with how many runs they get, but how quickly they get them.
-So we introduce a new way to measure batsman change in RRR.
+So we measure our batsmen's performances by how well they perform when there is a known target, that is in second innings only, and how they affect their teams required run rate while at bat.
 
@@
 Dhoni is considered the greatest closing batsman in the world.
-He has pulled of many spectacular saves, and has T20 and T20I strike rates after 14 years in the game.
-But how does his required run rate compare to other middle order and closing batsman
+He has pulled of many spectacular saves, and has excellent T20 and T20I strike rates after 14 years in the game.
+But how does his required run rate compare to other middle order and closing batsman.
+We compare him to some well known middle and closing batsmen from across T20 below.
+Starting in the middle overs where he plays most often, we compare the required run rates
 
@@
 
-Dhoni is renowed for being one of the worlds best closers but we can see it seems as though in the middle he is letting the game get out of hand.
-How can one of the worlds best closers let the run rate go up so consistently in the middle overs?
+It seems as though in the middle he is letting the game get out of hand. How can one of the worlds best closers let the run rate go up so consistently in the middle overs?
 
@@
 Dhoni comes in usually around overs 11/12, and from the plot below we can see that he is a slow starter, it takes several overs for him to get his strike rate up, which consistently stays high throughout the game even when other batsmen's begins to drop off.
+It seems he takes a while to warm up, which explains why he comes in early in the second half.
 
 ```{r,message=FALSE, echo=FALSE, warning=FALSE}
-
-cricjam_data_big %>%
-	dplyr::filter(batsman_id == 28081, 
-								batsman_balls_faced_this_match == 1) %>%
-	group_by(bucketed_over) %>% 
-	summarise(total = n()) %>% 
-	ungroup() %>% 
-	mutate(mean(total))
 
@@
 
-Another is the effort of Dhoni's partners in moving the game along while Dhoni is up
+Another we considered is the effort of Dhoni's partner to keep the required run rate down while he's at the crease, but those results couldn't show us that Dhoni is elite in these overs at all
 
@@
 
+Finally we looked at the performance in the end of the game, overs 17-20.
+THe required run rate is a poor metric to use here when the resources (overs and wickets) are low, so we restrict this to game situations where the DLS resources are greater than 10.
+Here we can see that Dhoni is quickly lowering the required run rate, consistently to the end of the match.
+It appears that this is his strength, being able to at the last minute quickly lower the run near the end of the game in T20
+
 ```{r,message=FALSE,echo=FALSE, warning=FALSE}

### 2020-10-03 paper_rob_working_good.Rmd
push
push

@@
 
-## James Stein Estimator - Why not a cricket example
-
-<https://bookdown.org/content/922/james-stein.html>
+## The Yorker
+
+The yorker is considered the most difficult delivery to hit for a number of reason [Rob can you fill this in].
+By pulling the delivery type from the play description on espncricinfo, every time a yorker was bowled we have the players who bowled and batted.
+We can take a look at the ability to hit yorkers for the closers we compared Dhoni to, and see how they all fare against the most difficult delivery in the game.
+
+```{r}
+middle_batsman_of_interest <- c(
+  230559, #pollard
+  32498, #yusef pathan
+  321777, #david miller
+  44936, #AB De Villiers
+  276298, #andre russell
+  8180, #shane watson
+  34102, #rohit sharma
+  28081, #Dhoni
+  325026 #maxwell
+  )
+
+cricjam_data_big %>% 
+  mutate(yorker=if_else(str_detect(play_description,"york"),1,0))%>%
+  select(batsman_id, earned_runs, play_description,yorker, tournament_type, batsman_name)%>%
+  filter(batsman_id%in% middle_batsman_of_interest) %>% 
+  group_by(earned_runs, yorker, batsman_name)%>%
+  tally() %>%
+  filter(yorker==1) %>% 
+  ungroup() %>% 
+	group_by(batsman_name) %>% 
+  mutate(total = sum(n), 
+         percent = n/total, 
+  			 earned_runs = as.factor(earned_runs), 
+  			 yorker_nah = "Yorker") %>% 
+	bind_rows(cricjam_data_big %>% 
+  mutate(yorker=if_else(str_detect(play_description,"york"),1,0))%>%
+  select(batsman_id, earned_runs, play_description,yorker, tournament_type, batsman_name)%>%
+  filter(batsman_id%in% middle_batsman_of_interest) %>% 
+  group_by(earned_runs, yorker, batsman_name)%>%
+  tally() %>%
+  filter(yorker==0) %>% 
+  ungroup() %>% 
+	group_by(batsman_name) %>% 
+  mutate(total = sum(n), 
+         percent = n/total, 
+  			 earned_runs = as.factor(earned_runs), 
+  			 yorker_nah = "Non Yorker")) %>% 
+	ggplot(aes(x = percent, y = batsman_name, color = earned_runs))+
+	geom_point() +
+	labs(color = "Runs", 
+			 y = "", 
+			 x = "%")+
+	facet_wrap(~yorker_nah)
+
+```
 
@@
 <https://twitter.com/cricvizanalyst/status/1311260989267087361>
-
-## Cool cricket examples
-
--   Dhoni spin/pace/yorkr
-
-    -   shows off joins to player information and extracting information from play by play
-
--   espncricinfo articles but instead of ipl do same tables for big bash
-
--   recreate stats from espncricinfo like this [page](https://www.espncricinfo.com/series/19297/statistics/1187665/new-zealand-vs-england-1st-t20i-england-in-new-zealand-2019-20)
 

### 2020-10-03 CopyOfpaper_rob_working-.Rmd
coin

@@
 ```{r}
-df<-data.frame(
-             Year = c(2010L,2011L,2012L,2013L,2014L,
-                      2015L,2016L,2017L,2018L,2019L,2020L,2021L,2022L),
-  number.of.heads = c(52L,55L,51L,70L,30L,40L,35L,
-                      43L,49L,37L,39L,45L,32L)
-)
-
-library(broom)
-df %>%
-  # create variables to pass to binom.test
-  mutate(n = 100, 
-         x = n - number.of.heads, 
+
+cricjam_data_big %>% 
+	dplyr::filter(inning == 1, 
+								batting_team_result != "super-over") %>% 
+	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
+	group_by(year, batting_team_result) %>% 
+	dplyr::select(batting_team_result, match_id, year) %>% 
+	distinct() %>% 
+	summarise(total = n()) %>% 
+	pivot_wider(names_from = batting_team_result, values_from = total) %>% 
+	ungroup() %>% 
+	mutate(n =win+lose, 
 				 p = 0.5, 
 				 alternative = "two.sided", 
-         conf.level = 0.95) %>%
-  # Run a binom.test for each row/year
+				 conf.level = 0.95, 
+				 x = win) %>% 
 	mutate(result = pmap(list(x, n,p), binom.test)) %>% 
@@
   # plot the annual confidence intervals
-  ggplot(aes(factor(Year), p.value)) + 
+  ggplot(aes(factor(year), p.value)) + 
     geom_errorbar(aes(ymin=conf.low, ymax=conf.high, color = (p.value > 0.05)), width=.1) +
   theme_minimal() +
-  labs(y = "P(coin is fair)", x = "Year") +
-  ggtitle("Annual coin fairness test")
+  labs(y = "P(Batting is fair)", x = "Year", color = "p > 0.05") +
+  ggtitle("Annual 1st Inning Batting Fairness")
 
@@
 One would assume that this was due to Dhoni's partners, but it appears he still performs at an elite level
+
+## Conclusion
+
+We believe that access to the play-by-play data available to cricket analysts and teams is vital to create a strong cricket analytics community.
+Without access fans are forced to accept the statistics given by analytics teams like Cricviz, who have released contradictory statistics on players to a community without the acesss to verify the work.
+The waRne package is at the moment a tool for scraping games that are complete, and lacks the functionality to live scrape and update.
+However we believe this framework can be adapted for live scraping as well, to give fans the ultimate tools for engaging in the game of cricket.

### 2020-10-03 paper_rob_working_good.Rmd
push
push

@@
 library(patchwork)
+library(cowplot)
 
@@
 
-```{r, echo=FALSE, warning=FALSE}
-library(tidyverse)
-cricjam_data_big %>% 
+```{r, echo=FALSE, message=FALSE,warning=FALSE,out.width = "80%"}
+
+p1<-cricjam_data_big %>% 
 	dplyr::filter(batsman_id == 267192, 
@@
 			 title = "Steve Smith's Distribution of Runs Against Spin V Pace")
-
-```
-
-```{r, echo=FALSE, warning=FALSE}
-library(tidyverse)
 
@@
 
-cricjam_data_big %>% 
+p2<-cricjam_data_big %>% 
 	dplyr::filter(batsman_id %in% rpo_players, 
@@
 
-```{r, echo=FALSE, warning=FALSE}
-
-cricjam_data_big %>% 
-	dplyr::filter(tournament_name %in% c("Twenty20 Big Bash", "Big Bash League"), 
-								inning == 2) %>% 
-	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
-	group_by(match_id) %>% 
-	slice(1) %>% 
-	dplyr::select(batting_team_result, year) %>% 
-	mutate(binary_result = if_else(batting_team_result=="win", 1, 0)) %>% 
-	group_by(batting_team_result, year) %>% 
-	summarise(win_count = n()) %>% 
-	ungroup() %>% 
-	pivot_wider(names_from = batting_team_result, values_from = win_count) %>% 
-	rename(superover = 'super-over') %>% 
-	replace_na(list(superover=0)) %>% 
-	mutate(total_games = lose+win+superover, 
-				 win_percentage = win/total_games, 
-				 year = as.factor(year)) %>% 
-	ggplot(aes(x = year, y = win_percentage, group = 1)) +
-	geom_line(color = 'red')
+plot_grid(p1,p2, labels=c("",""))
+
 ```
@@
 
--   Do teams that bat second have a higher winning percentage than those who bat first?
+-   Do teams that bat second have a different winning percentage than those who bat first?
 
@@
 ```{r, echo=FALSE, warning=FALSE}
-
-n<-900
-p<-0.5
-qbinom(c(0.025,0.975),size=n,prob=p)
-
-```
-
-```{r, echo=FALSE, warning=FALSE}
-df<-data.frame(
-             Year = c(2010L,2011L,2012L,2013L,2014L,
-                      2015L,2016L,2017L,2018L,2019L,2020L,2021L,2022L),
-  number.of.heads = c(52L,55L,51L,70L,30L,40L,35L,
-                      43L,49L,37L,39L,45L,32L)
-)
-
 library(broom)
-df %>%
-  # create variables to pass to binom.test
-  mutate(n = 100, 
-         x = n - number.of.heads, 
+cricjam_data_big %>% 
+	dplyr::filter(inning == 1, 
+								batting_team_result != "super-over") %>% 
+	dplyr::filter(tournament_type == "Twenty20", 
+                tournament_name %in% c("Twenty20 Big Bash", "Indian Premier League", "Big Bash League", 
+                                       "Champions League Twenty20", "Pepsi Indian Premier League")) %>% 
+	mutate(t_name = if_else(tournament_name %in% c("Big Bash League", "Twenty20 Big Bash"), "Big Bash", "IPL")) %>% 
+	mutate(year = if_else(nchar(year)==4, year, as.numeric(str_sub(as.character(year), 1, 4)))) %>% 
+	group_by(year, batting_team_result, t_name) %>% 
+	dplyr::select(batting_team_result, match_id, year, t_name) %>% 
+	distinct() %>% 
+	summarise(total = n()) %>% 
+	pivot_wider(names_from = batting_team_result, values_from = total) %>% 
+	ungroup() %>% 
+	mutate(n =win+lose, 
 				 p = 0.5, 
 				 alternative = "two.sided", 
-         conf.level = 0.95) %>%
-  # Run a binom.test for each row/year
+				 conf.level = 0.95, 
+				 x = win) %>% 
 	mutate(result = pmap(list(x, n,p), binom.test)) %>% 
@@
   # plot the annual confidence intervals
-  ggplot(aes(factor(Year), p.value)) + 
+  ggplot(aes(factor(year), p.value)) + 
     geom_errorbar(aes(ymin=conf.low, ymax=conf.high, color = (p.value > 0.05)), width=.1) +
   theme_minimal() +
-  labs(y = "P(coin is fair)", x = "Year") +
-  ggtitle("Annual coin fairness test")
+  labs(y = "P(Batting is fair)", x = "Year", color = "p > 0.05") +
+  ggtitle("Annual 1st Inning Batting Fairness")+ 
+  opts(axis.text.x=theme_text(angle=-90))+
+	facet_wrap(~t_name)
 
@@
 
-The yorker is considered the most difficult delivery to hit for a number of reason [Rob can you fill this in].
+The yorker is considered the most difficult delivery to hit for a number of reasons .
 By pulling the delivery type from the play description on espncricinfo, every time a yorker was bowled we have the players who bowled and batted.
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 middle_batsman_of_interest <- c(
@@
 ```
-
-## Easier engagement of fans
-
-To the surprise of many, there has not been a real push to engage fans in the analytics of the game of cricket.
-Unlike other sports which has seen a boom through the use of accessible data like Ice Hockey, NFL, Baseball and Basketball for example.
-
-Without an easy accessible medium how can crickets version of an analytics community grow.
-
-something something look towards how EPA has changed the way fans are engaged in NFL analytics - so maybe the change in run rate can be like EPA?
-
-<https://twitter.com/cricvizanalyst/status/1311260989267087361>
 
@@
 
-```{r, echo=FALSE}
-library(tidyverse)
-
-rpo_players <- c(
-	# 230559, #pollard
-	# 34102, #rohit sharma
-	33335, #suresh raina
-	28235, #siihkar dhawan
-  # 32498, #yusef pathan
-  # 321777, #david miller
-  # 44936, #AB De Villiers
-  # 276298, #andre russell
-  # 8180, #shane watson
-  34102, #rohit sharma
-  # 51880, #gayle
-  # 325012, #stoinis
-  # 272450, #MR Marsh
-  # 28081, #Dhoni
-  253802, #kholi
-  # 5334, #finch
-  219889 #warner
-  # 326434, #carey
-  # 267192, #smith
-  # 505120, #agar
-  # 530011, #head
-  # 5961, #henriques
-  # 325026 #maxwell
-  # 25913, #nabu
-  # 4864, #christian,
-  # 270484, #faulkner
-  # 230193, #wade
-  # 233680, #wells
-  # 500268, #turner
-  # 334337, #handscomb
-  # 603410, #mcdermott
-  # 308798  #short
-  # 787987, #labuschagne
-  # 1124282, # josh phillipe
-  # 326637, #Chris Lyrnn
-  # 5939, #hussey
-  # 931581 #pant
-  # 308967, # job buttler
-  # 249866  # alex hales
-)
-
-cricjam_data_big %>% 
-	dplyr::filter(batsman_id %in% rpo_players, 
-								!is.na(pace_spin)) %>% 
-	group_by(pace_spin, batsman_name) %>% 
-	mutate(rpo = sum(runs)/n()*6) %>% 
-	dplyr::summarise(mean_rpo = mean(rpo)) %>% 
-	ggplot(aes(x = mean_rpo, y = batsman_name, color = pace_spin))+
-	geom_point() +
-	theme_minimal()+
-	labs(x = "RPO", 
-			 title = "RPO By Bowling Style", 
-			 y = "", 
-			 color = "")
+## Conclusion - Easier fan engagement
+
+To the surprise of many, there has not been a real push to engage fans in the analytics of the game of cricket.
+Unlike other sports which has seen a boom through the use of accessible data like Ice Hockey, NFL, Baseball and Basketball for example.
+
+Without an easy accessible medium how can crickets version of an analytics community grow.
+
+\newpage

### 2020-10-03 paper_rob_working_good.Rmd
push
push

@@
   ggtitle("Annual 1st Inning Batting Fairness")+ 
-  opts(axis.text.x=theme_text(angle=-90))+
+  theme(axis.text.x=element_text(angle=-90))+
 	facet_wrap(~t_name)
@@
 Dhoni comes in usually around overs 11/12, and from the plot below we can see that he is a slow starter, it takes several overs for him to get his strike rate up, which consistently stays high throughout the game even when other batsmen's begins to drop off.
+It seems he takes a while to warm up, which explains why he comes in early in the second half.
 
@@
 
+Finally we looked at the performance in the end of the game, overs 17-20.
+THe required run rate is a poor metric to use here when the resources (overs and wickets) are low, so we restrict this to game situations where the DLS resources are greater than 10.
+Here we can see that Dhoni is quickly lowering the required run rate, consistently to the end of the match.
+It appears that this is his strength, being able to at the last minute quickly lower the run near the end of the game in T20
+
 What we can tell is that in second inning chases, the required run rate for Dhoni is lower than our other top middle order batsmen, contrasting our inital findings.
@@
 
+We believe that access to the play-by-play data available to cricket analysts and teams is vital to create a strong cricket analytics community.
+Without access fans are forced to accept the statistics given by analytics teams like Cricviz, who have released contradictory statistics on players to a community without the acesss to verify the work.
+The waRne package is at the moment a tool for scraping games that are complete, and lacks the functionality to live scrape and update.
+However we believe this framework can be adapted for live scraping as well, to give fans the ultimate tools for engaging in the game of cricket.
+
 \newpage

### 2020-10-03 paper_rob_working_good.Rmd
push
push

@@
   pdf_document:
-    toc: yes
+    toc: no
   html_document:
@@
   dplyr::filter(player_name != "Rohit Gurunath Sharma") %>% 
-  bind_rows(read_csv("data/new_player_database_crickjam (2).csv") %>% 
+  dplyr::bind_rows(read_csv("data/new_player_database_crickjam (2).csv") %>% 
               dplyr::filter(player_id == 34102) %>% 
@@
 
-By making data more accessible and more advanced metrics more accessible fans data journalism in sports has grown in recent years.
+By making data more accessible more advanced metrics more accessible fans data journalism in sports has grown in recent years.
 For example in [@horowitz2017nflscrapr] has enabled EPA to enter popular discussion among fans
@@
 
-Data accessibility enables fans, analysts and researchers to better understand the game. Through being able to reproduce common popular metrics, visualisations and article findings.
-
+Data accessibility enables fans, analysts and researchers to better understand the game.
 Through being able to reproduce, fans are able to make accessible findings for others and importantly they are able to extend and grow concepts.
@@
 	dplyr::filter(batsman_id == 267192, 
-								!is.na(bowling_style)) %>% 
-	group_by(bowling_style, earned_runs) %>% 
+								!is.na(pace_spin)) %>% 
+	group_by(pace_spin, earned_runs) %>% 
 	summarise(total = n()) %>% 
-	group_by(bowling_style) %>% 
+	group_by(pace_spin) %>% 
 	mutate(totals = sum(total),
@@
 				 earned_runs = as.factor(earned_runs)) %>% 
-	ggplot(aes(y = bowling_style, x = percent, color = earned_runs))+
+	ggplot(aes(y = pace_spin, x = percent, color = earned_runs))+
 	geom_point() +
@@
 
-plot_grid(p1,p2, labels=c("",""))
+# plot_grid(p1,p2, labels=c("",""))
+p1
+p2
 

### 2020-10-03 paper_rob_working_good.Rmd
push
push

@@
 
+```{r, echo=FALSE, warning=FALSE,message=FALSE,out.width = "80%"}
+p1<-cricjam_data_big %>% 
+	dplyr::filter(batsman_id == 267192, 
+								!is.na(pace_spin)) %>% 
+	group_by(pace_spin, earned_runs) %>% 
+	summarise(total = n()) %>% 
+	group_by(pace_spin) %>% 
+	mutate(totals = sum(total),
+				 percent = total/totals, 
+				 earned_runs = as.factor(earned_runs)) %>% 
+	ggplot(aes(y = pace_spin, x = percent, color = earned_runs))+
+	geom_point() +
+	labs(color = "Runs", 
+			 x = "% Balls Faced", 
+			 y = "Bowling Style", 
+			 title = "Steve Smith's Distribution of Runs Against Spin V Pace")
+p1
+
+```
+
 Another barrier for fans to engage in analytics is the inability to look up insights around their favourite players, for cricket many fans have to take insights as standalone facts, without the ability to explore other players who might be of interest.
@@
 # plot_grid(p1,p2, labels=c("",""))
-p1
 p2
@@
 
-For example using the above we can not only look at the answers graphically, but we can run a statistical test.
-A common classroom example for learning the binomial distribution is to ask if a coin is based given a proportion of heads and tails given a sample size [@mcelreath2020statistical].
+For example,a common classroom example for learning the binomial distribution is to ask if a coin is based given a proportion of heads and tails given a sample size [@mcelreath2020statistical].
 Instead of asking about coins, fans of sport and cricket might be interested to ask the question; after winning the coin toss, should a team decide to bat first or bat second [@kvam2004teaching].
@@
 
-```{r, echo=FALSE, warning=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE,out.width = "80%"}
 library(broom)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = "80%"}
 middle_batsman_of_interest <- c(
@@
 
-```{r,message=FALSE, echo=FALSE, warning=FALSE}
+```{r,message=FALSE, echo=FALSE, warning=FALSE,out.width = "80%"}
 # graph of powerplay
@@
 
-```{r,message=FALSE, echo=FALSE, warning=FALSE}
+```{r,message=FALSE, echo=FALSE, warning=FALSE,out.width = "80%"}
 # middle overs (enter in the middle)
@@
 
-```{r,message=FALSE, echo=FALSE, warning=FALSE}
+```{r,message=FALSE, echo=FALSE, warning=FALSE,out.width = "80%"}
 
@@
 
-```{r, message=FALSE,echo=FALSE, warning=FALSE}
+```{r, message=FALSE,echo=FALSE, warning=FALSE,out.width = "80%"}
 
@@
 
-```{r,message=FALSE,echo=FALSE, warning=FALSE}
+```{r,message=FALSE,echo=FALSE, warning=FALSE,out.width = "80%"}
 

### 2020-10-03 paper_rob_working_good.Rmd
push
push

@@
               drop_na() %>% 
-              distinct()) %>% 
+              dplyr::distinct()) %>% 
   dplyr::filter(player_name !="Sultan Ahmed")
@@
   drop_na()%>% 
-  mutate(pace_spin = case_when(bowling_style == "Right-arm fast" ~ "Pace",
+ dplyr:: mutate(pace_spin = case_when(bowling_style == "Right-arm fast" ~ "Pace",
                                 bowling_style == "Slow left-arm orthodox"~ "Spin",
@@
 			 title = "Steve Smith's Distribution of Runs Against Spin V Pace")
-p1
+p1 +theme_minimal()
 
@@
 ^[https://twitter.com/benbbaldwin/status/1311344809823657985].
-
-Without accessible data and a mechanism to get the data easily into peoples hands, how can cricket experience the growth that the NFL has throughout the past few years^[https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics].
-Through waRne, we hope to get the ball rolling on changing things for the cricket community.
-Fans of cricket might be interested to see if their own favourite players perform different against spin vs pace bowling, they might want to explore a different subset of players runs per over.
 
@@
 # plot_grid(p1,p2, labels=c("",""))
-p2
+p2 +theme_minimal()
 
 ```
+
+Without accessible data and a mechanism to get the data easily into peoples hands, how can cricket experience the growth that the NFL has throughout the past few years^[https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics].
+Through waRne, we hope to get the ball rolling on changing things for the cricket community.
+Fans of cricket might be interested to see if their own favourite players perform different against spin vs pace bowling, they might want to explore a different subset of players runs per over.
 
@@
 
-Fans and analysts of the game might want to make this decision based on a simple statistic, for example they might make it based on answer these questions instead.
-
--   Do teams that bat second have a different winning percentage than those who bat first?
-
--   Is this consistent across leagues and levels of cricket?
+Fans of cricket, instead of being mislead by [small samples](https://www.espncricinfo.com/story/_/id/30010174/kkr-vs-rr-ipl-2020-talking-points-why-teams-bowling-first?-did-andre-russell-promotion-work^[This article, has 6 matches being considered overwhelming evidence] might want to instead do a statistical test or visualise if batting first has a difference in terms of winning percentages and is this consistent across leagues.
 
@@
 
-The yorker is considered the most difficult delivery to hit for a number of reasons .
+The yorker is considered one of most difficult deliveries to hit for batsman for a few reasons.
+Yorkers are usually bowled quicker than a bowlers average pace and aimed at the batsmans feet.
+This makes it difficult for the batsman to get ''under'' the ball and hit it over the fence for six.
 By pulling the delivery type from the play description on espncricinfo, every time a yorker was bowled we have the players who bowled and batted.
-We can take a look at the ability to hit yorkers for the closers we compared Dhoni to, and see how they all fare against the most difficult delivery in the game.
-
+As yorkers are typically bowled during the closing stages of matches, we can look at how good closers such as [Dhoni](https://www.espncricinfo.com/india/content/player/28081.html) bat against them.
 ```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = "80%"}
@@
 			 y = "", 
-			 x = "%")+
+			 x = "%")+ theme_minimal()+
 	facet_wrap(~yorker_nah)
@@
 
-## Dhoni Dilemma
-
-With this new dataset, we allow fans to put on their analyst hats and to be able to ask themselves, what exactly makes a good closer and what exactly is the Dhoni Dilemma.
-To understand the Dhoni Deilemma, we need to see some interesting things to do with the change in required run rate.
+## [Dhoni](https://www.espncricinfo.com/india/content/player/28081.html) Dilemma
+
+With access to espncricinfo data, we allow fans to put on their analyst hats and to be able to ask themselves, what exactly makes a good closer and what exactly is the Dhoni Dilemma.
 
 Cricket is an obvious sort of sport, to win in the end you have to get one run more than your opposition.
-A commonly used statistic during the broadcast is the required run rate, for example if you need 120 to win off 20 overs your RRR is \$6\$ runs an over.
-What follows is then obvious is the chasing team wins if they are able to lower the RRR that is score quicker.
-
-Espncricinfo did article on teams that win games get a higher proportion of their runs in the powerplay and first 10 overs.
-But proportion of team runs in T20 games doesn't really make sense as a batting metric as we are not just concerned with how many runs they get, but how quickly they get them.
+A commonly used statistic during the broadcast is the required run rate, for example if you need 120 to win off 20 overs your RRR is $6$ runs an over.
+What follows is then obvious, the chasing team wins if they are able to lower the RRR to 0.
+
+Espncricinfo did a piece on successful T20 chases and found that teams who score a higher proportion of their needed runs in the first 10 overs did better than those who did not^[https://www.espncricinfo.com/story/_/id/26712889/t20-chases-won-lost-first-ten-overs-not-last-ten].
+But proportion of team runs in T20 games doesn't really make sense as an individual batting metric as we are not just concerned with how many runs they get, but how quickly they get them.
 So we measure our batsmen's performances by how well they perform when there is a known target, that is in second innings only, and how they affect their teams required run rate while at bat.
@@
        y = "Average Change in Required Run Rate", 
-       title = "Top Opening Batsmen T20")+
+       title = "Top Opening Batsmen T20 (Overs 1-6)")+
   theme_minimal()
@@
 
-This graph makes sense as we can see some of the best openers in the world lower the required run rate. So what is the Dhoni Dilemma?
-
-Dhoni is considered the greatest closing batsman in the world.
-He has pulled of many spectacular saves, and has excellent T20 and T20I strike rates after 14 years in the game.
+This graph makes sense as we can see some of the best openers in the world lower the required run rate^[https://www.icc-cricket.com/rankings/mens/player-rankings/t20i]. So what is the Dhoni Dilemma?
+
+Dhoni is considered the greatest closing batsman in the world^[https://www.smh.com.au/sport/cricket/australia-to-stick-with-stoinis-in-hope-he-can-be-their-dhoni-20200905-p55sqn.html].
+He has pulled of many spectacular wins, and has excellent T20 and T20I strike rates after 14 years in the game.
 But how does his required run rate compare to other middle order and closing batsman.
@@
        y = "Average Change in Required Run Rate", 
-       title = "Top Middle Batsmen T20")+
+       title = "Top Middle Batsmen T20 overs 7-16")+
   geom_text(vjust = 1.4, color = "black") +
@@
 
-Another we considered is the effort of Dhoni's partner to keep the required run rate down while he's at the crease, but those results couldn't show us that Dhoni is elite in these overs at all
+A common theory would be that if Dhoni is lowering the required run rate himself, then during this phase of play it must be his partner who is attacking the run rate, while Dhoni aims to survive to the closing overs.
+However looking at Dhonis partners we see this is not the case.
+Thus adding to the Dhoni Dilemma.
 
@@
 
+Finally we looked at the performance in the end of the game, overs 17-20.
+The required run rate is a poor metric to use here when the resources (overs and wickets) are low, so we restrict this to game situations where the [DLS](https://en.wikipedia.org/wiki/Duckworth%E2%80%93Lewis%E2%80%93Stern_method)^[The Duckworth Lewis tables are not publically accessible, through this example we hope someone somewhere will upload the tables for T20, for this example we had to use the ODI ones] resources are greater than 10.
+Here we can see that Dhoni is quickly lowering the required run rate, consistently to the end of the match.
+It appears that this is his strength, being able to at the last minute quickly lower the run near the end of the game in T20
+
 ```{r,message=FALSE,echo=FALSE, warning=FALSE,out.width = "80%"}
@@
 
-Finally we looked at the performance in the end of the game, overs 17-20.
-THe required run rate is a poor metric to use here when the resources (overs and wickets) are low, so we restrict this to game situations where the DLS resources are greater than 10.
-Here we can see that Dhoni is quickly lowering the required run rate, consistently to the end of the match.
-It appears that this is his strength, being able to at the last minute quickly lower the run near the end of the game in T20
-
-What we can tell is that in second inning chases, the required run rate for Dhoni is lower than our other top middle order batsmen, contrasting our inital findings.
-One would assume that this was due to Dhoni's partners, but it appears he still performs at an elite level
+It is clear in the data that while Dhoni is considered by many to be the worlds best closer, he is doing it in a unique counter intuitive way.
+While many would consider lowering the required run rate to be a priority, Dhoni and his partners consistently allow the required run rate to increase, however Dhoni is able to still bring his team home.
+This creates a dilemma for coaches, analysts and fans.
+Can their favourite player be a good closer if they try to bat like Dhoni?
 
@@
 We believe that access to the play-by-play data available to cricket analysts and teams is vital to create a strong cricket analytics community.
-Without access fans are forced to accept the statistics given by analytics teams like Cricviz, who have released contradictory statistics on players to a community without the acesss to verify the work.
+Without access fans are forced to accept the statistics given by analytics teams like Cricviz, who have released contradictory statistics on players to a community without the access to verify the work.
 The waRne package is at the moment a tool for scraping games that are complete, and lacks the functionality to live scrape and update.

### 2020-10-03 paper_rob_working_good.Rmd
push
push

@@
 title: "waRne - Putting cricket analytics in a spin"
-authors: Robert(1), James (2)
+authors: "unknown 1", "unknown 2"
 output:
@@
 
-By making data more accessible more advanced metrics more accessible fans data journalism in sports has grown in recent years.
+By making data available more advanced metrics have become more accessible to  fans through quality data journalists.
 For example in [@horowitz2017nflscrapr] has enabled EPA to enter popular discussion among fans
@@
 
-```{r, echo=FALSE, warning=FALSE,message=FALSE,out.width = "80%"}
+```{r, echo=FALSE, warning=FALSE,message=FALSE,out.width = "80%",fig.align='center'}
 p1<-cricjam_data_big %>% 
@@
 
-```{r, echo=FALSE, message=FALSE,warning=FALSE,out.width = "80%"}
+```{r, echo=FALSE, message=FALSE,warning=FALSE,out.width = "80%",fig.align='center'}
 
@@
 
-Fans of cricket, instead of being mislead by [small samples](https://www.espncricinfo.com/story/_/id/30010174/kkr-vs-rr-ipl-2020-talking-points-why-teams-bowling-first?-did-andre-russell-promotion-work^[This article, has 6 matches being considered overwhelming evidence] might want to instead do a statistical test or visualise if batting first has a difference in terms of winning percentages and is this consistent across leagues.
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE,out.width = "80%"}
+Fans of cricket, instead of being mislead by [small samples](https://www.espncricinfo.com/story/_/id/30010174/kkr-vs-rr-ipl-2020-talking-points-why-teams-bowling-first?-did-andre-russell-promotion-work)^[This article, has 6 matches being considered overwhelming evidence] might want to instead do a statistical test or visualise if batting first has a difference in terms of winning percentages and is this consistent across leagues.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE,out.width = "80%",fig.align='center'}
 library(broom)
@@
   theme_minimal() +
-  labs(y = "P(Batting is fair)", x = "Year", color = "p > 0.05") +
+  labs(y = "P(Batting is fair)", x = "Year", color = "p < 0.05") +
   ggtitle("Annual 1st Inning Batting Fairness")+ 
@@
 As yorkers are typically bowled during the closing stages of matches, we can look at how good closers such as [Dhoni](https://www.espncricinfo.com/india/content/player/28081.html) bat against them.
-```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = "80%"}
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = "80%",fig.align='center'}
 middle_batsman_of_interest <- c(
@@
 Cricket is an obvious sort of sport, to win in the end you have to get one run more than your opposition.
-A commonly used statistic during the broadcast is the required run rate, for example if you need 120 to win off 20 overs your RRR is $6$ runs an over.
+A commonly used statistic during the broadcast is the required run rate (RRR), for example if you need 120 to win off 20 overs your RRR is $6$ runs an over.
 What follows is then obvious, the chasing team wins if they are able to lower the RRR to 0.
@@
 
-```{r,message=FALSE, echo=FALSE, warning=FALSE,out.width = "80%"}
+```{r,message=FALSE, echo=FALSE, warning=FALSE,out.width = "80%",fig.align='center'}
 # graph of powerplay
@@
 
-```{r,message=FALSE, echo=FALSE, warning=FALSE,out.width = "80%"}
+```{r,message=FALSE, echo=FALSE, warning=FALSE,out.width = "80%",fig.align='center'}
 # middle overs (enter in the middle)
@@
 
-It seems as though in the middle he is letting the game get out of hand. How can one of the worlds best closers let the run rate go up so consistently in the middle overs?
+It seems as though in the middle he is letting the game get out of hand. How can one of the worlds best closers let the required run rate go up so consistently in the middle overs?
 
@@
 
-```{r,message=FALSE, echo=FALSE, warning=FALSE,out.width = "80%"}
+```{r,message=FALSE, echo=FALSE, warning=FALSE,out.width = "80%",fig.align='center'}
 
@@
 
-```{r, message=FALSE,echo=FALSE, warning=FALSE,out.width = "80%"}
+```{r, message=FALSE,echo=FALSE, warning=FALSE,out.width = "80%",fig.align='center'}
 
@@
 Here we can see that Dhoni is quickly lowering the required run rate, consistently to the end of the match.
-It appears that this is his strength, being able to at the last minute quickly lower the run near the end of the game in T20
-
-```{r,message=FALSE,echo=FALSE, warning=FALSE,out.width = "80%"}
+It appears that this is his strength, being able to at the last minute quickly lower the required run rate near the end of the game in T20
+
+```{r,message=FALSE,echo=FALSE, warning=FALSE,out.width = "80%",fig.align='center'}
 

### 2020-10-03 waRne paper.Rmd
push
push

--- 2020-10-03 paper_rob_working_good.Rmd
@@
 title: "waRne - Putting cricket analytics in a spin"
-authors: "unknown 1", "unknown 2"
+authors: unknown1, unknown2
 output:
